### 2017-09-05 analysis/Simulation.Rmd
Added Simulation.Rmd.

---
title: "Simulation for Correlation Motif"
author: "Zhiwei Ma"
date: "8/30/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model framework

The framework of Correlation Motif model is based on my write-up document and supplementary matrials by $Wei$ $and$ $Ji$ (2015). In $Wei$ $and$ $Ji$ (2015), they implemented an EM algorithm and introduced priors for both $\pi$ and $Q$. As a result, their iterative formulas for new $\pi$ and new $Q$ are:

\begin{eqnarray}
\pi_k^{(t+1)} = \frac{\sum_{i=1}^n p_{ik}+1}{n+K} , \\
q_{kr}^ {(t+1)}=\frac{\sum_{i=1}^n p_{ikr1}+1}{\sum_{i=1}^n p_{ik}+2}.
\end{eqnarray}

However, one can apply EM algorithm without introducing a prior and obtain a more interpretable formula:

\begin{eqnarray}
\pi_k^{(t+1)} = \frac{1}{n} \sum_{i=1}^n p_{ik}, \\
q_{kr}^ {(t+1)}=\frac{\sum_{i=1}^n p_{ikr1}}{\sum_{i=1}^n p_{ik}}.
\end{eqnarray}

Here $p_{ik}= Pr(z_i=k|X_i,\pi^{(t)}, Q^ {(t)})$ is the posterior probability for gene $i$ belonging to class $k$, and the mean of $p_{ik}$ over $i$ is a reasonable estimations for $\pi_{k}$. The argument is similar for $q_{kr}$. Thus, we could interpret the underlying meaning in our iterative formulas.

In fact, since $G$ is relatively large and $G \gg K$, the two formulas tend to give very similar results. We will illustrate this idea by implementing these two algorithms seperately on the same dataset: $simudata2$ in R package $Cormotif$.

Notice that right now we are using the model applying limma ($Smyth$, 2004). The results with prior is

```{r,cache=TRUE,message=FALSE,warning=FALSE}
library(Cormotif)
data(simudata2)
exprs.simu2<-as.matrix(simudata2[,2:25])
data(simu2_groupid)
data(simu2_compgroup)

motif.fitted<-cormotiffit(exprs.simu2,simu2_groupid,simu2_compgroup,
                          K=1:5,max.iter=1000,BIC=TRUE)
motif.fitted$bic
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

The results without prior is

```{r,echo=FALSE}
## Fit limma model
limmafit<-function(exprs,groupid,compid){
	compnum<-nrow(compid)
	genenum<-nrow(exprs)	
	limmat<-matrix(0,genenum,compnum)
	limmas2<-rep(0,compnum)
	limmadf<-rep(0,compnum)
	limmav0<-rep(0,compnum)
	limmag1num<-rep(0,compnum)
	limmag2num<-rep(0,compnum)

	for(i in 1:compnum) {
		selid1<-which(groupid == compid[i,1]) 
		selid2<-which(groupid == compid[i,2])
		eset<-new("ExpressionSet", exprs=cbind(exprs[,selid1],exprs[,selid2]))
		g1num<-length(selid1)
		g2num<-length(selid2)
		designmat<-cbind(base=rep(1,(g1num+g2num)), delta=c(rep(0,g1num),rep(1,g2num)))
		fit<-lmFit(eset,designmat)
		fit<-eBayes(fit)
		limmat[,i]<-fit$t[,2]
		limmas2[i]<-fit$s2.prior
		limmadf[i]<-fit$df.prior
		limmav0[i]<-fit$var.prior[2]
		limmag1num[i]<-g1num
		limmag2num[i]<-g2num

	}
      limmacompnum<-nrow(compid)
	result<-list(t=limmat, v0=limmav0, df0=limmadf, s20=limmas2, g1num=limmag1num, g2num=limmag2num,compnum=limmacompnum)
}

## Log-likelihood for moderated t under H0
modt.f0.loglike<-function(x,df){
	a<-dt(x, df, log=TRUE)
	result<-as.vector(a)
      	flag<-which(is.na(result)==TRUE)
      	result[flag]<-0
      	result
}

## Log-likelihood for moderated t under H1
## param=c(df,g1num,g2num,v0)
modt.f1.loglike<-function(x,param) {
	df<-param[1]
	g1num<-param[2]
	g2num<-param[3]
	v0<-param[4]
	w<-sqrt(1+v0/(1/g1num+1/g2num))
	dt(x/w, df, log=TRUE)-log(w)
	a<-dt(x/w, df, log=TRUE)-log(w)
      	result<-as.vector(a)
      	flag<-which(is.na(result)==TRUE)
      	result[flag]<-0
      	result
}

## Correlation Motif Fit
cmfit_1<-function(x, type, K=1, tol=1e-3, max.iter=100) {
	## initialize
	xrow<-nrow(x)
	xcol<-ncol(x)
	loglike0<-list()
	loglike1<-list()
	p<-rep(1,K)/K
	q<-matrix(runif(K*xcol), K, xcol)
	q[1,]<-rep(0.01,xcol)

	## compute loglikelihood
	for(i in 1:xcol) {
		f0<-type[[i]][[1]]
		f0param<-type[[i]][[2]]
		f1<-type[[i]][[3]]
		f1param<-type[[i]][[4]]
		loglike0[[i]]<-f0(x[,i],f0param)
		loglike1[[i]]<-f1(x[,i],f1param)	
	}

	## EM algorithm to get MLE of p and q
	condlike<-list()
	for(i in 1:xcol) {
		condlike[[i]]<-matrix(0,xrow,K)
	}

	loglike.old <- -1e10
	for(i.iter in 1:max.iter) {

		err<-tol+1

		## compute posterior cluster membership
		clustlike<-matrix(0,xrow,K)
		templike <- matrix(0,xrow,2)
		for(j in 1:K) {
			for(i in 1:xcol) {
				templike[,1]<-log(q[j,i])+loglike1[[i]]
				templike[,2]<-log(1-q[j,i])+loglike0[[i]]
				tempmax<-pmax(templike[,1],templike[,2])
				for(z in 1:2) {
					templike[,z]<-exp(templike[,z]-tempmax)
				}
				tempsum<-templike[,1]+templike[,2]
				clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
				condlike[[i]][,j]<-templike[,1]/tempsum
			}
			clustlike[,j]<-clustlike[,j]+log(p[j])
		}

		tempmax<-apply(clustlike,1,max)
		for(j in 1:K) {
			clustlike[,j]<-exp(clustlike[,j]-tempmax)
		}
		tempsum<-apply(clustlike,1,sum)

		## update motif occurrence rate
		for(j in 1:K) {
			clustlike[,j]<-clustlike[,j]/tempsum
		}

		p.new<-(apply(clustlike,2,sum))/(xrow)

		## update motifs
		q.new<-matrix(0, K, xcol)
		for(j in 1:K) {
			clustpsum<-sum(clustlike[,j])
			for(i in 1:xcol) {
				q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j]))/(clustpsum)
			}
		}

		## evaluate convergence
		err.p<-max(abs(p.new-p)/p)
		err.q<-max(abs(q.new-q)/q)
		err<-max(err.p, err.q)

		## evaluate whether the log.likelihood increases

		loglike.new<-(sum(tempmax+log(tempsum)))/xrow

		p<-p.new
		q<-q.new
		loglike.old<-loglike.new

		if(err<tol) {
			break;
		}
	}
	## compute posterior p
	clustlike<-matrix(0,xrow,K)
	for(j in 1:K) {
		for(i in 1:xcol) {
			templike[,1]<-log(q[j,i])+loglike1[[i]]
			templike[,2]<-log(1-q[j,i])+loglike0[[i]]
			tempmax<-pmax(templike[,1],templike[,2])
			for(z in 1:2) {
				templike[,z]<-exp(templike[,z]-tempmax)
			}
			tempsum<-templike[,1]+templike[,2]
			clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
			condlike[[i]][,j]<-templike[,1]/tempsum
		}
		clustlike[,j]<-clustlike[,j]+log(p[j])
	}

	tempmax<-apply(clustlike,1,max)
	for(j in 1:K) {
		clustlike[,j]<-exp(clustlike[,j]-tempmax)
	}
	tempsum<-apply(clustlike,1,sum)
	for(j in 1:K) {
		clustlike[,j]<-clustlike[,j]/tempsum
	}

	p.post<-matrix(0,xrow,xcol)
	for(j in 1:K) {
		for(i in 1:xcol) {
			p.post[,i]<-p.post[,i]+clustlike[,j]*condlike[[i]][,j]
		}
	}

	## return
	#calculate back loglikelihood
	loglike.old<-loglike.old#-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
	loglike.old<-loglike.old*xrow
	result<-list(p.post=p.post, motif.prior=p, motif.q=q, loglike=loglike.old)
}

generatetype<-function(limfitted)
{
	jtype<-list()
	df<-limfitted$g1num+limfitted$g2num-2+limfitted$df0
	for(j in 1:limfitted$compnum)
	{
  		jtype[[j]]<-list(f0=modt.f0.loglike, f0.param=df[j], f1=modt.f1.loglike, f1.param=c(df[j],limfitted$g1num[j],limfitted$g2num[j],limfitted$v0[j]))  
  	}
    	jtype
}

cormotiffit_1<-function(exprs,groupid,compid,K=1, tol=1e-3, max.iter=100,BIC=TRUE)
{
	limfitted<-limmafit(exprs,groupid,compid)
        jtype<-generatetype(limfitted)
	fitresult<-list()
	for(i in 1:length(K))
		fitresult[[i]]<-cmfit_1(limfitted$t,type=jtype,K=K[i],max.iter=max.iter,tol=tol)
	bic<-rep(0,length(K))
	aic<-rep(0,length(K))
	loglike<-rep(0,length(K))
	for(i in 1:length(K))
			loglike[i]<-fitresult[[i]]$loglike
	for(i in 1:length(K))
			bic[i]<--2*fitresult[[i]]$loglike+(K[i]-1+K[i]*limfitted$compnum)*log(dim(exprs)[1])
	for(i in 1:length(K))
			aic[i]<--2*fitresult[[i]]$loglike+2*(K[i]-1+K[i]*limfitted$compnum)
	if(BIC==TRUE)
	{
		bestflag=which(bic==min(bic))
	}
	else
	{
		bestflag=which(aic==min(aic))
	}
	result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
			aic=cbind(K,aic),loglike=cbind(K,loglike),allresult=fitresult)
	#result<-list(bestmotif=fitresult[[4]],bic=cbind(K,bic),
	#             aic=cbind(K,aic),loglike=cbind(K,loglike))
}

plotMotif_1<-function(bestmotif,title="")
{
	  layout(matrix(1:2,ncol=2))
          u<-1:dim(bestmotif$motif.q)[2]
          v<-1:dim(bestmotif$motif.q)[1]
          image(u,v,t(bestmotif$motif.q),
          col=gray(seq(from=1,to=0,by=-0.1)),xlab="Study",yaxt = "n",
		ylab="Corr. Motifs",main=paste(title,"pattern",sep=" "))
	  axis(2,at=1:length(v))
          for(i in 1:(length(u)+1))
          {
                abline(v=(i-0.5))
          }
          for(i in 1:(length(v)+1)) 
          {
                abline(h=(i-0.5))
          }
	  Ng=10000
	  if(is.null(bestmotif$p.post)!=TRUE)
		Ng=nrow(bestmotif$p.post)
	  genecount=floor(bestmotif$motif.p*Ng)
	  NK=nrow(bestmotif$motif.q)
	  plot(0,0.7,pch=".",xlim=c(0,1.2),ylim=c(0.75,NK+0.25),
		frame.plot=FALSE,axes=FALSE,xlab="No. of genes",ylab="", main=paste(title,"frequency",sep=" "))
	  segments(0,0.7,bestmotif$motif.p[1],0.7)
	  rect(0,1:NK-0.3,bestmotif$motif.p,1:NK+0.3,
		col="dark grey")
	  mtext(1:NK,at=1:NK,side=2,cex=0.8)
	  text(bestmotif$motif.p+0.15,1:NK,
	  labels=floor(bestmotif$motif.p*Ng))
}
```

```{r,cache=TRUE}
motif.fitted<-cormotiffit_1(exprs.simu2,simu2_groupid,simu2_compgroup,K=1:5,max.iter=1000,BIC=TRUE)

motif.fitted$bic
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

The two methods do give very similar results. However in practice, it takes much less iterations for methods with prior to achieve its accuracy. Here we prefer to apply CorMotif with prior. In the rest of document, without specification, we apply CorMOtif with prior.

## Simulation based on limma

We first duplicated several simulation from $Wei$ $and$ $Ji$ (2015), which was still based on limma.

### Simulation 1

In simulation 1, we generated 10,000 genes and four studies according to the four different patterns: 100 genes were differentially expressed in all four studies ($Y_i = [1,1,1,1]$); 400 genes were differential only in studies 1 and 2 ($Y_i = [1,1,0,0]$); 400 genes were differential only in studies 2 and 3 ($Y_i = [0,1,1,0]$); 9100 genes were non-differential ($Y_i = [0,0,0,0]$). Each study had six samples: three cases and three controls. The variances $\sigma_{ir}^2$ were simulated from $n_{0r}s_{0r}^2/\chi^2(n_{0r})$, where $n_{0r}=4$ and $s_{0r}^2=0.02$. The expression values were generated using $x_{irlj}\sim N(0,\sigma_{ir}^2)$. When $y_{ir}=1$, drew $\mu_{ir}$ from $N(0,16\sigma_{ir}^2)$, and $\mu_{ir}$ was added to the three cases.

Notice that in $Wei$ $and$ $Ji$ (2015), they drew $\mu_{ir}$ from $N(0,4\sigma_{ir}^2)$. I think it is a typo, since with this setting, the simulation results are quite unsatisfying and only two patterns were recognized. The reason for that is the difference between cases and controls will be minor if the variance is not large enough, which will weaken the pattern.

```{r,cache=TRUE}
K = 4
D = 4
G = 10000
pi0 = c(0.01,0.04,0.04,0.91)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,1,1,0)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=6*D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(1)
for(g in 1:G){
  for(d in 1:D){
    sigma2 = 4*0.02/rchisq(1,4)
    x = rnorm(6,0,sqrt(sigma2))
    if(A[g,d]==1){
      dx = rnorm(1,0,4*sqrt(sigma2))
      x[1:3] = x[1:3]+dx
    }
    X[g,(6*d-5):(6*d)] = x
  }
}
groupid = rep(1:8,each=3)
compgroup = t(matrix(1:8,ncol=4))
motif.fitted<-cormotiffit(X,groupid,compgroup,K=1:5,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

The result matches well with $Wei$ $and$ $Ji$ (2015).

### Simulation 2

In simulation 2, we generated 10,000 genes and four studies according to the four different patterns: 300 genes were differentially expressed in all four studies ($Y_i = [1,1,1,1]$); 300 genes were differential only in studies 1 and 2 ($Y_i = [1,1,0,0]$); 300 genes were differential only in studies 3 and 4 ($Y_i = [0,0,1,1]$); 9100 genes were non-differential ($Y_i = [0,0,0,0]$).

```{r,cache=TRUE}
K = 4
D = 4
G = 10000
pi0 = c(0.03,0.03,0.03,0.91)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,0,1,1)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=6*D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(2)
for(g in 1:G){
  for(d in 1:D){
    sigma2 = 4*0.02/rchisq(1,4)
    x = rnorm(6,0,sqrt(sigma2))
    if(A[g,d]==1){
      dx = rnorm(1,0,4*sqrt(sigma2))
      x[1:3] = x[1:3]+dx
    }
    X[g,(6*d-5):(6*d)] = x
  }
}
groupid = rep(1:8,each=3)
compgroup = t(matrix(1:8,ncol=4))
motif.fitted<-cormotiffit(X,groupid,compgroup,K=1:5,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

Then we could check if we change the proportion of four patterns to $200:50:50:9700$.

```{r,cache=TRUE}
K = 4
D = 4
G = 10000
pi0 = c(0.02,0.005,0.005,0.97)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,0,1,1)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=6*D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(1)
for(g in 1:G){
  for(d in 1:D){
    sigma2 = 4*0.02/rchisq(1,4)
    x = rnorm(6,0,sqrt(sigma2))
    if(A[g,d]==1){
      dx = rnorm(1,0,4*sqrt(sigma2))
      x[1:3] = x[1:3]+dx
    }
    X[g,(6*d-5):(6*d)] = x
  }
}
groupid = rep(1:8,each=3)
compgroup = t(matrix(1:8,ncol=4))
motif.fitted<-cormotiffit(X,groupid,compgroup,K=1:5,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

We could find that the second and third pattern merged togethor, since their proportions were relatively small compared to the first and forth pattern.

### Simulation 3

In simulation 3, we generated 10,000 genes and eight studies according to the four different patterns: 200 genes were differentially expressed in all four studies ($Y_i = [1,1,1,1,1,1,1,1]$); 200 genes were differential only in studies 1, 2, 3, 4 ($Y_i = [1,1,1,1,0,0,0,0]$); 200 genes were differential only in studies 5, 6, 7, 8 ($Y_i = [0,0,0,0,1,1,1,1]$); 200 genes were differential only in studies 3, 4, 5, 6 ($Y_i = [0,0,1,1,1,1,0,0]$); 9200 genes were non-differential ($Y_i = [0,0,0,0,0,0,0,0,0]$).

```{r,cache=TRUE}
K = 5
D = 8
G = 10000
pi0 = c(0.02,0.02,0.02,0.02,0.92)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1,1,1,1,1)
q0[2,] = c(1,1,1,1,0,0,0,0)
q0[3,] = c(0,0,0,0,1,1,1,1)
q0[4,] = c(0,0,1,1,1,1,0,0)
q0[5,] = c(0,0,0,0,0,0,0,0)
X = matrix(0,nrow=G,ncol=6*D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(3)
for(g in 1:G){
  for(d in 1:D){
    sigma2 = 4*0.02/rchisq(1,4)
    x = rnorm(6,0,sqrt(sigma2))
    if(A[g,d]==1){
      dx = rnorm(1,0,4*sqrt(sigma2))
      x[1:3] = x[1:3]+dx
    }
    X[g,(6*d-5):(6*d)] = x
  }
}
groupid = rep(1:16,each=3)
compgroup = t(matrix(1:16,nrow=2))
motif.fitted<-cormotiffit(X,groupid,compgroup,K=1:6,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

CorMotif tends to merge weaker patterns with small proportion. Just as $Wei$ $and$ $Ji$ (2015) states, correlation motifs only represent a parsimonious representation of the correlation structure supported by the available data. On the other hand, the two stage method may also be blame. Adding $limma$ structure can increase inaccuracy for estimation. In the next section, we will generate $x_{ir}$ directly from a Gaussian distribution. In fact, any distribution can be applied under our framework. In addition, we can also estimate the parameters for Gaussian distribution.

## CorMotif based on Gaussian model

Let

\begin{eqnarray*}
f_{r0}(x_{ir}) &=& N(x_{ir};0,1), \\
f_{r1}(x_{ir}) &=& N(x_{ir};0,1+\sigma_r^2).
\end{eqnarray*}

First, we checked the performance of CorMotif by assuming we know $\sigma^2$ exactly. Then we dealed with the situation where we were given no information about $\sigma^2$. Apply

$$\sigma_r^{2(t+1)}= \frac{\sum_{i=1}^n \sum_{k=1}^K (x_{ir}^2-1)p_{ikr1}}{\sum_{i=1}^n \sum_{k=1}^K p_{ikr1}},$$

from my write-up document, we could iteratively update $\sigma_r^2$.

### Simulation 4

We used the same patterns and proportions as Simulation 1 in the previous section and let $\sigma^2=(4^2,4^2,4^2,4^2)$. Assume we know $\sigma^2$, the estimated results were:

```{r,echo=FALSE}
## Log-likelihood for norm under H0 and H1
modt.f.loglike<-function(x,m,s2) {
	a<-dnorm(x, m, sqrt(s2), log=TRUE)
	result<-as.vector(a)
  flag<-which(is.na(result)==TRUE)
  result[flag]<-0
  result
}

```

```{r,echo=FALSE}
cmfit.norm.known<-function(x, sigma2, K=1, tol=1e-3, max.iter=100) {
	## initialize
	xrow<-nrow(x)
	xcol<-ncol(x)
	loglike0<-list()
	loglike1<-list()
	p<-rep(1,K)/K
	q<-matrix(runif(K*xcol), K, xcol)
	q[1,]<-rep(0.01,xcol)

	## compute loglikelihood
	for(i in 1:xcol) {
  		loglike0[[i]]<-modt.f.loglike(x[,i],0,1)
  		loglike1[[i]]<-modt.f.loglike(x[,i],0,1+sigma2[i])	
  	}

	## EM algorithm to get MLE of p, q and sigma2
	condlike<-list()
	for(i in 1:xcol) {
		condlike[[i]]<-matrix(0,xrow,K)
	}

	loglike.old <- -1e10
	for(i.iter in 1:max.iter) {
		err<-tol+1

		## compute posterior cluster membership
		clustlike<-matrix(0,xrow,K)
		templike <- matrix(0,xrow,2)
		for(j in 1:K) {
			for(i in 1:xcol) {
				templike[,1]<-log(q[j,i])+loglike1[[i]]
				templike[,2]<-log(1-q[j,i])+loglike0[[i]]
				tempmax<-pmax(templike[,1],templike[,2])
				for(z in 1:2) {
					templike[,z]<-exp(templike[,z]-tempmax)
				}
				tempsum<-templike[,1]+templike[,2]
				clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
				condlike[[i]][,j]<-templike[,1]/tempsum
			}
			clustlike[,j]<-clustlike[,j]+log(p[j])
		}

		tempmax<-apply(clustlike,1,max)
		for(j in 1:K) {
			clustlike[,j]<-exp(clustlike[,j]-tempmax)
		}
		tempsum<-apply(clustlike,1,sum)

		## update motif occurrence rate
		for(j in 1:K) {
			clustlike[,j]<-clustlike[,j]/tempsum
		}

		#p.new<-(apply(clustlike,2,sum))/(xrow)
		p.new<-(apply(clustlike,2,sum)+1)/(xrow+K)
		## update motifs
		q.new<-matrix(0, K, xcol)
		for(j in 1:K) {
			clustpsum<-sum(clustlike[,j])
			for(i in 1:xcol) {
				#q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j]))/(clustpsum)
				q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j])+1)/(clustpsum+2)
			}
		}
		## evaluate convergence
		err.p<-max(abs(p.new-p)/p)
		err.q<-max(abs(q.new-q)/q)
		err<-max(err.p, err.q)

		## evaluate whether the log.likelihood increases
		loglike.new<-(sum(tempmax+log(tempsum))+sum(log(p.new))+sum(log(q.new)+log(1-q.new)))/xrow
		#loglike.new<-(sum(tempmax+log(tempsum)))/xrow

		p<-p.new
		q<-q.new
		loglike.old<-loglike.new

		if(err<tol) {
			break;
		}
	}
	## compute posterior p
	clustlike<-matrix(0,xrow,K)
	for(j in 1:K) {
		for(i in 1:xcol) {
			templike[,1]<-log(q[j,i])+loglike1[[i]]
			templike[,2]<-log(1-q[j,i])+loglike0[[i]]
			tempmax<-pmax(templike[,1],templike[,2])
			for(z in 1:2) {
				templike[,z]<-exp(templike[,z]-tempmax)
			}
			tempsum<-templike[,1]+templike[,2]
			clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
			condlike[[i]][,j]<-templike[,1]/tempsum
		}
		clustlike[,j]<-clustlike[,j]+log(p[j])
	}

	tempmax<-apply(clustlike,1,max)
	for(j in 1:K) {
		clustlike[,j]<-exp(clustlike[,j]-tempmax)
	}
	tempsum<-apply(clustlike,1,sum)
	for(j in 1:K) {
		clustlike[,j]<-clustlike[,j]/tempsum
	}

	p.post<-matrix(0,xrow,xcol)
	for(j in 1:K) {
		for(i in 1:xcol) {
			p.post[,i]<-p.post[,i]+clustlike[,j]*condlike[[i]][,j]
		}
	}

	## return
	#calculate back loglikelihood
	loglike.old<-loglike.old-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
	#loglike.old<-loglike.old#-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
	loglike.old<-loglike.old*xrow
  result<-list(p.post=p.post, motif.prior=p, motif.q=q,loglike=loglike.old,motif.s = sigma2)
}

cormotiffit.norm.known<-function(x,sigma2,K=1, tol=1e-3, max.iter=100,BIC=TRUE)
{
	fitresult<-list()
	for(i in 1:length(K))
		fitresult[[i]]<-cmfit.norm.known(x,sigma2,K=K[i],max.iter=max.iter,tol=tol)
	bic<-rep(0,length(K))
	aic<-rep(0,length(K))
	loglike<-rep(0,length(K))
	for(i in 1:length(K))
			loglike[i]<-fitresult[[i]]$loglike
	for(i in 1:length(K))
			bic[i]<--2*fitresult[[i]]$loglike+(K[i]-1+K[i]*ncol(x))*log(dim(x)[1])
	for(i in 1:length(K))
			aic[i]<--2*fitresult[[i]]$loglike+2*(K[i]-1+K[i]*ncol(x))
	if(BIC==TRUE)
	{
		bestflag=which(bic==min(bic))
	}
	else
	{
		bestflag=which(aic==min(aic))
	}
	#result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
	#		aic=cbind(K,aic),loglike=cbind(K,loglike))
	result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
	             aic=cbind(K,aic),loglike=cbind(K,loglike),allresult=fitresult)
} 
```

```{r,cache=TRUE}
K = 4
D = 4
G = 10000
sigma2 = c(16,16,16,16)
pi0 = c(0.01,0.04,0.04,0.91)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,1,1,0)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(4)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
motif.fitted<-cormotiffit.norm.known(X,sigma2,K=1:5,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

CorMotif could correctly recognize four patterns and the proportion is also reasonable. Then we checked the performace without any prior information about $\sigma^2$ with the same simulated data.

```{r,echo=FALSE}
cmfit.norm<-function(x, K=1, tol=1e-3, max.iter=100) {
	## initialize
	xrow<-nrow(x)
	xcol<-ncol(x)
	loglike0<-list()
	loglike1<-list()
	p<-rep(1,K)/K
	q<-matrix(runif(K*xcol), K, xcol)
	q[1,]<-rep(0.01,xcol)
	sigma2 = rep(1,xcol)
#	sigma2 = c(16,16,16,16)

	## EM algorithm to get MLE of p, q and sigma2
	condlike<-list()
	for(i in 1:xcol) {
		condlike[[i]]<-matrix(0,xrow,K)
	}

	loglike.old <- -1e10

	for(i.iter in 1:max.iter) {

	  	err<-tol+1
		## compute loglikelihood
  	for(i in 1:xcol) {
    		loglike0[[i]]<-modt.f.loglike(x[,i],0,1)
    		loglike1[[i]]<-modt.f.loglike(x[,i],0,1+sigma2[i])	
    	}
		## compute posterior cluster membership
		clustlike<-matrix(0,xrow,K)
		templike <- matrix(0,xrow,2)
		for(j in 1:K) {
			for(i in 1:xcol) {
				templike[,1]<-log(q[j,i])+loglike1[[i]]
				templike[,2]<-log(1-q[j,i])+loglike0[[i]]
				tempmax<-pmax(templike[,1],templike[,2])
				for(z in 1:2) {
					templike[,z]<-exp(templike[,z]-tempmax)
				}
				tempsum<-templike[,1]+templike[,2]
				clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
				condlike[[i]][,j]<-templike[,1]/tempsum
			}
			clustlike[,j]<-clustlike[,j]+log(p[j])
		}

		tempmax<-apply(clustlike,1,max)
		for(j in 1:K) {
			clustlike[,j]<-exp(clustlike[,j]-tempmax)
		}
		tempsum<-apply(clustlike,1,sum)

		## update motif occurrence rate
		for(j in 1:K) {
			clustlike[,j]<-clustlike[,j]/tempsum
		}

		#p.new<-(apply(clustlike,2,sum))/(xrow)
		p.new<-(apply(clustlike,2,sum)+1)/(xrow+K)
		## update motifs
		q.new<-matrix(0, K, xcol)
		for(j in 1:K) {
			clustpsum<-sum(clustlike[,j])
			for(i in 1:xcol) {
				#q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j]))/(clustpsum)
				q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j])+1)/(clustpsum+2)
			}
		}
    ## update sigma2
		sigma2.new<-rep(0,xcol)
		for(i in 1:xcol){
		  pikr1 = clustlike*condlike[[i]]
		  xx = x[,i]^2-1
		  sigma2.new[i] = sum(t(xx)%*%pikr1)/sum(pikr1)
		}
		## evaluate convergence
		err.p<-max(abs(p.new-p)/p)
		err.q<-max(abs(q.new-q)/q)
		err.sigma2<-max(abs((sigma2.new-sigma2)/sigma2))
		err<-max(err.p, err.q, err.sigma2)
    #err<-max(err.p, err.q)
		## evaluate whether the log.likelihood increases
		loglike.new<-(sum(tempmax+log(tempsum))+sum(log(p.new))+sum(log(q.new)+log(1-q.new)))/xrow
		#loglike.new<-(sum(tempmax+log(tempsum)))/xrow

		p<-p.new
		q<-q.new
		sigma2<-sigma2.new
		loglike.old<-loglike.new

		if(err<tol) {
			break;
		}
	}
	## compute posterior p
	clustlike<-matrix(0,xrow,K)
	for(j in 1:K) {
		for(i in 1:xcol) {
			templike[,1]<-log(q[j,i])+loglike1[[i]]
			templike[,2]<-log(1-q[j,i])+loglike0[[i]]
			tempmax<-pmax(templike[,1],templike[,2])
			for(z in 1:2) {
				templike[,z]<-exp(templike[,z]-tempmax)
			}
			tempsum<-templike[,1]+templike[,2]
			clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
			condlike[[i]][,j]<-templike[,1]/tempsum
		}
		clustlike[,j]<-clustlike[,j]+log(p[j])
	}

	tempmax<-apply(clustlike,1,max)
	for(j in 1:K) {
		clustlike[,j]<-exp(clustlike[,j]-tempmax)
	}
	tempsum<-apply(clustlike,1,sum)
	for(j in 1:K) {
		clustlike[,j]<-clustlike[,j]/tempsum
	}

	p.post<-matrix(0,xrow,xcol)
	for(j in 1:K) {
		for(i in 1:xcol) {
			p.post[,i]<-p.post[,i]+clustlike[,j]*condlike[[i]][,j]
		}
	}

	## return
	#calculate back loglikelihood
	loglike.old<-loglike.old-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
	#loglike.old<-loglike.old#-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
	loglike.old<-loglike.old*xrow
  result<-list(p.post=p.post, motif.prior=p, motif.q=q,loglike=loglike.old,motif.s = sigma2)
}

cormotiffit.norm<-function(x,K=1, tol=1e-3, max.iter=100,BIC=TRUE)
{
	fitresult<-list()
	for(i in 1:length(K))
		fitresult[[i]]<-cmfit.norm(x,K=K[i],max.iter=max.iter,tol=tol)
	bic<-rep(0,length(K))
	aic<-rep(0,length(K))
	loglike<-rep(0,length(K))
	for(i in 1:length(K))
			loglike[i]<-fitresult[[i]]$loglike
	for(i in 1:length(K))
			bic[i]<--2*fitresult[[i]]$loglike+(K[i]-1+K[i]*ncol(x))*log(dim(x)[1])
	for(i in 1:length(K))
			aic[i]<--2*fitresult[[i]]$loglike+2*(K[i]-1+K[i]*ncol(x))
	if(BIC==TRUE)
	{
		bestflag=which(bic==min(bic))
	}
	else
	{
		bestflag=which(aic==min(aic))
	}
	#result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
	#		aic=cbind(K,aic),loglike=cbind(K,loglike))
	result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
	             aic=cbind(K,aic),loglike=cbind(K,loglike),allresult=fitresult)
} 
```

```{r,cache=TRUE}
motif.fitted<-cormotiffit.norm(X,K=1:5,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

And the estimated $\sigma^2$ is

```{r,cache=TRUE}
motif.fitted$bestmotif$motif.s
```

### Simulation 5

We used the same patterns and proportions as Simulation 2 in the previous section and let $\sigma^2=(4^2,4^2,4^2,4^2)$. Assume we know $\sigma^2$, the estimated results were:

```{r,cache=TRUE}
K = 4
D = 4
G = 10000
sigma2 = c(16,16,16,16)
pi0 = c(0.03,0.03,0.03,0.91)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,0,1,1)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(4)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
motif.fitted<-cormotiffit.norm.known(X,sigma2,K=1:5,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

Again, CorMotif could correctly recognize four patterns and the proportion is also accurate. Then we checked the performace without any prior information about $\sigma^2$ with the same simulated data.

```{r,cache=TRUE}
motif.fitted<-cormotiffit.norm(X,K=1:5,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

And the estimated $\sigma^2$ is

```{r,cache=TRUE}
motif.fitted$bestmotif$motif.s
```

### Simulation 6

We used the same patterns and proportions as Simulation 3 in the previous section and let $\sigma^2=(4^2,4^2,4^2,4^2,4^2,4^2,4^2,4^2)$. Assume we know $\sigma^2$, the estimated results were:

```{r,cache=TRUE}
K = 5
D = 8
G = 10000
sigma2 = rep(16,D)
pi0 = c(0.02,0.02,0.02,0.02,0.92)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1,1,1,1,1)
q0[2,] = c(1,1,1,1,0,0,0,0)
q0[3,] = c(0,0,0,0,1,1,1,1)
q0[4,] = c(0,0,1,1,1,1,0,0)
q0[5,] = c(0,0,0,0,0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
motif.fitted<-cormotiffit.norm.known(X,sigma2,K=1:6,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

Then we checked the performace without any prior information about $\sigma^2$ with the same simulated data.

```{r,cache=TRUE}
motif.fitted<-cormotiffit.norm(X,K=1:6,max.iter=1000,BIC=TRUE)
plotIC(motif.fitted)
plotMotif(motif.fitted)
```

And the estimated $\sigma^2$ is

```{r,cache=TRUE}
motif.fitted$bestmotif$motif.s
```

Under all situation, CorMotif without knowing $\sigma^2$ performed equally well with knowing $\sigma^2$.

## Discussion

We tested CorMotif for both $limma$ and Guassion model. Through our simulation, we find that correlation motifs only represent a parsimonious representation of the correlation structure. The reason is twofold. On the one hand, CorMotif tends to merge patterns with small proportion. On the other hand, if the differences between cases and controls (for Gaussian model, the variance $\sigma^2$) are not significant, CorMotif cannot detect the pattern either.

### 2017-09-05 analysis/Simulation.Rmd
wflow_publish("Simulation.Rmd")

@@
 date: "8/30/2017"
-output: pdf_document
+output:
+  html_document:
+    toc: false
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
 ```

### 2017-09-05 analysis/Simulation.Rmd
wflow_publish("Simulation.Rmd")

@@
 ---
+<!-- The file analysis/chunks.R contains chunks that define default settings
+shared across the workflowr files. -->
+```{r read-chunk, include=FALSE, cache=FALSE}
+knitr::read_chunk("chunks.R")
+```
+
+<!-- Update knitr chunk options -->
+```{r knitr-opts-chunk, include=FALSE}
+```
+
+<!-- Insert the date the file was last updated -->
+```{r last-updated, echo=FALSE, results='asis'}
+```
+
+<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
+ package git2r is installed -->
+```{r code-version, echo=FALSE, results='asis'}
+```
+
+<!-- Add your analysis here -->
 

### 2017-09-10 analysis/Simulation.Rmd
wflow_publish("Simulation.Rmd")

@@
 date: "8/30/2017"
-output:
-  html_document:
-    toc: false
+output: html_document
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
+knitr::opts_chunk$set(echo = TRUE)
 ```

### 2017-09-16 paper/main.tex
add paper file

\documentclass[11pt]{article}   	
\usepackage{geometry}                		
\geometry{letterpaper}                   		
\usepackage{setspace}
\doublespacing

\usepackage{graphicx}					
\usepackage{amssymb}
\usepackage{amsmath,amsfonts}
\usepackage{bm}

\newtheorem{assumption}{Assumption}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

\title{Generic Cormotif}
\author{Zhiwei Ma}
\date{September 2017}

\begin{document}

\maketitle

\section{Introduction}
%\textit{lfdr}
\section{Methods}

\subsection{Model Outline}

Our method is designed to estimate the effects of multiple units in multiple studies. Suppose there are totally $n$ units and $R$ studies. Let $\beta = [\beta_{jr}]_{J\times R}$ denote "effects" of interest. For instance, $\beta_{jr}$ could be the difference in the mean expression of gene $j$ at study $r$ under two conditions.

Assume that the available data are estimates $\hat \beta = [\hat \beta_{jr}]_{J\times R}$ of the effects, and the corresponding estimated standard errors $\hat s = [\hat s_{jr}]_{J\times R}$. Let $\beta_j:=(\beta_{j1},\ldots,\beta_{jr})'$, where $j=1,2,\ldots,n$. Similarly for $\hat \beta_j$.

Similar to the original $Cormotif$, we assume that all units fall into $K$ different classes. Besides, we have the following assumptions:

\begin{assumption}Each units $j$ is randomly and independently assigned to a class label $z_j$ according to probability $\pi=(\pi_1,\ldots,\pi_K)$. Here $\pi_k=P(z_j=k)$ is the prior probability that a unit belongs to class $k$. We have $\sum_k\pi_k = 1$.
\end{assumption}

\begin{assumption}Given unit's class label $z_j=k$, the effects $\beta_{j1},\beta_{j2},\ldots,\beta_{jR}$ are independent from unimodal distribution $g_{kr}$. 
\end{assumption}

Here we describe the simplest version of our model. First we assume the effect $\beta_{jr}$ is independent of their standard errors $\hat s_{jr}$. Then by Assumption 1 and 2, we have
\begin{equation}
p(\beta_j|\pi,g,\hat s) = \sum_{k=1}^K\pi_k\prod_{r=1}^R  g_{kr}(\beta_{jr}),
\label{eq:1}
\end{equation}
where $g$ represents $\{g_{kr}|k = 1,\ldots,K$, $r=1,\ldots,R\}$. A simple way to implement the unimodal assumption (UA) is to assume that $g_{kr}$ is a mixture of a point mass at 0 and a mixture of $zero-mean$ normal distribution:
\begin{equation}
g_{kr}(\cdot) =  w_0^{kr}\delta_0(\cdot)+\sum_{l=1}^L w_l^{kr}N(\cdot;0,\sigma_l^2),
\label{eq:2}
\end{equation}
where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot;\mu,\sigma^2)$ denotes the density of Normal distribution with mean $\mu$ and variance $\sigma^2$. Here we assume  $\sigma_1,\sigma_2,\ldots,\sigma_L$ are fixed positive numbers forming a wide and dense grid.

For likelihood $p(\hat \beta|\beta,\hat s)$, we assume a Normal approximation:
\begin{equation}
p(\hat \beta|\beta,\hat s) = \prod_{j=1}^np(\hat \beta_j|\beta_j,\hat s) = \prod_{j=1}^n\prod_{r=1}^RN(\hat\beta_{jr};\beta_{jr},\hat s_{jr}^2).
\label{eq:3}
\end{equation}

Together, (\ref{eq:1})-(\ref{eq:3}) imply that 
\begin{eqnarray}
p(\hat\beta|\pi,g,\hat s) &=& \prod_{j=1}^n\left[\sum_{k=1}^K\pi_k\prod_{r=1}^R(g_{kr}*N_{jr})(\hat\beta_{jr})\right] \nonumber \\ 
&=& \prod_{j=1}^n\left\{\sum_{k=1}^K\pi_k\prod_{r=1}^R\left[ \sum_{l=0}^Lw_l^{kr}N(\hat\beta_{jr};0,\sigma_l^2+\hat s_{jr}^2)\right]\right\}.
\label{eq:4}
\end{eqnarray}
Here $N_{jr}$ denotes $N(\cdot;0,\hat s_{jr}^2)$ and $*$ means the convolution of two functions. We define $\sigma_0:=0$.

Model (\ref{eq:4}) is the extension of original Cormotif: set $L=1$ and assume $\hat s_{jr}^2$s are identical for all $j$, (\ref{eq:4}) is just Cormotif under Normal distribution. Similar to Cormotif, our model can capture the correlation among multiple studies. To see this, consider the likelihood for unit $j$. Based on our model, $p(\hat\beta_j|\pi,g,\hat s) = \sum_{k=1}^K\pi_k\prod_{r=1}^R(g_{kr}*N_{jr})(\hat\beta_{jr})$. The distribution for unit $j$ under study $r$ is $p(\hat\beta_{jr}|\pi,g,\hat s) = \sum_{k=1}^K\pi_k(g_{kr}*N_{jr})(\hat\beta_{jr})$. It is clear that $p(\hat\beta_j|\pi,g,\hat s) \neq \prod_{r=1}^Rp(\hat\beta_{jr}|\pi,g,\hat s)$, so different studies are dependent.

Another advantage of our generic model is that we could compute the posterior distribution for effects $\beta_{jr}$. By Bayes theorem, we have
\begin{equation}
    p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}) \propto p(\beta_{jr}|\hat s_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})
\end{equation}

\subsection{Fitting the model}

\section{Detailed Method}

\subsection{Embellishments}

\subsection{Implementation Details}

\subsubsection{Optimization}
This section We presents the EM algorithm used to estimate both $\pi$ and $g$. First compute the log likelihood function for $\hat\beta$ and group label $z=(z_1,\ldots,z_n)'$:
\begin{eqnarray}
    &&\log p(\hat\beta,z|\pi,g,\hat s) = \sum_{j=1}^n\log p(\hat\beta_j|z_j,\pi,g,\hat s)+\sum_{j=1}^n\log p(z_j|\pi) \nonumber\\
    &=& \sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\log p(\hat\beta_j|z_j=k,\pi,g,\hat s)
    +\sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\log p(z_j=k|\pi)\nonumber\\
    &=& \sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\sum_{r=1}^R\log \left[(g_{kr}*N_{jr})(\hat\beta_{jr}) \right]+\sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\log\pi_k.
\end{eqnarray}

Here $z$ is a latent variable. The EM algorithm seeks to find the MLE of the marginal likelihood (\ref{eq:4}) by iteratively applying the E-step and the M-step.

In the E-step, one evaluates the $Q$-function $Q(\pi,g|\pi^{(t)},g^{(t)})$, here $(\pi^{(t)},g^{(t)})$ is the current estimation. We have
\begin{eqnarray}
    &&Q(\pi,g|\pi^{(t)},g^{(t)}) = E_{z|\hat\beta,\hat s,\pi^{(t)},g^{(t)}}\left[\log p(\hat\beta,z|\pi,g,\hat s)\right]\nonumber\\
    &=&\sum_{j=1}^n\sum_{k=1}^K\sum_{r=1}^Rp_{jk}\log \left[(g_{kr}*N_{jr})(\hat\beta_{jr}) \right]+\sum_{j=1}^n\sum_{k=1}^Kp_{jk}\log\pi_k,
\label{eq:Q}
\end{eqnarray}
where we denote
\begin{eqnarray}
    p_{jk}&=&E_{z|\hat\beta,\hat s,\pi^{(t)},g^{(t)}}[\mathbb{I}(z_j=k)]=p(z_j=k|\hat\beta_j,\hat s,\pi^{(t)},g^{(t)})\nonumber\\
    &=& \frac{p(\hat\beta_j,z_j=k|\hat s,\pi^{(t)},g^{(t)})}{p(\hat\beta_j|\hat s,\pi^{(t)},g^{(t)})}\nonumber\\
    &=& \frac{\pi_k^{(t)}\prod_{r=1}^R(g_{kr}^{(t)}*N_{jr})(\hat\beta_{jr})}{\sum_{k'=1}^K\pi_{k'}^{(t)}\prod_{r=1}^R(g_{k'r}^{(t)}*N_{jr})(\hat\beta_{jr})}
\end{eqnarray}

In the M-step, one finds $\pi$ and $g$ that maximize the $Q$-function $Q(\pi,g|\pi^{(t)},g^{(t)})$, and denote them as $\pi^{(t+1)}$ and $g^{(t+1)}$, that is 
\begin{equation}
(\pi^{(t+1)}, g^ {(t+1)}) =\argmax_{(\pi,g)} Q(\pi,g|\pi^{(t)},g^{(t)}).
\end{equation}
For $\pi^{(t+1)}$, we could optimize it from (\ref{eq:Q}) directly and get
\begin{equation}
    \pi_k^{(t+1)} = \frac{1}{n} \sum_{j=1}^n p_{jk}.
\end{equation}
Notice in (\ref{eq:Q}), we could separately optimize $g_{kr}$ for fixed $k$ and $r$, that is 
\begin{equation}
    g_{kr}^ {(t+1)}=\argmax_{g_{kr}}\sum_{j=1}^np_{jk}\log \left[(g_{kr}*N_{jr})(\hat\beta_{jr}) \right].
\label{eq:g_kr}
\end{equation}
Optimizing (\ref{eq:g_kr}) is a convex problem, which we solve using an EM algorithm, accelerated using R package {\tt SQUAREM} .

\subsubsection{Model Selection: Bayesian Information Criterion}

To determine the class number of $K$ , we use Bayesian Information Criterion(BIC). The BIC in our setting is written as 
\begin{eqnarray}
{\rm BIC}(K) &=& -2\log p(\hat\beta|\pi,g,\hat s) + (K\times R\times L+K-1)\times \log n \nonumber \\
&=& -2 \sum_{j=1}^n\log\left[\sum_{k=1}^K\pi_k\prod_{r=1}^R(g_{kr}*N_{jr})(\hat\beta_{jr})\right] \nonumber\\
&+& (K\times R\times L+K-1)\times \log n.
\end{eqnarray}
Here $K-1$ is the number of parameters for $\pi$, $K\times R\times L$ is the number of parameters involved in $g$ and $n$ is the unit number. We choose the K with the smallest BIC, that is 
\begin{equation}
\hat K = \argmin_{K\geq1} {\rm BIC}(K).
\end{equation}

\end{document}

### 2017-09-16 analysis/license.Rmd
wflow_publish("license.Rmd")

---
title: "License"
output:
  html_document:
    toc: false
---

MIT License

Copyright (c) 2017 Zhiwei Ma

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.

### 2017-09-16 analysis/Simulation.Rmd
wflow_publish("analysis/Simulation.Rmd")
wflow_publish("Simulation.Rmd")

@@
 
-## Model framework
-
-The framework of Correlation Motif model is based on my write-up document and supplementary matrials by $Wei$ $and$ $Ji$ (2015).
+## Model Framework
+
+The framework of Correlation Motif model is based on my [write-up document](../paper/Cormotif.pdf) and supplementary matrials by [$Wei$ $and$ $Ji$ (2015)][wj].
 In $Wei$ $and$ $Ji$ (2015), they implemented an EM algorithm and introduced priors for both $\pi$ and $Q$.
@@
 On the other hand, if the differences between cases and controls (for Gaussian model, the variance $\sigma^2$) are not significant, CorMotif cannot detect the pattern either.
+
+## Session information.
+
+```{r info}
+print(sessionInfo())
+```
+
+[wj]: https://doi.org/10.1093/biostatistics/kxu038

### 2017-09-16 analysis/Simulation.Rmd
wflow_publish("Simulation.Rmd")

@@
 
-The framework of Correlation Motif model is based on my [write-up document](../paper/Cormotif.pdf) and supplementary matrials by [$Wei$ $and$ $Ji$ (2015)][wj].
+The framework of Correlation Motif model is based on my [write-up document][wud] and supplementary matrials by [$Wei$ $and$ $Ji$ (2015)][wj].
 In $Wei$ $and$ $Ji$ (2015), they implemented an EM algorithm and introduced priors for both $\pi$ and $Q$.
@@
 
+[wud]: https://github.com/stephenslab/cormotif/blob/master/paper/Cormotif.pdf
 [wj]: https://doi.org/10.1093/biostatistics/kxu038

### 2017-09-25 paper/main.tex
change paper

@@
 \newtheorem{assumption}{Assumption}
+\newtheorem{remark}{Remark}
+
+\def\lfdr{\textit{lfdr}}
+\def\lfsr{\textit{lfsr}}
+\def\ash{{\tt ash}\xspace}
 
@@
 Let $\beta_j:=(\beta_{j1},\ldots,\beta_{jr})'$, where $j=1,2,\ldots,n$.
-Similarly for $\hat \beta_j$.
+Similarly for $\hat \beta_j$ and $\hat s_j$,
 
-Similar to the original $Cormotif$, we assume that all units fall into $K$ different classes. Besides, we have the following assumptions:
+Similar to the original \textit{Cormotif} \cite{cormotif}, we assume that all units fall into $K$ different classes.
+Besides, we have the following assumptions, which distinguish our model from existing ones:
 
@@
 
-\begin{assumption}Given unit's class label $z_j=k$, the effects $\beta_{j1},\beta_{j2},\ldots,\beta_{jR}$ are independent from unimodal distribution $g_{kr}$. 
+\begin{assumption}Given unit's class label $z_j=k$, the effects $\beta_{j1},\beta_{j2},\ldots,\beta_{jR}$ are independent from unimodal distribution $g_{kr}$, that is $p(\beta_{j}|z_j =k,\hat s)=\prod_{r=1}^R p(\beta_{jr}|z_j =k,\hat s)=\prod_{r=1}^Rg_{kr}( \beta_{jr};\hat s)$
 \end{assumption}
 
-Here we describe the simplest version of our model. First we assume the effect $\beta_{jr}$ is independent of their standard errors $\hat s_{jr}$. Then by Assumption 1 and 2, we have
+\begin{assumption} The effects $\beta_1,\beta_2,\ldots,\beta_n$ are independent given $\pi$ and $g$.
+\end{assumption}
+
+Here we describe the simplest version of our model. First we assume the effect $\beta_{jr}$ is independent of their standard errors $\hat s_{jr}$. Then by Assumption 1, 2 and 3, we have
 \begin{equation}
-p(\beta_j|\pi,g,\hat s) = \sum_{k=1}^K\pi_k\prod_{r=1}^R  g_{kr}(\beta_{jr}),
+p(\beta|\pi,g,\hat s) = \prod_{j=1}^np(\beta_j|\pi,g,\hat s) = \prod_{j=1}^n \left [\sum_{k=1}^K\pi_k\prod_{r=1}^R  g_{kr}(\beta_{jr})\right ]
+%p(\beta_j|\pi,g,\hat s) = \sum_{k=1}^K\pi_k\prod_{r=1}^R  g_{kr}(\beta_{jr}),
 \label{eq:1}
 \end{equation}
+
 where $g$ represents $\{g_{kr}|k = 1,\ldots,K$, $r=1,\ldots,R\}$.
-A simple way to implement the unimodal assumption (UA) is to assume that $g_{kr}$ is a mixture of a point mass at 0 and a mixture of $zero-mean$ normal distribution:
+A simple way to implement the unimodal assumption (UA) is to assume that $g_{kr}$ is a mixture of a point mass at 0 and a mixture of $zero$-$mean$ normal distribution \cite{ash}:
 \begin{equation}
@@
 where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot;\mu,\sigma^2)$ denotes the density of Normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here we assume  $\sigma_1,\sigma_2,\ldots,\sigma_L$ are fixed positive numbers forming a wide and dense grid.
+Here we assume  $\sigma_1,\sigma_2,\ldots,\sigma_L$ are known and fixed positive numbers forming a wide and dense grid, which are built via a data-driven approach.
 
@@
 
-Model (\ref{eq:4}) is the extension of original Cormotif: set $L=1$ and assume $\hat s_{jr}^2$s are identical for all $j$, (\ref{eq:4}) is just Cormotif under Normal distribution.
+Model (\ref{eq:4}) is the extension of original  \textit{Cormotif}: set $L=1$ and assume for any fixed $r$, $\hat s_{jr}^2$s are identical for all $j$, (\ref{eq:4}) is just Cormotif under Normal distribution.
 Similar to Cormotif, our model can capture the correlation among multiple studies.
@@
 
+When setting $K =1$, our model is equivalent to applying {\tt ash} \cite{ash} to each study separately.
+The advantage of our model is that it allows correlations among studies.
+This advantage leads to higher accuracy for effect estimation in multiple studies.
+
 Another advantage of our generic model is that we could compute the posterior distribution for effects $\beta_{jr}$. By Bayes theorem, we have
@@
     p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}) \propto p(\beta_{jr}|\hat s_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})
+    \label{eq:5}
 \end{equation}
@@
 \subsection{Fitting the model}
+
+Our method involves three steps:
+\begin{enumerate}
+    \item For a given $K$, estimate the parameters $\pi$ and $g$ by maximizing the likelihood $L(\pi,g)$, given by (\ref{eq:4}).
+    \item Choose the best $K$ by minimizing the Bayesian Information Criterion (BIC).
+    \item Compute, for each $j$ and $r$, the posterior distribution $p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)$ by (\ref{eq:5}).
+\end{enumerate}
+
+All the details can be found in Section (\ref{sec:detail}).
+Step 1 is a convex problem, and we apply a EM algorithm to solve it.
+Step 2 is straightforward.
+The conditional distributions $p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)$ by (\ref{eq:5}) in Step 3 are analytically available, each a mixture of a point mass on zero and $L$ normal distribution.
+
+\subsection{Local False Sign Rate}
+
+To measure "significance" of an effect $\beta_{jr}$ we use the local false sign rate ($\lfsr$), which is defined as:
+\begin{equation}
+    \lfsr_{jr} := \min\{p(\beta_{jr}\geq 0|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g), p(\beta_{jr}\leq 0|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)\}
+\end{equation}
+
+Intutively, $\lfsr_{jr}$ is the probability that we would get the sign of effect $\beta_{jr}$ incorrect if we were to use our best guess of the sign.
+Therefore, a small $\lfsr$ indicates high confidence in determining the sign of an effect.
+Notice that $\lfsr$ is more conservative than the local false discovery rate ($\lfdr$), since we can infer $\lfsr_{jr}\geq\lfdr_{jr}$ from the definition.
 
@@
 
-\subsection{Implementation Details}
+\subsection{Implementation Details} \label{sec:detail}
 
@@
 
+\subsubsection{Posterior distribution}
+
+For a more general case, suppose
+\begin{equation}
+g_{kr}(\cdot;w) = \sum_{l=0}^L w_l^{kr}f_l(\cdot),
+\label{eq:geng}
+\end{equation}
+where $f_0$ is a point mass on 0. Then we have 
+\begin{equation}
+    p(\beta_{jr}|\pi,g,\hat s) = \sum_{k=1}^K\pi_k g_{kr}(\beta_{jr})=  \sum_{k=1}^K\pi_k\sum_{l=0}^Lw_l^{kr}f_l(\beta_{jr})
+    = \sum_{l=0}^L\eta_{lr}f_l(\beta_{jr}),
+\end{equation}
+where $\eta_{lr}=\sum_{k=1}^K\pi_kw_l^{kr}$. Thus, by (\ref{eq:5}), we get
+\begin{equation}
+    p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\pi, g) = \sum_{l=0}^L\theta_{ljr}p_{l}(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}).
+\end{equation}
+The posterior mixture component $p_l$ is the posterior on $\beta_{jr}$ that would be obtained using prior $f_l(\beta_{jr})$ and likelihood $p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})$.
+To compute the mixture weights $\theta_{ljr}$, denote
+\begin{equation}
+    \tilde f_{ljr}(\hat\beta_{jr}) := \int f_l(\beta_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})d\beta_{jr},
+\end{equation}
+the convolution of $f_l$ and a density function. Then $\theta_{ljr}$ can be obtained from
+\begin{equation}
+    \theta_{ljr} = \frac{\eta_{lr}\tilde f_{ljr}(\hat\beta_{jr})}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}.
+\end{equation}
+
+\begin{thebibliography}{99}
+
+\bibitem{cormotif} 
+Wei, Y. and Ji, H. (2014). Joint analysis of differential gene expression in multiple studies using correlation motif. {\it Biostatistics}, doi:10.1093/biostatistics/
+kxu038.
+
+\bibitem{ash} 
+Stephens, M. (2016). False discovery rate: A new deal. {\it Biostatistics} 18 (2): 275-294
+
+\bibitem{mash}
+Urbut, S.M., Wang, G. and Stephens, M. (2017). Flexible statistical methods for estimating and testing effects in genomic studies with multiple conditions. {\it bioRxiv}, doi:10.1101/096552.
+
+\end{thebibliography}
+
 \end{document}

### 2017-09-25 analysis/Vignette.Rmd
wflow_publish("Vignette.Rmd")

--- 2017-09-10 analysis/Simulation.Rmd
@@
 ---
-title: "Simulation for Correlation Motif"
+title: "Generic Cormotif Vignette"
 author: "Zhiwei Ma"
-date: "8/30/2017"
+date: "9/24/2017"
 output: html_document
 ---
+
 <!-- The file analysis/chunks.R contains chunks that define default settings
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+## Data preparation
+
+## Model Fitting
+
+## Session information
+
+<!-- Insert the session information into the document -->
+```{r session-info}
 ```
-
-## Model framework
-
-The framework of Correlation Motif model is based on my write-up document and supplementary matrials by $Wei$ $and$ $Ji$ (2015).
-In $Wei$ $and$ $Ji$ (2015), they implemented an EM algorithm and introduced priors for both $\pi$ and $Q$.
-As a result, their iterative formulas for new $\pi$ and new $Q$ are:
-
-\begin{eqnarray}
-\pi_k^{(t+1)} = \frac{\sum_{i=1}^n p_{ik}+1}{n+K} , \\
-q_{kr}^ {(t+1)}=\frac{\sum_{i=1}^n p_{ikr1}+1}{\sum_{i=1}^n p_{ik}+2}.
-\end{eqnarray}
-
-However, one can apply EM algorithm without introducing a prior and obtain a more interpretable formula:
-
-\begin{eqnarray}
-\pi_k^{(t+1)} = \frac{1}{n} \sum_{i=1}^n p_{ik}, \\
-q_{kr}^ {(t+1)}=\frac{\sum_{i=1}^n p_{ikr1}}{\sum_{i=1}^n p_{ik}}.
-\end{eqnarray}
-
-Here $p_{ik}= Pr(z_i=k|X_i,\pi^{(t)}, Q^ {(t)})$ is the posterior probability for gene $i$ belonging to class $k$, and the mean of $p_{ik}$ over $i$ is a reasonable estimations for $\pi_{k}$.
-The argument is similar for $q_{kr}$.
-Thus, we could interpret the underlying meaning in our iterative formulas.
-
-In fact, since $G$ is relatively large and $G \gg K$, the two formulas tend to give very similar results.
-We will illustrate this idea by implementing these two algorithms seperately on the same dataset: $simudata2$ in R package $Cormotif$.
-
-Notice that right now we are using the model applying limma ($Smyth$, 2004). The results with prior is
-
-```{r,cache=TRUE,message=FALSE,warning=FALSE}
-library(Cormotif)
-data(simudata2)
-exprs.simu2<-as.matrix(simudata2[,2:25])
-data(simu2_groupid)
-data(simu2_compgroup)
-
-motif.fitted<-cormotiffit(exprs.simu2,simu2_groupid,simu2_compgroup,
-                          K=1:5,max.iter=1000,BIC=TRUE)
-motif.fitted$bic
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-The results without prior is
-
-```{r,echo=FALSE}
-## Fit limma model
-limmafit<-function(exprs,groupid,compid){
-	compnum<-nrow(compid)
-	genenum<-nrow(exprs)	
-	limmat<-matrix(0,genenum,compnum)
-	limmas2<-rep(0,compnum)
-	limmadf<-rep(0,compnum)
-	limmav0<-rep(0,compnum)
-	limmag1num<-rep(0,compnum)
-	limmag2num<-rep(0,compnum)
-
-	for(i in 1:compnum) {
-		selid1<-which(groupid == compid[i,1]) 
-		selid2<-which(groupid == compid[i,2])
-		eset<-new("ExpressionSet", exprs=cbind(exprs[,selid1],exprs[,selid2]))
-		g1num<-length(selid1)
-		g2num<-length(selid2)
-		designmat<-cbind(base=rep(1,(g1num+g2num)), delta=c(rep(0,g1num),rep(1,g2num)))
-		fit<-lmFit(eset,designmat)
-		fit<-eBayes(fit)
-		limmat[,i]<-fit$t[,2]
-		limmas2[i]<-fit$s2.prior
-		limmadf[i]<-fit$df.prior
-		limmav0[i]<-fit$var.prior[2]
-		limmag1num[i]<-g1num
-		limmag2num[i]<-g2num
-
-	}
-      limmacompnum<-nrow(compid)
-	result<-list(t=limmat, v0=limmav0, df0=limmadf, s20=limmas2, g1num=limmag1num, g2num=limmag2num,compnum=limmacompnum)
-}
-
-## Log-likelihood for moderated t under H0
-modt.f0.loglike<-function(x,df){
-	a<-dt(x, df, log=TRUE)
-	result<-as.vector(a)
-      	flag<-which(is.na(result)==TRUE)
-      	result[flag]<-0
-      	result
-}
-
-## Log-likelihood for moderated t under H1
-## param=c(df,g1num,g2num,v0)
-modt.f1.loglike<-function(x,param) {
-	df<-param[1]
-	g1num<-param[2]
-	g2num<-param[3]
-	v0<-param[4]
-	w<-sqrt(1+v0/(1/g1num+1/g2num))
-	dt(x/w, df, log=TRUE)-log(w)
-	a<-dt(x/w, df, log=TRUE)-log(w)
-      	result<-as.vector(a)
-      	flag<-which(is.na(result)==TRUE)
-      	result[flag]<-0
-      	result
-}
-
-## Correlation Motif Fit
-cmfit_1<-function(x, type, K=1, tol=1e-3, max.iter=100) {
-	## initialize
-	xrow<-nrow(x)
-	xcol<-ncol(x)
-	loglike0<-list()
-	loglike1<-list()
-	p<-rep(1,K)/K
-	q<-matrix(runif(K*xcol), K, xcol)
-	q[1,]<-rep(0.01,xcol)
-
-	## compute loglikelihood
-	for(i in 1:xcol) {
-		f0<-type[[i]][[1]]
-		f0param<-type[[i]][[2]]
-		f1<-type[[i]][[3]]
-		f1param<-type[[i]][[4]]
-		loglike0[[i]]<-f0(x[,i],f0param)
-		loglike1[[i]]<-f1(x[,i],f1param)	
-	}
-
-	## EM algorithm to get MLE of p and q
-	condlike<-list()
-	for(i in 1:xcol) {
-		condlike[[i]]<-matrix(0,xrow,K)
-	}
-
-	loglike.old <- -1e10
-	for(i.iter in 1:max.iter) {
-
-		err<-tol+1
-
-		## compute posterior cluster membership
-		clustlike<-matrix(0,xrow,K)
-		templike <- matrix(0,xrow,2)
-		for(j in 1:K) {
-			for(i in 1:xcol) {
-				templike[,1]<-log(q[j,i])+loglike1[[i]]
-				templike[,2]<-log(1-q[j,i])+loglike0[[i]]
-				tempmax<-pmax(templike[,1],templike[,2])
-				for(z in 1:2) {
-					templike[,z]<-exp(templike[,z]-tempmax)
-				}
-				tempsum<-templike[,1]+templike[,2]
-				clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
-				condlike[[i]][,j]<-templike[,1]/tempsum
-			}
-			clustlike[,j]<-clustlike[,j]+log(p[j])
-		}
-
-		tempmax<-apply(clustlike,1,max)
-		for(j in 1:K) {
-			clustlike[,j]<-exp(clustlike[,j]-tempmax)
-		}
-		tempsum<-apply(clustlike,1,sum)
-
-		## update motif occurrence rate
-		for(j in 1:K) {
-			clustlike[,j]<-clustlike[,j]/tempsum
-		}
-
-		p.new<-(apply(clustlike,2,sum))/(xrow)
-
-		## update motifs
-		q.new<-matrix(0, K, xcol)
-		for(j in 1:K) {
-			clustpsum<-sum(clustlike[,j])
-			for(i in 1:xcol) {
-				q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j]))/(clustpsum)
-			}
-		}
-
-		## evaluate convergence
-		err.p<-max(abs(p.new-p)/p)
-		err.q<-max(abs(q.new-q)/q)
-		err<-max(err.p, err.q)
-
-		## evaluate whether the log.likelihood increases
-
-		loglike.new<-(sum(tempmax+log(tempsum)))/xrow
-
-		p<-p.new
-		q<-q.new
-		loglike.old<-loglike.new
-
-		if(err<tol) {
-			break;
-		}
-	}
-	## compute posterior p
-	clustlike<-matrix(0,xrow,K)
-	for(j in 1:K) {
-		for(i in 1:xcol) {
-			templike[,1]<-log(q[j,i])+loglike1[[i]]
-			templike[,2]<-log(1-q[j,i])+loglike0[[i]]
-			tempmax<-pmax(templike[,1],templike[,2])
-			for(z in 1:2) {
-				templike[,z]<-exp(templike[,z]-tempmax)
-			}
-			tempsum<-templike[,1]+templike[,2]
-			clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
-			condlike[[i]][,j]<-templike[,1]/tempsum
-		}
-		clustlike[,j]<-clustlike[,j]+log(p[j])
-	}
-
-	tempmax<-apply(clustlike,1,max)
-	for(j in 1:K) {
-		clustlike[,j]<-exp(clustlike[,j]-tempmax)
-	}
-	tempsum<-apply(clustlike,1,sum)
-	for(j in 1:K) {
-		clustlike[,j]<-clustlike[,j]/tempsum
-	}
-
-	p.post<-matrix(0,xrow,xcol)
-	for(j in 1:K) {
-		for(i in 1:xcol) {
-			p.post[,i]<-p.post[,i]+clustlike[,j]*condlike[[i]][,j]
-		}
-	}
-
-	## return
-	#calculate back loglikelihood
-	loglike.old<-loglike.old#-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
-	loglike.old<-loglike.old*xrow
-	result<-list(p.post=p.post, motif.prior=p, motif.q=q, loglike=loglike.old)
-}
-
-generatetype<-function(limfitted)
-{
-	jtype<-list()
-	df<-limfitted$g1num+limfitted$g2num-2+limfitted$df0
-	for(j in 1:limfitted$compnum)
-	{
-  		jtype[[j]]<-list(f0=modt.f0.loglike, f0.param=df[j], f1=modt.f1.loglike, f1.param=c(df[j],limfitted$g1num[j],limfitted$g2num[j],limfitted$v0[j]))  
-  	}
-    	jtype
-}
-
-cormotiffit_1<-function(exprs,groupid,compid,K=1, tol=1e-3, max.iter=100,BIC=TRUE)
-{
-	limfitted<-limmafit(exprs,groupid,compid)
-        jtype<-generatetype(limfitted)
-	fitresult<-list()
-	for(i in 1:length(K))
-		fitresult[[i]]<-cmfit_1(limfitted$t,type=jtype,K=K[i],max.iter=max.iter,tol=tol)
-	bic<-rep(0,length(K))
-	aic<-rep(0,length(K))
-	loglike<-rep(0,length(K))
-	for(i in 1:length(K))
-			loglike[i]<-fitresult[[i]]$loglike
-	for(i in 1:length(K))
-			bic[i]<--2*fitresult[[i]]$loglike+(K[i]-1+K[i]*limfitted$compnum)*log(dim(exprs)[1])
-	for(i in 1:length(K))
-			aic[i]<--2*fitresult[[i]]$loglike+2*(K[i]-1+K[i]*limfitted$compnum)
-	if(BIC==TRUE)
-	{
-		bestflag=which(bic==min(bic))
-	}
-	else
-	{
-		bestflag=which(aic==min(aic))
-	}
-	result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
-			aic=cbind(K,aic),loglike=cbind(K,loglike),allresult=fitresult)
-	#result<-list(bestmotif=fitresult[[4]],bic=cbind(K,bic),
-	#             aic=cbind(K,aic),loglike=cbind(K,loglike))
-}
-
-plotMotif_1<-function(bestmotif,title="")
-{
-	  layout(matrix(1:2,ncol=2))
-          u<-1:dim(bestmotif$motif.q)[2]
-          v<-1:dim(bestmotif$motif.q)[1]
-          image(u,v,t(bestmotif$motif.q),
-          col=gray(seq(from=1,to=0,by=-0.1)),xlab="Study",yaxt = "n",
-		ylab="Corr. Motifs",main=paste(title,"pattern",sep=" "))
-	  axis(2,at=1:length(v))
-          for(i in 1:(length(u)+1))
-          {
-                abline(v=(i-0.5))
-          }
-          for(i in 1:(length(v)+1)) 
-          {
-                abline(h=(i-0.5))
-          }
-	  Ng=10000
-	  if(is.null(bestmotif$p.post)!=TRUE)
-		Ng=nrow(bestmotif$p.post)
-	  genecount=floor(bestmotif$motif.p*Ng)
-	  NK=nrow(bestmotif$motif.q)
-	  plot(0,0.7,pch=".",xlim=c(0,1.2),ylim=c(0.75,NK+0.25),
-		frame.plot=FALSE,axes=FALSE,xlab="No. of genes",ylab="", main=paste(title,"frequency",sep=" "))
-	  segments(0,0.7,bestmotif$motif.p[1],0.7)
-	  rect(0,1:NK-0.3,bestmotif$motif.p,1:NK+0.3,
-		col="dark grey")
-	  mtext(1:NK,at=1:NK,side=2,cex=0.8)
-	  text(bestmotif$motif.p+0.15,1:NK,
-	  labels=floor(bestmotif$motif.p*Ng))
-}
-```
-
-```{r,cache=TRUE}
-motif.fitted<-cormotiffit_1(exprs.simu2,simu2_groupid,simu2_compgroup,K=1:5,max.iter=1000,BIC=TRUE)
-
-motif.fitted$bic
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-The two methods do give very similar results.
-However in practice, it takes much less iterations for methods with prior to achieve its accuracy.
-Here we prefer to apply CorMotif with prior.
-In the rest of document, without specification, we apply CorMOtif with prior.
-
-## Simulation based on limma
-
-We first duplicated several simulation from $Wei$ $and$ $Ji$ (2015), which was still based on limma.
-
-### Simulation 1
-
-In simulation 1, we generated 10,000 genes and four studies according to the four different patterns: 100 genes were differentially expressed in all four studies ($Y_i = [1,1,1,1]$); 400 genes were differential only in studies 1 and 2 ($Y_i = [1,1,0,0]$); 400 genes were differential only in studies 2 and 3 ($Y_i = [0,1,1,0]$); 9100 genes were non-differential ($Y_i = [0,0,0,0]$).
-Each study had six samples: three cases and three controls.
-The variances $\sigma_{ir}^2$ were simulated from $n_{0r}s_{0r}^2/\chi^2(n_{0r})$, where $n_{0r}=4$ and $s_{0r}^2=0.02$.
-The expression values were generated using $x_{irlj}\sim N(0,\sigma_{ir}^2)$.
-When $y_{ir}=1$, drew $\mu_{ir}$ from $N(0,16\sigma_{ir}^2)$, and $\mu_{ir}$ was added to the three cases.
-
-Notice that in $Wei$ $and$ $Ji$ (2015), they drew $\mu_{ir}$ from $N(0,4\sigma_{ir}^2)$.
-I think it is a typo, since with this setting, the simulation results are quite unsatisfying and only two patterns were recognized.
-The reason for that is the difference between cases and controls will be minor if the variance is not large enough, which will weaken the pattern.
-
-```{r,cache=TRUE}
-K = 4
-D = 4
-G = 10000
-pi0 = c(0.01,0.04,0.04,0.91)
-q0 = matrix(0,nrow=K,ncol=D)
-q0[1,] = c(1,1,1,1)
-q0[2,] = c(1,1,0,0)
-q0[3,] = c(0,1,1,0)
-q0[4,] = c(0,0,0,0)
-X = matrix(0,nrow=G,ncol=6*D)
-rows = rep(1:K,times=pi0*G)
-A = q0[rows,]
-set.seed(1)
-for(g in 1:G){
-  for(d in 1:D){
-    sigma2 = 4*0.02/rchisq(1,4)
-    x = rnorm(6,0,sqrt(sigma2))
-    if(A[g,d]==1){
-      dx = rnorm(1,0,4*sqrt(sigma2))
-      x[1:3] = x[1:3]+dx
-    }
-    X[g,(6*d-5):(6*d)] = x
-  }
-}
-groupid = rep(1:8,each=3)
-compgroup = t(matrix(1:8,ncol=4))
-motif.fitted<-cormotiffit(X,groupid,compgroup,K=1:5,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-The result matches well with $Wei$ $and$ $Ji$ (2015).
-
-### Simulation 2
-
-In simulation 2, we generated 10,000 genes and four studies according to the four different patterns: 300 genes were differentially expressed in all four studies ($Y_i = [1,1,1,1]$); 300 genes were differential only in studies 1 and 2 ($Y_i = [1,1,0,0]$); 300 genes were differential only in studies 3 and 4 ($Y_i = [0,0,1,1]$); 9100 genes were non-differential ($Y_i = [0,0,0,0]$).
-
-```{r,cache=TRUE}
-K = 4
-D = 4
-G = 10000
-pi0 = c(0.03,0.03,0.03,0.91)
-q0 = matrix(0,nrow=K,ncol=D)
-q0[1,] = c(1,1,1,1)
-q0[2,] = c(1,1,0,0)
-q0[3,] = c(0,0,1,1)
-q0[4,] = c(0,0,0,0)
-X = matrix(0,nrow=G,ncol=6*D)
-rows = rep(1:K,times=pi0*G)
-A = q0[rows,]
-set.seed(2)
-for(g in 1:G){
-  for(d in 1:D){
-    sigma2 = 4*0.02/rchisq(1,4)
-    x = rnorm(6,0,sqrt(sigma2))
-    if(A[g,d]==1){
-      dx = rnorm(1,0,4*sqrt(sigma2))
-      x[1:3] = x[1:3]+dx
-    }
-    X[g,(6*d-5):(6*d)] = x
-  }
-}
-groupid = rep(1:8,each=3)
-compgroup = t(matrix(1:8,ncol=4))
-motif.fitted<-cormotiffit(X,groupid,compgroup,K=1:5,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-Then we could check if we change the proportion of four patterns to $200:50:50:9700$.
-
-```{r,cache=TRUE}
-K = 4
-D = 4
-G = 10000
-pi0 = c(0.02,0.005,0.005,0.97)
-q0 = matrix(0,nrow=K,ncol=D)
-q0[1,] = c(1,1,1,1)
-q0[2,] = c(1,1,0,0)
-q0[3,] = c(0,0,1,1)
-q0[4,] = c(0,0,0,0)
-X = matrix(0,nrow=G,ncol=6*D)
-rows = rep(1:K,times=pi0*G)
-A = q0[rows,]
-set.seed(1)
-for(g in 1:G){
-  for(d in 1:D){
-    sigma2 = 4*0.02/rchisq(1,4)
-    x = rnorm(6,0,sqrt(sigma2))
-    if(A[g,d]==1){
-      dx = rnorm(1,0,4*sqrt(sigma2))
-      x[1:3] = x[1:3]+dx
-    }
-    X[g,(6*d-5):(6*d)] = x
-  }
-}
-groupid = rep(1:8,each=3)
-compgroup = t(matrix(1:8,ncol=4))
-motif.fitted<-cormotiffit(X,groupid,compgroup,K=1:5,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-We could find that the second and third pattern merged togethor, since their proportions were relatively small compared to the first and forth pattern.
-
-### Simulation 3
-
-In simulation 3, we generated 10,000 genes and eight studies according to the four different patterns: 200 genes were differentially expressed in all four studies ($Y_i = [1,1,1,1,1,1,1,1]$); 200 genes were differential only in studies 1, 2, 3, 4 ($Y_i = [1,1,1,1,0,0,0,0]$); 200 genes were differential only in studies 5, 6, 7, 8 ($Y_i = [0,0,0,0,1,1,1,1]$); 200 genes were differential only in studies 3, 4, 5, 6 ($Y_i = [0,0,1,1,1,1,0,0]$); 9200 genes were non-differential ($Y_i = [0,0,0,0,0,0,0,0,0]$).
-
-```{r,cache=TRUE}
-K = 5
-D = 8
-G = 10000
-pi0 = c(0.02,0.02,0.02,0.02,0.92)
-q0 = matrix(0,nrow=K,ncol=D)
-q0[1,] = c(1,1,1,1,1,1,1,1)
-q0[2,] = c(1,1,1,1,0,0,0,0)
-q0[3,] = c(0,0,0,0,1,1,1,1)
-q0[4,] = c(0,0,1,1,1,1,0,0)
-q0[5,] = c(0,0,0,0,0,0,0,0)
-X = matrix(0,nrow=G,ncol=6*D)
-rows = rep(1:K,times=pi0*G)
-A = q0[rows,]
-set.seed(3)
-for(g in 1:G){
-  for(d in 1:D){
-    sigma2 = 4*0.02/rchisq(1,4)
-    x = rnorm(6,0,sqrt(sigma2))
-    if(A[g,d]==1){
-      dx = rnorm(1,0,4*sqrt(sigma2))
-      x[1:3] = x[1:3]+dx
-    }
-    X[g,(6*d-5):(6*d)] = x
-  }
-}
-groupid = rep(1:16,each=3)
-compgroup = t(matrix(1:16,nrow=2))
-motif.fitted<-cormotiffit(X,groupid,compgroup,K=1:6,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-CorMotif tends to merge weaker patterns with small proportion.
-Just as $Wei$ $and$ $Ji$ (2015) states, correlation motifs only represent a parsimonious representation of the correlation structure supported by the available data.
-On the other hand, the two stage method may also be blame.
-Adding $limma$ structure can increase inaccuracy for estimation.
-In the next section, we will generate $x_{ir}$ directly from a Gaussian distribution.
-In fact, any distribution can be applied under our framework.
-In addition, we can also estimate the parameters for Gaussian distribution.
-
-## CorMotif based on Gaussian model
-
-Let
-
-\begin{eqnarray*}
-f_{r0}(x_{ir}) &=& N(x_{ir};0,1), \\
-f_{r1}(x_{ir}) &=& N(x_{ir};0,1+\sigma_r^2).
-\end{eqnarray*}
-
-First, we checked the performance of CorMotif by assuming we know $\sigma^2$ exactly. Then we dealed with the situation where we were given no information about $\sigma^2$. Apply
-
-$$\sigma_r^{2(t+1)}= \frac{\sum_{i=1}^n \sum_{k=1}^K (x_{ir}^2-1)p_{ikr1}}{\sum_{i=1}^n \sum_{k=1}^K p_{ikr1}},$$
-
-from my write-up document, we could iteratively update $\sigma_r^2$.
-
-### Simulation 4
-
-We used the same patterns and proportions as Simulation 1 in the previous section and let $\sigma^2=(4^2,4^2,4^2,4^2)$. Assume we know $\sigma^2$, the estimated results were:
-
-```{r,echo=FALSE}
-## Log-likelihood for norm under H0 and H1
-modt.f.loglike<-function(x,m,s2) {
-	a<-dnorm(x, m, sqrt(s2), log=TRUE)
-	result<-as.vector(a)
-  flag<-which(is.na(result)==TRUE)
-  result[flag]<-0
-  result
-}
-
-```
-
-```{r,echo=FALSE}
-cmfit.norm.known<-function(x, sigma2, K=1, tol=1e-3, max.iter=100) {
-	## initialize
-	xrow<-nrow(x)
-	xcol<-ncol(x)
-	loglike0<-list()
-	loglike1<-list()
-	p<-rep(1,K)/K
-	q<-matrix(runif(K*xcol), K, xcol)
-	q[1,]<-rep(0.01,xcol)
-
-	## compute loglikelihood
-	for(i in 1:xcol) {
-  		loglike0[[i]]<-modt.f.loglike(x[,i],0,1)
-  		loglike1[[i]]<-modt.f.loglike(x[,i],0,1+sigma2[i])	
-  	}
-
-	## EM algorithm to get MLE of p, q and sigma2
-	condlike<-list()
-	for(i in 1:xcol) {
-		condlike[[i]]<-matrix(0,xrow,K)
-	}
-
-	loglike.old <- -1e10
-	for(i.iter in 1:max.iter) {
-		err<-tol+1
-
-		## compute posterior cluster membership
-		clustlike<-matrix(0,xrow,K)
-		templike <- matrix(0,xrow,2)
-		for(j in 1:K) {
-			for(i in 1:xcol) {
-				templike[,1]<-log(q[j,i])+loglike1[[i]]
-				templike[,2]<-log(1-q[j,i])+loglike0[[i]]
-				tempmax<-pmax(templike[,1],templike[,2])
-				for(z in 1:2) {
-					templike[,z]<-exp(templike[,z]-tempmax)
-				}
-				tempsum<-templike[,1]+templike[,2]
-				clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
-				condlike[[i]][,j]<-templike[,1]/tempsum
-			}
-			clustlike[,j]<-clustlike[,j]+log(p[j])
-		}
-
-		tempmax<-apply(clustlike,1,max)
-		for(j in 1:K) {
-			clustlike[,j]<-exp(clustlike[,j]-tempmax)
-		}
-		tempsum<-apply(clustlike,1,sum)
-
-		## update motif occurrence rate
-		for(j in 1:K) {
-			clustlike[,j]<-clustlike[,j]/tempsum
-		}
-
-		#p.new<-(apply(clustlike,2,sum))/(xrow)
-		p.new<-(apply(clustlike,2,sum)+1)/(xrow+K)
-		## update motifs
-		q.new<-matrix(0, K, xcol)
-		for(j in 1:K) {
-			clustpsum<-sum(clustlike[,j])
-			for(i in 1:xcol) {
-				#q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j]))/(clustpsum)
-				q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j])+1)/(clustpsum+2)
-			}
-		}
-		## evaluate convergence
-		err.p<-max(abs(p.new-p)/p)
-		err.q<-max(abs(q.new-q)/q)
-		err<-max(err.p, err.q)
-
-		## evaluate whether the log.likelihood increases
-		loglike.new<-(sum(tempmax+log(tempsum))+sum(log(p.new))+sum(log(q.new)+log(1-q.new)))/xrow
-		#loglike.new<-(sum(tempmax+log(tempsum)))/xrow
-
-		p<-p.new
-		q<-q.new
-		loglike.old<-loglike.new
-
-		if(err<tol) {
-			break;
-		}
-	}
-	## compute posterior p
-	clustlike<-matrix(0,xrow,K)
-	for(j in 1:K) {
-		for(i in 1:xcol) {
-			templike[,1]<-log(q[j,i])+loglike1[[i]]
-			templike[,2]<-log(1-q[j,i])+loglike0[[i]]
-			tempmax<-pmax(templike[,1],templike[,2])
-			for(z in 1:2) {
-				templike[,z]<-exp(templike[,z]-tempmax)
-			}
-			tempsum<-templike[,1]+templike[,2]
-			clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
-			condlike[[i]][,j]<-templike[,1]/tempsum
-		}
-		clustlike[,j]<-clustlike[,j]+log(p[j])
-	}
-
-	tempmax<-apply(clustlike,1,max)
-	for(j in 1:K) {
-		clustlike[,j]<-exp(clustlike[,j]-tempmax)
-	}
-	tempsum<-apply(clustlike,1,sum)
-	for(j in 1:K) {
-		clustlike[,j]<-clustlike[,j]/tempsum
-	}
-
-	p.post<-matrix(0,xrow,xcol)
-	for(j in 1:K) {
-		for(i in 1:xcol) {
-			p.post[,i]<-p.post[,i]+clustlike[,j]*condlike[[i]][,j]
-		}
-	}
-
-	## return
-	#calculate back loglikelihood
-	loglike.old<-loglike.old-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
-	#loglike.old<-loglike.old#-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
-	loglike.old<-loglike.old*xrow
-  result<-list(p.post=p.post, motif.prior=p, motif.q=q,loglike=loglike.old,motif.s = sigma2)
-}
-
-cormotiffit.norm.known<-function(x,sigma2,K=1, tol=1e-3, max.iter=100,BIC=TRUE)
-{
-	fitresult<-list()
-	for(i in 1:length(K))
-		fitresult[[i]]<-cmfit.norm.known(x,sigma2,K=K[i],max.iter=max.iter,tol=tol)
-	bic<-rep(0,length(K))
-	aic<-rep(0,length(K))
-	loglike<-rep(0,length(K))
-	for(i in 1:length(K))
-			loglike[i]<-fitresult[[i]]$loglike
-	for(i in 1:length(K))
-			bic[i]<--2*fitresult[[i]]$loglike+(K[i]-1+K[i]*ncol(x))*log(dim(x)[1])
-	for(i in 1:length(K))
-			aic[i]<--2*fitresult[[i]]$loglike+2*(K[i]-1+K[i]*ncol(x))
-	if(BIC==TRUE)
-	{
-		bestflag=which(bic==min(bic))
-	}
-	else
-	{
-		bestflag=which(aic==min(aic))
-	}
-	#result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
-	#		aic=cbind(K,aic),loglike=cbind(K,loglike))
-	result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
-	             aic=cbind(K,aic),loglike=cbind(K,loglike),allresult=fitresult)
-} 
-```
-
-```{r,cache=TRUE}
-K = 4
-D = 4
-G = 10000
-sigma2 = c(16,16,16,16)
-pi0 = c(0.01,0.04,0.04,0.91)
-q0 = matrix(0,nrow=K,ncol=D)
-q0[1,] = c(1,1,1,1)
-q0[2,] = c(1,1,0,0)
-q0[3,] = c(0,1,1,0)
-q0[4,] = c(0,0,0,0)
-X = matrix(0,nrow=G,ncol=D)
-rows = rep(1:K,times=pi0*G)
-A = q0[rows,]
-set.seed(4)
-for(g in 1:G){
-  for(d in 1:D){
-    if(A[g,d]==1){
-      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
-    } else{
-      X[g,d] = rnorm(1,0,1)
-    }
-  }
-}
-motif.fitted<-cormotiffit.norm.known(X,sigma2,K=1:5,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-CorMotif could correctly recognize four patterns and the proportion is also reasonable. Then we checked the performace without any prior information about $\sigma^2$ with the same simulated data.
-
-```{r,echo=FALSE}
-cmfit.norm<-function(x, K=1, tol=1e-3, max.iter=100) {
-	## initialize
-	xrow<-nrow(x)
-	xcol<-ncol(x)
-	loglike0<-list()
-	loglike1<-list()
-	p<-rep(1,K)/K
-	q<-matrix(runif(K*xcol), K, xcol)
-	q[1,]<-rep(0.01,xcol)
-	sigma2 = rep(1,xcol)
-#	sigma2 = c(16,16,16,16)
-
-	## EM algorithm to get MLE of p, q and sigma2
-	condlike<-list()
-	for(i in 1:xcol) {
-		condlike[[i]]<-matrix(0,xrow,K)
-	}
-
-	loglike.old <- -1e10
-
-	for(i.iter in 1:max.iter) {
-
-	  	err<-tol+1
-		## compute loglikelihood
-  	for(i in 1:xcol) {
-    		loglike0[[i]]<-modt.f.loglike(x[,i],0,1)
-    		loglike1[[i]]<-modt.f.loglike(x[,i],0,1+sigma2[i])	
-    	}
-		## compute posterior cluster membership
-		clustlike<-matrix(0,xrow,K)
-		templike <- matrix(0,xrow,2)
-		for(j in 1:K) {
-			for(i in 1:xcol) {
-				templike[,1]<-log(q[j,i])+loglike1[[i]]
-				templike[,2]<-log(1-q[j,i])+loglike0[[i]]
-				tempmax<-pmax(templike[,1],templike[,2])
-				for(z in 1:2) {
-					templike[,z]<-exp(templike[,z]-tempmax)
-				}
-				tempsum<-templike[,1]+templike[,2]
-				clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
-				condlike[[i]][,j]<-templike[,1]/tempsum
-			}
-			clustlike[,j]<-clustlike[,j]+log(p[j])
-		}
-
-		tempmax<-apply(clustlike,1,max)
-		for(j in 1:K) {
-			clustlike[,j]<-exp(clustlike[,j]-tempmax)
-		}
-		tempsum<-apply(clustlike,1,sum)
-
-		## update motif occurrence rate
-		for(j in 1:K) {
-			clustlike[,j]<-clustlike[,j]/tempsum
-		}
-
-		#p.new<-(apply(clustlike,2,sum))/(xrow)
-		p.new<-(apply(clustlike,2,sum)+1)/(xrow+K)
-		## update motifs
-		q.new<-matrix(0, K, xcol)
-		for(j in 1:K) {
-			clustpsum<-sum(clustlike[,j])
-			for(i in 1:xcol) {
-				#q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j]))/(clustpsum)
-				q.new[j,i]<-(sum(clustlike[,j]*condlike[[i]][,j])+1)/(clustpsum+2)
-			}
-		}
-    ## update sigma2
-		sigma2.new<-rep(0,xcol)
-		for(i in 1:xcol){
-		  pikr1 = clustlike*condlike[[i]]
-		  xx = x[,i]^2-1
-		  sigma2.new[i] = sum(t(xx)%*%pikr1)/sum(pikr1)
-		}
-		## evaluate convergence
-		err.p<-max(abs(p.new-p)/p)
-		err.q<-max(abs(q.new-q)/q)
-		err.sigma2<-max(abs((sigma2.new-sigma2)/sigma2))
-		err<-max(err.p, err.q, err.sigma2)
-    #err<-max(err.p, err.q)
-		## evaluate whether the log.likelihood increases
-		loglike.new<-(sum(tempmax+log(tempsum))+sum(log(p.new))+sum(log(q.new)+log(1-q.new)))/xrow
-		#loglike.new<-(sum(tempmax+log(tempsum)))/xrow
-
-		p<-p.new
-		q<-q.new
-		sigma2<-sigma2.new
-		loglike.old<-loglike.new
-
-		if(err<tol) {
-			break;
-		}
-	}
-	## compute posterior p
-	clustlike<-matrix(0,xrow,K)
-	for(j in 1:K) {
-		for(i in 1:xcol) {
-			templike[,1]<-log(q[j,i])+loglike1[[i]]
-			templike[,2]<-log(1-q[j,i])+loglike0[[i]]
-			tempmax<-pmax(templike[,1],templike[,2])
-			for(z in 1:2) {
-				templike[,z]<-exp(templike[,z]-tempmax)
-			}
-			tempsum<-templike[,1]+templike[,2]
-			clustlike[,j]<-clustlike[,j]+tempmax+log(tempsum)
-			condlike[[i]][,j]<-templike[,1]/tempsum
-		}
-		clustlike[,j]<-clustlike[,j]+log(p[j])
-	}
-
-	tempmax<-apply(clustlike,1,max)
-	for(j in 1:K) {
-		clustlike[,j]<-exp(clustlike[,j]-tempmax)
-	}
-	tempsum<-apply(clustlike,1,sum)
-	for(j in 1:K) {
-		clustlike[,j]<-clustlike[,j]/tempsum
-	}
-
-	p.post<-matrix(0,xrow,xcol)
-	for(j in 1:K) {
-		for(i in 1:xcol) {
-			p.post[,i]<-p.post[,i]+clustlike[,j]*condlike[[i]][,j]
-		}
-	}
-
-	## return
-	#calculate back loglikelihood
-	loglike.old<-loglike.old-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
-	#loglike.old<-loglike.old#-(sum(log(p))+sum(log(q)+log(1-q)))/xrow
-	loglike.old<-loglike.old*xrow
-  result<-list(p.post=p.post, motif.prior=p, motif.q=q,loglike=loglike.old,motif.s = sigma2)
-}
-
-cormotiffit.norm<-function(x,K=1, tol=1e-3, max.iter=100,BIC=TRUE)
-{
-	fitresult<-list()
-	for(i in 1:length(K))
-		fitresult[[i]]<-cmfit.norm(x,K=K[i],max.iter=max.iter,tol=tol)
-	bic<-rep(0,length(K))
-	aic<-rep(0,length(K))
-	loglike<-rep(0,length(K))
-	for(i in 1:length(K))
-			loglike[i]<-fitresult[[i]]$loglike
-	for(i in 1:length(K))
-			bic[i]<--2*fitresult[[i]]$loglike+(K[i]-1+K[i]*ncol(x))*log(dim(x)[1])
-	for(i in 1:length(K))
-			aic[i]<--2*fitresult[[i]]$loglike+2*(K[i]-1+K[i]*ncol(x))
-	if(BIC==TRUE)
-	{
-		bestflag=which(bic==min(bic))
-	}
-	else
-	{
-		bestflag=which(aic==min(aic))
-	}
-	#result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
-	#		aic=cbind(K,aic),loglike=cbind(K,loglike))
-	result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
-	             aic=cbind(K,aic),loglike=cbind(K,loglike),allresult=fitresult)
-} 
-```
-
-```{r,cache=TRUE}
-motif.fitted<-cormotiffit.norm(X,K=1:5,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-And the estimated $\sigma^2$ is
-
-```{r,cache=TRUE}
-motif.fitted$bestmotif$motif.s
-```
-
-### Simulation 5
-
-We used the same patterns and proportions as Simulation 2 in the previous section and let $\sigma^2=(4^2,4^2,4^2,4^2)$. Assume we know $\sigma^2$, the estimated results were:
-
-```{r,cache=TRUE}
-K = 4
-D = 4
-G = 10000
-sigma2 = c(16,16,16,16)
-pi0 = c(0.03,0.03,0.03,0.91)
-q0 = matrix(0,nrow=K,ncol=D)
-q0[1,] = c(1,1,1,1)
-q0[2,] = c(1,1,0,0)
-q0[3,] = c(0,0,1,1)
-q0[4,] = c(0,0,0,0)
-X = matrix(0,nrow=G,ncol=D)
-rows = rep(1:K,times=pi0*G)
-A = q0[rows,]
-set.seed(4)
-for(g in 1:G){
-  for(d in 1:D){
-    if(A[g,d]==1){
-      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
-    } else{
-      X[g,d] = rnorm(1,0,1)
-    }
-  }
-}
-motif.fitted<-cormotiffit.norm.known(X,sigma2,K=1:5,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-Again, CorMotif could correctly recognize four patterns and the proportion is also accurate. Then we checked the performace without any prior information about $\sigma^2$ with the same simulated data.
-
-```{r,cache=TRUE}
-motif.fitted<-cormotiffit.norm(X,K=1:5,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-And the estimated $\sigma^2$ is
-
-```{r,cache=TRUE}
-motif.fitted$bestmotif$motif.s
-```
-
-### Simulation 6
-
-We used the same patterns and proportions as Simulation 3 in the previous section and let $\sigma^2=(4^2,4^2,4^2,4^2,4^2,4^2,4^2,4^2)$. Assume we know $\sigma^2$, the estimated results were:
-
-```{r,cache=TRUE}
-K = 5
-D = 8
-G = 10000
-sigma2 = rep(16,D)
-pi0 = c(0.02,0.02,0.02,0.02,0.92)
-q0 = matrix(0,nrow=K,ncol=D)
-q0[1,] = c(1,1,1,1,1,1,1,1)
-q0[2,] = c(1,1,1,1,0,0,0,0)
-q0[3,] = c(0,0,0,0,1,1,1,1)
-q0[4,] = c(0,0,1,1,1,1,0,0)
-q0[5,] = c(0,0,0,0,0,0,0,0)
-X = matrix(0,nrow=G,ncol=D)
-rows = rep(1:K,times=pi0*G)
-A = q0[rows,]
-for(g in 1:G){
-  for(d in 1:D){
-    if(A[g,d]==1){
-      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
-    } else{
-      X[g,d] = rnorm(1,0,1)
-    }
-  }
-}
-motif.fitted<-cormotiffit.norm.known(X,sigma2,K=1:6,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-Then we checked the performace without any prior information about $\sigma^2$ with the same simulated data.
-
-```{r,cache=TRUE}
-motif.fitted<-cormotiffit.norm(X,K=1:6,max.iter=1000,BIC=TRUE)
-plotIC(motif.fitted)
-plotMotif(motif.fitted)
-```
-
-And the estimated $\sigma^2$ is
-
-```{r,cache=TRUE}
-motif.fitted$bestmotif$motif.s
-```
-
-Under all situation, CorMotif without knowing $\sigma^2$ performed equally well with knowing $\sigma^2$.
-
-## Discussion
-
-We tested CorMotif for both $limma$ and Guassion model.
-Through our simulation, we find that correlation motifs only represent a parsimonious representation of the correlation structure.
-The reason is twofold.
-On the one hand, CorMotif tends to merge patterns with small proportion.
-On the other hand, if the differences between cases and controls (for Gaussian model, the variance $\sigma^2$) are not significant, CorMotif cannot detect the pattern either.

### 2017-09-25 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

---
title: "Untitled"
author: "First Last"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
```

### Simulation

We set unit number $n = 5000$, and study number $R = 4$. We set every observation has the same standard error, $s_{jr}=1$. That is, $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1) $. the 5000 units come from 4 patterns ($K=4$): 2000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 2000 genes have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.

```{r,cache=TRUE}
source('~/Git/cormotif/code/function.R')
K = 4
D = 4
G = 5000
sigma2 = c(100,100,100,100)
pi0 = c(0.02,0.04,0.04,0.9)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,1,1,0)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(111)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
betahat=X
sebetahat=matrix(1,nrow=G,ncol=D)
fit = generic.cormotif(betahat,sebetahat,K=1:5)
```

After fitting the generic cormotif model, we can check the BIC values obtained by all cluster numbers:

```{r,cache=TRUE}
fit$bic
```

$K = 4$ has the smallest BIC. The standard deviation built by our method is

```{r,cache=TRUE}
fit$mixsd
```

And the estimated $\pi$ for $K = 4$ is

```{r,cache=TRUE}
fit$bestmotif$pi
```

Then we could check the estimation for $g$, i.e. the values of $w$. The estimated $w$ for studies 1 is

```{r,cache=TRUE}
fit$bestmotif$W[[1]]
```

The estimated $w$ for studies 2 is

```{r,cache=TRUE}
fit$bestmotif$W[[2]]
```

The estimated $w$ for studies 3 is

```{r,cache=TRUE}
fit$bestmotif$W[[3]]
```

The estimated $w$ for studies 4 is

```{r,cache=TRUE}
fit$bestmotif$W[[4]]
```

All of them matches well with our setting, which verifies our algorithm.

Also, we could compute the \textit{lfdr}:

```{r,cache=TRUE}
head(cbind(fit$result[[1]]$lfdr,fit$result[[2]]$lfdr,fit$result[[3]]$lfdr,fit$result[[4]]$lfdr))
```

And \textit{lfsr}:

```{r,cache=TRUE}
head(cbind(fit$result[[1]]$lfsr,fit$result[[2]]$lfsr,fit$result[[3]]$lfsr,fit$result[[4]]$lfsr))
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-09-25 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

@@
 ---
-title: "Untitled"
+title: "Simulation for generic Cormotif model"
 author: "First Last"
-date: YYYY-MM-DD
+date: "9/24/2017"
 output: html_document
@@
 We set every observation has the same standard error, $s_{jr}=1$.
-That is, $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1) $.
-the 5000 units come from 4 patterns ($K=4$): 2000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 2000 genes have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
+That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
+The 5000 units come from 4 patterns ($K=4$): 2000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 2000 genes have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
 

### 2017-09-25 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

@@
 
-### Simulation
+## Simulation
 
@@
 That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 5000 units come from 4 patterns ($K=4$): 2000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 2000 genes have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
+The 5000 units come from 4 patterns ($K=4$): 2000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 1000 unitss have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
 
@@
 
-Also, we could compute the \textit{lfdr}:
+Also, we could compute the lfdr:
 
@@
 
-And \textit{lfsr}:
+And lfsr:
 

### 2017-09-25 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
+knitr::opts_chunk$set(echo = TRUE,eval=TRUE)
 ```
@@
 sigma2 = c(100,100,100,100)
-pi0 = c(0.02,0.04,0.04,0.9)
+pi0 = c(0.2,0.2,0.2,0.4)
 q0 = matrix(0,nrow=K,ncol=D)

### 2017-09-25 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

@@
 
-```{r,cache=TRUE}
+```{r}
 fit$bic
@@
 
-$K = 4$ has the smallest BIC. The standard deviation built by our method is
+$K = 4$ has the smallest BIC. The standard deviation $\sigma_0,\sigma_1,\ldots,\sigma_L$ built by our method is
 
-```{r,cache=TRUE}
+```{r}
 fit$mixsd
@@
 
-```{r,cache=TRUE}
+```{r}
 fit$bestmotif$pi
@@
 
-```{r,cache=TRUE}
+```{r}
 fit$bestmotif$W[[1]]
@@
 
-```{r,cache=TRUE}
+```{r}
 fit$bestmotif$W[[2]]
@@
 
-```{r,cache=TRUE}
+```{r}
 fit$bestmotif$W[[3]]
@@
 
-```{r,cache=TRUE}
+```{r}
 fit$bestmotif$W[[4]]
@@
 
-```{r,cache=TRUE}
+```{r}
 head(cbind(fit$result[[1]]$lfdr,fit$result[[2]]$lfdr,fit$result[[3]]$lfdr,fit$result[[4]]$lfdr))
@@
 
-```{r,cache=TRUE}
+```{r}
 head(cbind(fit$result[[1]]$lfsr,fit$result[[2]]$lfsr,fit$result[[3]]$lfsr,fit$result[[4]]$lfsr))

### 2017-09-25 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

@@
 sebetahat=matrix(1,nrow=G,ncol=D)
+```
+
+```{r,cache=TRUE}
 fit = generic.cormotif(betahat,sebetahat,K=1:5)

### 2017-09-25 code/Untitled.R
add main function

@@ -1,0 +1,184 @@
+library("SQUAREM")
+
+mixcompdist = "normal"
+optmethod = "w_mixEM"
+df = NULL
+nullweight = 10
+pointmass = TRUE
+mixsd = NULL
+gridmult = sqrt(2)
+outputlevel = 2
+g = NULL
+fixg = FALSE
+mode = 0
+alpha = 0
+grange = c(-Inf,Inf)
+control = list()
+lik = NULL
+weights=NULL
+pi_thresh = 1e-10
+
+n = 2000
+R = 5
+K = 3
+
+#betahat = matrix(rnorm(n*R,0,sqrt(5)),ncol=R)
+betahat = matrix(0,nrow=n,ncol=R)
+betahat[1:(n/10),] = matrix(rnorm(n*R/10,0,sqrt(301)),ncol=R)
+betahat[(n/10+1):(n/5),] = matrix(rnorm(n*R/10,0,sqrt(37)),ncol=R)
+betahat[(n/5+1):n,] = matrix(rnorm(4*n*R/5,0,1),ncol=R)
+#betahat[1:(n/4),] = matrix(rnorm(n*R/4,0,sqrt(101)),ncol=R)
+#betahat[(n/4+1):n,] = matrix(rnorm(3*n*R/4,0,1),ncol=R)
+sebetahat = matrix(1,nrow=n,ncol=R)
+
+
+source('~/Git/ashr/R/lik.R')
+# set likelihood based on defaults if missing
+if(is.null(lik)){ 
+  if(is.null(df)){
+    lik = lik_normal()
+  } else {lik = lik_t(df)}
+}
+
+## determine mixsd
+betahat_v = as.vector(betahat)
+sebetahat_v = as.vector(sebetahat)
+source('~/Git/ashr/R/set_data.R')
+data_v = set_data(betahat_v, sebetahat_v, lik, alpha)
+source('~/Git/ashr/R/ash.R')
+if(is.null(mixsd)){
+  mixsd = autoselect.mixsd(data_v,gridmult,mode,grange,mixcompdist)
+}
+if(pointmass){
+  mixsd = c(0,mixsd)
+}
+L = length(mixsd)
+null.comp = which.min(mixsd)
+
+# compute conv_dens matrix
+source('~/Git/ashr/R/mix.R')
+source('~/Git/ashr/R/normalmix.R')
+source('~/Git/ashr/R/mix_opt.R')
+matrix_llik = list()
+for(r in 1:R){
[…]

### 2017-09-25 code/function.R

@@ -1,0 +1,179 @@
+library("SQUAREM")
+
+source('~/Git/ashr/R/lik.R')
+source('~/Git/ashr/R/set_data.R')
+source('~/Git/ashr/R/ash.R')
+source('~/Git/ashr/R/mix.R')
+source('~/Git/ashr/R/normalmix.R')
+source('~/Git/ashr/R/mix_opt.R')
+source('~/Git/ashr/R/output.R')
+source('~/Git/ashr/R/ashutility.R')
+
+generic.cormotif=function(betahat,sebetahat,mixcompdist = "normal",optmethod = "w_mixEM",df = NULL
+       ,nullweight = 10,pointmass = TRUE,mixsd = NULL,gridmult = sqrt(2),outputlevel = 2,g = NULL
+       ,fixg = FALSE,mode = 0,alpha = 0,grange = c(-Inf,Inf),control = list(),lik = NULL
+       ,weights=NULL,pi_thresh = 1e-10,K=1,BIC=TRUE, mess = TRUE) {
+  
+  n = dim(betahat)[1]
+  R = dim(betahat)[2]
+  
+  # set likelihood based on defaults if missing
+  if(is.null(lik)){ 
+    if(is.null(df)){
+      lik = lik_normal()
+    } else {lik = lik_t(df)}
+  }
+  
+  ## determine mixsd
+  betahat_v = as.vector(betahat)
+  sebetahat_v = as.vector(sebetahat)
+  data_v = set_data(betahat_v, sebetahat_v, lik, alpha)
+  if(is.null(mixsd)){
+    mixsd = autoselect.mixsd(data_v,gridmult,mode,grange,mixcompdist)
+  }
+  if(pointmass){
+    mixsd = c(0,mixsd)
+  }
+  L = length(mixsd)
+  null.comp = which.min(mixsd)
+  w = initpi(L,n,null.comp)
+  
+  # compute conv_dens matrix
+  matrix_llik = list()
+  for(r in 1:R){
+    data = set_data(betahat[,r], sebetahat[,r], lik, alpha)
+    if(mixcompdist == "normal") g=normalmix(w,rep(mode,L),mixsd)
+    matrix_llik[[r]] =  t(log_comp_dens_conv(g,data))
+  }
+  
+  ## fit the model via EM 
+  fitresult<-list()
+  for(i in 1:length(K)){
[…]

### 2017-09-26 paper/main.tex
change paper

--- 2017-09-16 paper/main.tex
@@
 \newtheorem{assumption}{Assumption}
+
+\def\lfdr{\textit{lfdr}}
+\def\lfsr{\textit{lfsr}}
 
@@
 where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot;\mu,\sigma^2)$ denotes the density of Normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here we assume  $\sigma_1,\sigma_2,\ldots,\sigma_L$ are fixed positive numbers forming a wide and dense grid.
+Here we assume  $\sigma_1,\sigma_2,\ldots,\sigma_L$ are known and fixed positive numbers forming a wide and dense grid, which are built via a data-driven method.
 
@@
 \subsection{Fitting the model}
+
+Our method involves three steps:
+\begin{enumerate}
+    \item For a given $K$, estimate the parameters $\pi$ and $g$ by maximizing the likelihood $L(\pi,g)$, given by (\ref{eq:4}).
+    \item Choose the best $K$ by minimizing the Bayesian Information Criterion (BIC).
+    \item Compute, for each $j$ and $r$, the posterior distribution $p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)$ by (\ref{eq:5}).
+\end{enumerate}
+
+All the details can be found in Section(\ref{sec:detail}).
+Step 1 is a convex problem, and we apply a EM algorithm to solve it.
+Step 2 is straightforward.
+The conditional distributions $p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)$ by (\ref{eq:5}) in Step 3 are analytically available, each a mixture of a point mass on zero and $L$ normal distribution.
+
+\subsection{Local False Sign Rate}
+
+To measure "significance" of an effect $\beta_{jr}$ we use the local false sign rate ($\lfsr$), which is defined as:
+\begin{equation}
+    \lfsr_{jr} := \min\{p(\beta_{jr}\geq 0|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g), p(\beta_{jr}\leq 0|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)\}
+\end{equation}
+
+Intutively, $\lfsr_{jr}$ is the probability that we would get the sign of effect $\beta_{jr}$ incorrect if we were to use our best guess of the sign.
+Therefore, a small $\lfsr$ indicates high confidence in determining the sign of an effect.
+Notice that $\lfsr$ is more conservative than the local false discovery rate ($\lfdr$), since we can infer $\lfsr_{jr}\geq\lfdr_{jr}$ from the definition.
 
@@
 
+\subsubsection{Posterior distribution}
+
+For a more general case, suppose
+\begin{equation}
+g_{kr}(\cdot;w) = \sum_{l=0}^L w_l^{kr}f_l(\cdot),
+\label{eq:geng}
+\end{equation}
+where $f_0$ is a point mass on 0. Then we have 
+\begin{equation}
+    p(\beta_{jr}|\pi,g,\hat s) = \sum_{k=1}^K\pi_k g_{kr}(\beta_{jr})=  \sum_{k=1}^K\pi_k\sum_{l=0}^Lw_l^{kr}f_l(\beta_{jr})
+    = \sum_{l=0}^L\eta_{lr}f_l(\beta_{jr}),
+\end{equation}
+where $\eta_{lr}=\sum_{k=1}^K\pi_kw_l^{kr}$. Thus, by (\ref{eq:5}), we get
+\begin{equation}
+    p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\pi, g) = \sum_{l=0}^L\theta_{ljr}p_{l}(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}).
+\end{equation}
+The posterior mixture component $p_l$ is the posterior on $\beta_{jr}$ that would be obtained using prior $f_l(\beta_{jr})$ and likelihood $p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})$.
+To compute the mixture weights $\theta_{ljr}$, denote
+\begin{equation}
+    \tilde f_{ljr}(\hat\beta_{jr}) := \int f_l(\beta_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})d\beta_{jr},
+\end{equation}
+the convolution of $f_l$ and a density function. Then $\theta_{ljr}$ can be obtained from
+\begin{equation}
+    \theta_{ljr} = \frac{\eta_{lr}\tilde f_{ljr}(\hat\beta_{jr})}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}
+\end{equation}
+
 \end{document}

### 2017-09-26 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

--- 2017-09-25 analysis/Simulation_generic_Cormotif.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,eval=TRUE)
+knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
 ```
@@
 That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 5000 units come from 4 patterns ($K=4$): 2000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 1000 unitss have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
+The 5000 units come from 4 patterns ($K=4$): 4500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 200 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 200 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 100 unitss have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
+
+### Known $K$
+
+First assume we know $K=4$ and fit the model.
 
@@
 sigma2 = c(100,100,100,100)
-pi0 = c(0.2,0.2,0.2,0.4)
+pi0 = c(0.02,0.04,0.04,0.9)
 q0 = matrix(0,nrow=K,ncol=D)
@@
 sebetahat=matrix(1,nrow=G,ncol=D)
-fit = generic.cormotif(betahat,sebetahat,K=1:5)
 ```
 
-After fitting the generic cormotif model, we can check the BIC values obtained by all cluster numbers:
+```{r,cache=TRUE}
+fit1 = generic.cormotif(betahat,sebetahat,K=4,mess=FALSE)
+```
+
+After fit the generic cormotif model, the standard deviation $\sigma_0,\sigma_1,\ldots,\sigma_L$ built by our method is
 
 ```{r}
-fit$bic
+fit1$mixsd
 ```
 
-$K = 4$ has the smallest BIC. The standard deviation $\sigma_0,\sigma_1,\ldots,\sigma_L$ built by our method is
+The estimated $\pi$ for $K = 4$ is
 
 ```{r}
-fit$mixsd
+fit1$bestmotif$pi
 ```
 
-And the estimated $\pi$ for $K = 4$ is
+Then we could check the estimation for $g$, i.e. the values of $w$, which is a $K\times L$ matrix.
+
+The estimated $w$ for studies 1 is
 
 ```{r}
-fit$bestmotif$pi
-```
-
-Then we could check the estimation for $g$, i.e. the values of $w$. The estimated $w$ for studies 1 is
-
-```{r}
-fit$bestmotif$W[[1]]
+fit1$bestmotif$W[[1]]
 ```
@@
 ```{r}
-fit$bestmotif$W[[2]]
+fit1$bestmotif$W[[2]]
 ```
@@
 ```{r}
-fit$bestmotif$W[[3]]
+fit1$bestmotif$W[[3]]
 ```
@@
 ```{r}
-fit$bestmotif$W[[4]]
+fit1$bestmotif$W[[4]]
 ```
@@
 ```{r}
-head(cbind(fit$result[[1]]$lfdr,fit$result[[2]]$lfdr,fit$result[[3]]$lfdr,fit$result[[4]]$lfdr))
+head(fit1$lfdr)
 ```
@@
 ```{r}
-head(cbind(fit$result[[1]]$lfsr,fit$result[[2]]$lfsr,fit$result[[3]]$lfsr,fit$result[[4]]$lfsr))
+head(fit1$lfsr)
+```
+
+### Unknown $K$
+
+Then suppose we don't know $k$. Set $K = 1:6$ and apply BIC and AIC.
+
+```{r,cache=TRUE}
+fit2 = generic.cormotif(betahat,sebetahat,K=1:5,mess=FALSE)
+```
+
+we can check the BIC and AIC values obtained by all cluster numbers:
+
+```{r}
+fit2$bic
+```
+
+The AIC values:
+
+```{r}
+fit2$aic
+```
+
+$K$ with the smallest BIC is
+
+```{r}
+fit2$bestmotif$K
+```
+
+We can see that $K=4$ doesn't give the smallest BIC or AIC.
+
+And the "best" estimated $\pi$ is
+
+```{r}
+fit2$bestmotif$pi
+```
+
+The estimated $w$ is
+
+```{r}
+fit2$bestmotif$W
+```
+
+Also, we could compute the lfdr:
+
+```{r}
+head(fit2$lfdr)
+```
+
+And lfsr:
+
+```{r}
+head(fit2$lfsr)
 ```

### 2017-09-26 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

@@
 
+## Discussion
+
+Our approach can conduct accurate estimation with known $K$, but without any information about $K$, neither AIC and BIC could give accurate estimation of $K$.
+
 ## Session information

### 2017-09-26 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

@@
 That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 5000 units come from 4 patterns ($K=4$): 4500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 200 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 200 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 100 unitss have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
+The 5000 units come from 4 patterns ($K=4$): 4000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 400 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 400 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 200 unitss have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
 
@@
 sigma2 = c(100,100,100,100)
-pi0 = c(0.02,0.04,0.04,0.9)
+pi0 = c(0.04,0.08,0.08,0.8)
 q0 = matrix(0,nrow=K,ncol=D)
@@
 sebetahat=matrix(1,nrow=G,ncol=D)
-```
-
-```{r,cache=TRUE}
+# fit the model
 fit1 = generic.cormotif(betahat,sebetahat,K=4,mess=FALSE)
@@
 ```{r,cache=TRUE}
-fit2 = generic.cormotif(betahat,sebetahat,K=1:5,mess=FALSE)
+fit2 = generic.cormotif(betahat,sebetahat,K=1:6,mess=FALSE)
 ```

### 2017-09-26 code/Untitled.R
change code

@@ -17,10 +17,12 @@ control = list()
 lik = NULL
 weights=NULL
 pi_thresh = 1e-10
+BIC=TRUE
+mess = TRUE
 
-n = 2000
-R = 5
-K = 3
+n = 1000
+R = 4
+K = 4
 
 #betahat = matrix(rnorm(n*R,0,sqrt(5)),ncol=R)
 betahat = matrix(0,nrow=n,ncol=R)
@@ -153,11 +155,15 @@ for(r in 1:R){
   matrix_llik[[r]] =  t(log_comp_dens_conv(g,data))
 }
 
+
+
+
+## sim1
 K = 4
 D = 4
-G = 5000
-sigma2 = c(100,100,100,100)
-pi0 = c(0.2,0.2,0.2,0.4)
+G = 1000
+sigma2 = c(16,16,16,16)
+pi0 = c(0.02,0.04,0.04,0.9)
 q0 = matrix(0,nrow=K,ncol=D)
 q0[1,] = c(1,1,1,1)
 q0[2,] = c(1,1,0,0)
@@ -179,6 +185,54 @@ for(g in 1:G){
 betahat=X
 sebetahat=matrix(1,nrow=G,ncol=D)
 
+## sim 2
+K = 4
+D = 4
+G = 5000
+sigma2 = c(16,16,16,16)
+pi0 = c(0.03,0.03,0.03,0.91)
+q0 = matrix(0,nrow=K,ncol=D)
+q0[1,] = c(1,1,1,1)
+q0[2,] = c(1,1,0,0)
+q0[3,] = c(0,0,1,1)
+q0[4,] = c(0,0,0,0)
+X = matrix(0,nrow=G,ncol=D)
+rows = rep(1:K,times=pi0*G)
+A = q0[rows,]
+set.seed(4)
+for(g in 1:G){
+  for(d in 1:D){
+    if(A[g,d]==1){
+      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
+    } else{
+      X[g,d] = rnorm(1,0,1)
+    }
+  }
+}
+betahat=X
+sebetahat=matrix(1,nrow=G,ncol=D)
 
-
-
+## sim3
+K = 5
+D = 8
+G = 10000
+sigma2 = rep(16,D)
+pi0 = c(0.02,0.02,0.02,0.02,0.92)
+q0 = matrix(0,nrow=K,ncol=D)
[…]

### 2017-09-26 code/function.R

@@ -37,13 +37,14 @@ generic.cormotif=function(betahat,sebetahat,mixcompdist = "normal",optmethod = "
   L = length(mixsd)
   null.comp = which.min(mixsd)
   w = initpi(L,n,null.comp)
+  if(mixcompdist == "normal") g=normalmix(w,rep(mode,L),mixsd)
   
   # compute conv_dens matrix
   matrix_llik = list()
+  data = list()
   for(r in 1:R){
-    data = set_data(betahat[,r], sebetahat[,r], lik, alpha)
-    if(mixcompdist == "normal") g=normalmix(w,rep(mode,L),mixsd)
-    matrix_llik[[r]] =  t(log_comp_dens_conv(g,data))
+    data[[r]] = set_data(betahat[,r], sebetahat[,r], lik, alpha)
+    matrix_llik[[r]] =  t(log_comp_dens_conv(g,data[[r]]))
   }
   
   ## fit the model via EM 
@@ -66,25 +67,25 @@ generic.cormotif=function(betahat,sebetahat,mixcompdist = "normal",optmethod = "
     bestflag=which(aic==min(aic))
   }
   bestmotif=fitresult[[bestflag]]
-  ## output
-  output = set_output(outputlevel) #sets up flags for what to output
-  resfns = set_resfns(output)
-  result = list()
-  if(length(resfns)>0){
-    for(r in 1:R){
-      data = set_data(betahat[,r], sebetahat[,r], lik, alpha)
-      what = t(bestmotif$W[[r]])%*%bestmotif$pi
-      if(mixcompdist == "normal") ghat=normalmix(what,rep(mode,L),mixsd)
-      result[[r]] = as.data.frame(lapply(resfns,do.call,list(g=prune(ghat,pi_thresh),data=data)))
-    }
-   
-  }
-  
+  ## lfdr and lsdr
+  ds = comp_lfdsr(bestmotif, g, data)
+  # output = set_output(outputlevel) #sets up flags for what to output
+  # resfns = set_resfns(output)
+  # result = list()
+  # if(length(resfns)>0){
+  #   for(r in 1:R){
+  #     data = set_data(betahat[,r], sebetahat[,r], lik, alpha)
[…]

### 2017-09-28 code/ash.R
add function

@@ -1,0 +1,795 @@
+#' @useDynLib ashr
+#' @import Matrix truncnorm SQUAREM doParallel pscl Rcpp foreach parallel
+#' @title Adaptive Shrinkage
+#' 
+#' @description Implements Empirical Bayes shrinkage and false discovery rate 
+#' methods based on unimodal prior distributions.
+#'
+#' @details The ash function provides a number of ways to perform Empirical Bayes shrinkage
+#' estimation and false discovery rate estimation. The main assumption is that 
+#' the underlying distribution of effects is unimodal. Novice users are recommended
+#' to start with the examples provided below.
+#' 
+#' In the simplest case the inputs to ash are a vector of estimates (betahat)
+#' and their corresponding standard errors (sebetahat), and degrees of freedom (df).
+#' The method assumes that for some (unknown) "true" vector of effects beta, the statistic 
+#' (betahat[j]-beta[j])/sebetahat[j] has a $t$ distribution on $df$ degrees of freedom.
+#' (The default of df=NULL assumes a normal distribution instead of a t.)
+#' 
+#' By default the method estimates the vector beta under the assumption that beta ~ g for a distribution
+#' g in G, where G is some unimodal family of distributions to be specified (see parameter \code{mixcompdist}).
+#' By default is to assume the mode is 0, and this is suitable for settings where you are interested in testing which beta[j]
+#' are non-zero. To estimate the mode see parameter \code{mode}.
+#'  
+#' As is standard in empirical Bayes methods, the fitting proceeds in two stages: 
+#' i) estimate g by maximizing a (possibly penalized) likelihood; 
+#' ii) compute the posterior distribution for each beta[j] | betahat[j],sebetahat[j] 
+#' using the estimated g as the prior distribution.
+#' 
+#' A more general case allows that beta[j]/sebetahat[j]^alpha | sebetahat[j] ~ g.
+#' 
+#' @param betahat a p vector of estimates
+#' 
+#' @param sebetahat a p vector of corresponding standard errors
+#' 
+#' @param mixcompdist distribution of components in mixture used to represent the family G.
+#' Depending on the choice of mixture component, the family G becomes more or less flexible. 
+#' Options are:\cr
+#' \describe{
+#' \item{uniform}{G is (approximately) any symmetric unimodal distribution}
[…]

### 2017-09-28 code/ashn.R

@@ -1,0 +1,10 @@
+# a wrapper function that estimates the mode, using optim
+# called by ash if mode="estimate"
+ash.estmode = function(betahat,modemin,modemax,...){
+  test.op = function(c){return(-ash(betahat=betahat,mode=c,outputlevel="loglik",...)$loglik)}
+  #opt = stats::optimize(test.op,interval=c(min(betahat),max(betahat)))
+  opt = stats::optimize(test.op,interval=c(modemin,modemax))
+  return(opt$minimum)
+}
+
+

### 2017-09-28 code/ashutility.R

@@ -1,0 +1,318 @@
+
+################################## ASH UTILITY FUNCTIONS ############################
+
+#' @title Summary method for ash object
+#'
+#' @description Print summary of fitted ash object
+#'
+#' @details \code{\link{summary}} prints the fitted mixture, the
+#'     fitted log likelihood with 10 digits and a flag to indicate
+#'     convergence
+#' @param object the fitted ash object
+#' @param ... not used, included for consistency as an S3
+#'     generic/method.
+#'
+#' @export
+#' @method summary ash
+#'
+summary.ash=function(object,...){
+  if(!is.null(object$fitted_g)){print(object$fitted_g)}
+  print("Ash object has elements: ")
+  print(paste(names(object),sep=","))
+}
+
+
+#' @title Print method for ash object
+#'
+#' @description Print the fitted distribution of beta values in the EB
+#'     hierarchical model
+#'
+#' @details None
+#' @param x the fitted ash object
+#' @param ... not used, included for consistency as an S3
+#'     generic/method.
+#'
+#' @export
+#' @method print ash
+#' 
+print.ash =function(x,...){
+  print(summary(x,...))
+}
+
+#' @title Plot method for ash object
+#'
+#' @description Plot the cdf of the underlying fitted distribution
+#'
+#' @param x the fitted ash object
+#' @param ... Arguments to be passed to methods,such as graphical parameters (see \code{\link[graphics]{plot}})
+#' @param xmin xlim lower range, default is the lowest value of betahat
+#' @param xmax xlim upper range, default is the highest value of betahat
+#' @details None
+#'
+#' @method plot ash
+#' @export
+#'
+plot.ash = function(x,...,xmin,xmax){
+  if(missing(xmin)){xmin=min(x$data$betahat)}
+  if(missing(xmax)){xmax=max(x$data$betahat)}
+  xgrid = seq(xmin,xmax,length=1000)
+  y = cdf.ash(x,xgrid)
+  graphics::plot(y,type="l",...)
+}
+
+#' @title Diagnostic plots for ash object
+#'
+#' @description Generate several plots to diagnose the fitness of ASH on the data
+#'
+#' @param x the fitted ash object
+#' @param plot.it logical. whether to plot the diagnostic result
+#' @param sebetahat.tol tolerance to test the equality of betahat
[…]

### 2017-09-28 code/igmix.R

@@ -1,0 +1,103 @@
+
+############################### METHODS FOR igmix class ###########################
+
+#' @title Constructor for igmix class
+#'
+#' @description Creates an object of class igmix (finite mixture of
+#'     univariate inverse-gammas)
+#'
+#' @details None
+#'
+#' @param pi vector of mixture proportions
+#' @param alpha vector of shape parameters
+#' @param beta vector of rate parameters
+#'
+#' @return an object of class igmix
+#'
+#' @export
+#'
+#' @examples igmix(c(0.5,0.5),c(1,1),c(1,2))
+#'
+igmix = function(pi,alpha,beta){
+  structure(data.frame(pi,alpha,beta),class="igmix")
+}
+
+comp_sd.igmix = function(m){
+  m$beta/(m$alpha-1)/sqrt(m$alpha-2)
+}
+
+#' @export
+comp_mean.igmix = function(m){
+  m$beta/(m$alpha-1)
+}
+
+comp_dens.igmix = function(m,y,log=FALSE){
+  k=ncomp(m)
+  n=length(y)
+  d = matrix(rep(y,rep(k,n)),nrow=k)
+  return(matrix(stats::dgamma(1/d, shape=m$alpha, rate=outer(m$beta,1/y^2),log),nrow=k))
+}
+
+#density of product of each component of a inverse-gamma mixture with Gamma(v/2,v/2) at s
+# s an n-vector at which density is to be evaluated
+#return a k by n matrix
+comp_dens_conv.igmix = function(m,data,FUN="+"){
+  k=ncomp(m)
+  x = data$x
+  s = data$s
+  v = data$v
+  n=length(s)
+  dens = t(exp(v/2*log(v/2)-lgamma(v/2)
+               +(v/2-1)*outer(log(s^2),rep(1,k))
+               +outer(rep(1,n),m$alpha*log(m$beta)-lgamma(m$alpha)+lgamma(m$alpha+v/2))
+               -outer(rep(1,n),m$alpha+v/2)*log(outer(v/2*s^2,m$beta,FUN="+"))))
+  return(dens)
+  
+}
+
+#' @export
+comp_cdf.igmix = function(m,y,lower.tail=TRUE){
+  vapply(y,pigamma,m$alpha,m$alpha,m$beta,lower.tail)
+}
+
+
+#' @export
[…]

### 2017-09-28 code/lik.R

@@ -1,0 +1,192 @@
+#' @title Likelihood object for normal error distribution
+#' @description Creates a likelihood object for ash for use with normal error distribution
+#' 
+#' @examples 
+#'    z = rnorm(100) + rnorm(100) # simulate some data with normal error
+#'    ash(z,1,lik=lik_normal())
+#' @export
+lik_normal = function(){
+  list(name="normal",
+       const = TRUE, #used to indicate whether the likelihood function is constant for all observations (some parts of ash only work in this case)
+       lcdfFUN = function(x){stats::pnorm(x,log=TRUE)},
+       lpdfFUN = function(x){stats::dnorm(x,log=TRUE)},
+       etruncFUN = function(a,b){my_etruncnorm(a,b)},
+       e2truncFUN = function(a,b){my_e2truncnorm(a,b)}
+       )
+}
+
+#' @title Likelihood object for t error distribution
+#' @description Creates a likelihood object for ash for use with t error distribution
+#' @param df degree of freedom parameter of t distribution
+#' 
+#' @examples 
+#'    z = rnorm(100) + rt(100,df=4) # simulate some data with t error
+#'    ash(z,1,lik=lik_t(df=4))
+#' @export
+lik_t = function(df){
+  list(name = "t",
+      const = (length(unique(df))==1),
+      lcdfFUN = function(x){stats::pt(x,df=df,log=TRUE)},
+      lpdfFUN = function(x){stats::dt(x,df=df,log=TRUE)},
+      etruncFUN = function(a,b){etrunct::e_trunct(a,b,df=df,r=1)},
+      e2truncFUN = function(a,b){etrunct::e_trunct(a,b,df=df,r=2)}
+      )
+}
+
+#' @title Likelihood object for logF error distribution
+#' @description Creates a likelihood object for ash for use with logF error distribution
+#' @param df1 first degree of freedom parameter of F distribution
+#' @param df2 second degree of freedom parameter of F distribution
+#' 
+#' @examples 
+#'    e = rnorm(100) + log(rf(100,df1=10,df2=10)) # simulate some data with log(F) error
+#'    ash(e,1,lik=lik_logF(df1=10,df2=10))
+#' @export
+lik_logF = function(df1,df2){
[…]

### 2017-09-28 code/mix.R

@@ -1,0 +1,559 @@
+
+################################## GENERIC FUNCTIONS ############################
+
+#' Generic function of calculating the component densities of the
+#'     mixture
+#' @param m mixture of k components generated by normalmix() or
+#'     unimix() or igmix()
+#' @param y is an n-vector of location
+#' @param log whether to use log-scale on densities
+#' @return A k by n matrix of densities
+#' @export
+comp_dens = function(m,y,log=FALSE){
+  UseMethod("comp_dens")
+}
+#' @export
+comp_dens.default = function(m,y,log=FALSE){
+  stop(paste("Invalid class", class(m), "for first argument in",  match.call()))
+}
+
+#' Generic function to extract the standard deviations of
+#'     components of the mixture
+#' @param m a mixture of k components generated by normalmix() or
+#'     unimix() or igmix()
+#' @return it returns a vector of standard deviations
+#' @export
+comp_sd = function(m){
+  UseMethod("comp_sd")
+}
+#' @export
+comp_sd.default = function(m){
+  stop("method comp_sd not written for this class")
+}
+
+#' Generic function to extract which components of mixture are point mass on 0
+#' @param m a mixture of k components generated by normalmix() or unimix() or igmix()
+#' @return a boolean vector indicating which components are point mass on 0 
+#' @export
+pm_on_zero = function(m){
+  UseMethod("pm_on_zero")
+}
+#' @export
+pm_on_zero.default = function(m){
+  return(comp_mean(m)==0 & comp_sd(m)==0)
+}
+
+#' Generic function of calculating the second moment of components of
+#'     the mixture
+#' @param m a mixture of k components generated by normalmix() or
+#'     unimix() or igmix()
+#' @return it returns a vector of second moments.
+comp_mean2 = function(m){
+  UseMethod("comp_mean2")
+}
+comp_mean2.default = function(m){
+  comp_sd(m)^2 + comp_mean(m)^2
+}
+
+
+#' Generic function of calculating the overall mean of the mixture
+#'
+#'
+#' @param m a mixture of k components generated by normalmix() or
+#'     unimix() or igmix()
+#' @return it returns scalar, the mean of the mixture distribution.
+calc_mixmean = function(m){
+  UseMethod("calc_mixmean")
+}
[…]

### 2017-09-28 code/mix_opt.R

@@ -1,0 +1,328 @@
+# This file contains methods for optimizing the mixture proportions given data
+# from known mixture components.
+# The penalized likelihood being maximized
+# is \sum_j \log \sum_k \pi_k f_{jk} + \sum_j (prior_j-1) \log \pi_k
+
+#' @title Estimate mixture proportions of a mixture model by Interior Point method
+#'
+#' @description Given the individual component likelihoods for a mixture model, estimates the mixture proportions.
+#'
+#' @details Optimizes \deqn{L(pi)= sum_j w_j log(sum_k pi_k f_{jk}) + h(pi)} 
+#' subject to pi_k non-negative and sum_k pi_k = 1. Here \deqn{h(pi)} is
+#' a penalty function h(pi) = sum_k (prior_k-1) log pi_k.
+#' Calls REBayes::KWDual in the REBayes package, which is in turn a wrapper to the mosek 
+#' convex optimization software. So REBayes must be installed to use this. 
+#' Used by the ash main function; there is no need for a user to call this 
+#' function separately, but it is exported for convenience.
+#'
+#' 
+#' @param matrix_lik, a n by k matrix with (j,k)th element equal to \eqn{f_k(x_j)}.
+#' @param prior, a k vector of the parameters of the Dirichlet prior on \eqn{\pi}. Recommended to be rep(1,k)
+#' @param pi_init, the initial value of \eqn{\pi} to use. If not specified defaults to (1/k,...,1/k).
+#' @param control A list of control parameters for the SQUAREM algorithm, default value is set to be control.default=list(K = 1, method=3, square=TRUE, step.min0=1, step.max0=1, mstep=4, kr=1, objfn.inc=1,tol=1.e-07, maxiter=5000, trace=FALSE). 
+#' @param weights weights to be assigned to the observations (an n vector)
+#' 
+#' @return A list, including the estimates (pihat), the log likelihood for each interation (B)
+#' and a flag to indicate convergence
+#'  
+#' @export
+mixIP = function(matrix_lik, prior, pi_init = NULL, control = list(), weights=NULL){
+ 
+  # This is the smallest value allowed for the mixture weights.
+  min.f <- 0
+    
+  if(!requireNamespace("REBayes",quietly=TRUE)){stop("mixIP requires installation of package REBayes")}
[…]

### 2017-09-28 code/normalmix.R

@@ -1,0 +1,145 @@
+############################### METHODS FOR normalmix class ###########################
+
+#' @title Constructor for normalmix class
+#'
+#' @description Creates an object of class normalmix (finite mixture
+#'     of univariate normals)
+#'
+#' @details None
+#'
+#' @param pi vector of mixture proportions
+#' @param mean vector of means
+#' @param sd vector of standard deviations
+#'
+#' @return an object of class normalmix
+#'
+#' @export
+#'
+#' @examples normalmix(c(0.5,0.5),c(0,0),c(1,2))
+#'
+normalmix = function(pi,mean,sd){
+  structure(data.frame(pi,mean,sd),class="normalmix")
+}
+
+
+#' @title comp_sd.normalmix
+#' @description returns sds of the normal mixture
+#' @param m a normal mixture distribution with k components
+#' @return a vector of length k
+#' @export
+comp_sd.normalmix = function(m){
+  m$sd
+}
+
+#' @title comp_mean.normalmix
+#' @description returns mean of the normal mixture
+#' @param m a normal mixture distribution with k components
+#' @return a vector of length k
+#' @export
+comp_mean.normalmix = function(m){
+  m$mean
+}
+
+#' @export
+comp_dens.normalmix = function(m,y,log=FALSE){
+  k=ncomp(m)
+  n=length(y)
+  d = matrix(rep(y,rep(k,n)),nrow=k)
+  return(matrix(stats::dnorm(d, m$mean, m$sd, log),nrow=k))
+}
+
+#' @export
+comp_cdf.normalmix = function(m,y,lower.tail=TRUE){
+  vapply(y,stats::pnorm,m$mean,m$mean,m$sd,lower.tail)
+}
+
+
+
+#' @title comp_dens_conv.normalmix
+#' @description returns density of convolution of each component of a
+#'     normal mixture with N(0,s^2) at x. Note that
+#'     convolution of two normals is normal, so it works that way
+#' @param m mixture distribution with k components
+#' @param data a list with components x and s to be interpreted as a normally-distributed observation and its standard error
+#' @return a k by n matrix
+comp_dens_conv.normalmix = function(m,data){
+  if(!is_normal(data$lik)){
+    stop("Error: normal mixture for non-normal likelihood is not yet implemented")
+  }
[…]

### 2017-09-28 code/output.R

@@ -1,0 +1,124 @@
+# this file contains functions to compute the columns of
+# the result data.frame from ash
+# all such functions must take parameters (g,data)
+
+# Posterior Mean
+calc_pm = function(g,data){
+  exclude = get_exclusions(data)
+  PosteriorMean = rep(0,length = n_obs(data))
+  PosteriorMean[!exclude] = postmean(g,data)[!exclude]
+  PosteriorMean[exclude] = calc_mixmean(g)
+  PosteriorMean = PosteriorMean * (data$s_orig^data$alpha)
+  return(PosteriorMean)
+}
+
+
+# Posterior Standard Deviation
+calc_psd = function(g,data){
+  exclude  = get_exclusions(data)
+  PosteriorSD = rep(0,length = n_obs(data))
+  PosteriorSD[!exclude] = postsd(g,data)[!exclude]
+  PosteriorSD[exclude] = calc_mixsd(g)
+  PosteriorSD= PosteriorSD * (data$s_orig^data$alpha)
+  return(PosteriorSD)
+}
+
+# Local FDR
+calc_lfdr = function(g,data){
+  exclude  = get_exclusions(data)
+  ZeroProb = rep(0,length = n_obs(data))
+  ZeroProb[!exclude] = colSums(comp_postprob(g,data)[pm_on_zero(g),,drop = FALSE])[!exclude]
+  ZeroProb[exclude] = sum(mixprop(g)[pm_on_zero(g)])
+  return(ZeroProb)
+}
+
+#negative probability
+calc_np = function(g,data){
+  exclude  = get_exclusions(data)
+  NegativeProb = rep(0,length = n_obs(data))
+  NegativeProb[!exclude] = cdf_post(g, 0, data)[!exclude] - calc_lfdr(g,data)[!exclude]
+  NegativeProb[exclude] = mixcdf(g,0)
+  return(NegativeProb)
+}
+
+
+#positive probability
+calc_pp = function(g,data){
+  pp=(1-calc_np(g,data)-calc_lfdr(g,data))
+  ifelse(pp<0,0,pp) #deal with numerical issues that lead to numbers <0
+}
+
+# local False Sign Rate
+calc_lfsr = function(g,data){
+  compute_lfsr(calc_np(g,data),calc_lfdr(g,data))
+}
+
+calc_svalue = function(g,data){
+  return(qval.from.lfdr(calc_lfsr(g,data)))
+}
+
[…]

### 2017-09-28 code/process_args.R

@@ -1,0 +1,73 @@
+#sets optimization method
+#also checks if necessary tools installed for optmethod specified
+set_optmethod = function(optmethod){
+  # Fallbacks for optmethod - checks required packages are installed
+  if(optmethod == "mixIP"){
+    if (!requireNamespace("REBayes", quietly = TRUE)) {  # check whether REBayes package is present
+    # If REBayes package missing
+      message("Due to absence of package REBayes, switching to EM algorithm")
+      optmethod = "cxxMixSquarem"
+    }
+  }
+  
+  if(optmethod == "cxxMixSquarem"){
+    if (!requireNamespace("Rcpp", quietly = TRUE)) {
+      message("Due to absence of Rcpp using vanilla EM; for faster performance install REBayes (preferred) or Rcpp")
+      optmethod = "mixEM"
+    }
+  }
+  
+  if (optmethod == "mixIP") assertthat::assert_that(requireNamespace("REBayes", quietly = TRUE))
+  if (optmethod == "cxxMixSquarem") assertthat::assert_that(requireNamespace("Rcpp", quietly = TRUE))
+  return(optmethod)
+}
+
+
+check_lik = function(lik, betahat, sebetahat, df, mixcompdist){
+  if(is.null(lik$lcdfFUN)){stop("Likelihood must have lcdfFUN")}
+  if(is.null(lik$lpdfFUN)){stop("Likelihood must have lpdfFUN")}
+  
+  if(!(lik$name %in% c("normal","t")) & !is.null(df)){
+    warning("Input df is ignored for this likelihood function")
+  }
+  if(!(lik$name %in% c("normal","normalmix")) & mixcompdist == "normal"){
+    stop("Error: Normal mixture for non-normal likelihood is not yet implemented")
+  }
+  if(lik$name %in% c("pois","binom")){
+    if (!sum(betahat==0) | !sum(sebetahat==1)){
+      stop("Error: betahat must be rep(0,n) and sebetaht must be 1 for 
+           Poisson/Binomial likelihood")
+    }
+  }
+  
+}
+
+
+check_args = function(mixcompdist,df,prior,optmethod,gridmult,sebetahat,betahat){
+  if(!is.numeric(betahat)){
+    stop("Error: betahat must be numeric")
+  }
+  if(!is.numeric(sebetahat)){
+    stop("Error: sebetahat must be numeric")
+  }
+  
+  if (mixcompdist == "normal" & !is.null(df))
[…]

### 2017-09-28 code/truncnorm.R

@@ -1,0 +1,154 @@
+#' @title my_etruncnorm
+#' @description Compute expectation of truncated normal.
+#'
+#' @param a Left limit of distribution.
+#' @param b Right limit of distribution.
+#' @param mean The mean of the untruncated normal.
+#' @param sd The standard deviation of the untruncated normal.
+#' @export
+my_etruncnorm= function(a,b,mean=0,sd=1){
+  alpha = (a-mean)/sd
+  beta =  (b-mean)/sd
+  #Flip the onese where both are positive, as the computations are more stable
+  #when both negative
+  flip = (alpha>0 & beta>0)
+  flip[is.na(flip)]=FALSE #deal with NAs
+  alpha[flip]= -alpha[flip]
+  beta[flip]=-beta[flip]
+  
+  #Fix a bug of quoting the truncnorm package
+  #E(X|a<X<b)=a when a==b is a natural result
+  #while etruncnorm would simply return NaN,causing PosteriorMean also NaN
+  # ZMZ:  when a and b are both negative and far from 0, etruncnorm can't compute
+  # the mean and variance. Also we should deal with 0/0 situation caused by sd = 0.
+  tmp1=etruncnorm(alpha,beta,0,1)
+  
+  isequal=is.equal(alpha,beta)
+  tmp1[isequal]=alpha[isequal]
+  
+  tmp= (-1)^flip * (mean+sd*tmp1)
+  
+  max_alphabeta = ifelse(alpha<beta, beta,alpha)
+  max_ab = ifelse(alpha<beta,b,a)
+  toobig = max_alphabeta<(-30)
+  toobig[is.na(toobig)]=FALSE
+  tmp[toobig] = max_ab[toobig]
+
+  # muzhe: this part consider many cases when 
+  # truncnorm expectation outcome is NA. For example
+  # when sd=0, or when the mean lies outside the given
+  # interval with extremely small sd, etc. This part 
+  # deals with all these problems. The concrete example
+  # can be found in test_myetruncnorm.R file.
+  # Also we need the function to be adaptive to 
+  # various forms of input: scaler, vector, matrix.
+  # To unify all these possibility we need to wrap
[…]

### 2017-09-28 code/unimix.R

@@ -1,0 +1,188 @@
+
+
+############################### METHODS FOR unimix class ###########################
+
+#' @title Constructor for unimix class
+#'
+#' @description Creates an object of class unimix (finite mixture of
+#'     univariate uniforms)
+#'
+#' @details None
+#'
+#' @param pi vector of mixture proportions
+#' @param a vector of left hand ends of uniforms
+#' @param b vector of right hand ends of uniforms
+#'
+#' @return an object of class unimix
+#'
+#' @export
+#'
+#' @examples unimix(c(0.5,0.5),c(0,0),c(1,2))
+unimix = function(pi,a,b){
+  structure(data.frame(pi,a,b),class="unimix")
+}
+
+#' @export
+comp_cdf.unimix = function(m,y,lower.tail=TRUE){
+  vapply(y,stats::punif,m$a,min=m$a,max=m$b,lower.tail)
+}
+
+comp_sd.unimix = function(m){
+  (m$b-m$a)/sqrt(12)
+}
+
+#' @export
+comp_mean.unimix = function(m){
+  (m$a+m$b)/2
+}
+
+
+
+comp_dens.unimix = function(m,y,log=FALSE){
+  k=ncomp(m)
+  n=length(y)
+  d = matrix(rep(y,rep(k,n)),nrow=k)
+  return(matrix(stats::dunif(d, m$a, m$b, log),nrow=k))
+}
+
+#' density of convolution of each component of a unif mixture 
+#' @param m a mixture of class unimix
+#' @param data, see set_data()
+#'
+#' @return a k by n matrix
+#'
+#' @export
+comp_dens_conv.unimix = function(m,data){
+  return(exp(log_comp_dens_conv(m,data)))
+}
+
+#' log density of convolution of each component of a unif mixture 
+#' @inheritParams comp_dens_conv.unimix
+#' @return a k by n matrix of densities
+log_comp_dens_conv.unimix = function(m,data){
+  b = pmax(m$b,m$a) #ensure a<b
+  a = pmin(m$b,m$a)
+  lik = data$lik
+  
+  lpa = do.call(lik$lcdfFUN, list(outer(data$x,a,FUN="-")/data$s))
+  lpb = do.call(lik$lcdfFUN, list(outer(data$x,b,FUN="-")/data$s))
+  
+  if (sum(lpa-lpb,na.rm=TRUE)<0){
[…]

### 2017-09-28 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

@@
 
-### Known $K$
+### Known K
 
@@
 ```{r,cache=TRUE}
-source('~/Git/cormotif/code/function.R')
+source('../code/function.R')
 K = 4
@@
 
-### Unknown $K$
+### Unknown K
 

### 2017-09-28 analysis/Simulation_generic_Cormotif.Rmd
change function

@@
 That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 5000 units come from 4 patterns ($K=4$): 4000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 400 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 400 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 200 unitss have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.
+The 5000 units come from 4 patterns ($K=4$): 4000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 400 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 400 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 200 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.
 
@@
 G = 5000
-sigma2 = c(100,100,100,100)
+sigma2 = c(16,16,16,16)
 pi0 = c(0.04,0.08,0.08,0.8)

### 2017-09-28 code/function.R

@@ -1,13 +1,13 @@
 library("SQUAREM")
 
-source('~/Git/ashr/R/lik.R')
-source('~/Git/ashr/R/set_data.R')
-source('~/Git/ashr/R/ash.R')
-source('~/Git/ashr/R/mix.R')
-source('~/Git/ashr/R/normalmix.R')
-source('~/Git/ashr/R/mix_opt.R')
-source('~/Git/ashr/R/output.R')
-source('~/Git/ashr/R/ashutility.R')
+source('../code/lik.R')
+source('../code/set_data.R')
+source('../code/ash.R')
+source('../code/mix.R')
+source('../code/normalmix.R')
+source('../code/mix_opt.R')
+source('../code/output.R')
+source('../code/ashutility.R')
 
 generic.cormotif=function(betahat,sebetahat,mixcompdist = "normal",optmethod = "w_mixEM",df = NULL
        ,nullweight = 10,pointmass = TRUE,mixsd = NULL,gridmult = sqrt(2),outputlevel = 2,g = NULL

### 2017-09-28 analysis/Simulation_generic_Cormotif.Rmd
show loglike

@@
 title: "Simulation for generic Cormotif model"
-author: "First Last"
+author: "Zhiwei Ma"
 date: "9/24/2017"
@@
 
-Then suppose we don't know $k$. Set $K = 1:6$ and apply BIC and AIC.
+Then suppose we don't know $k$. Set $K = 1:10$ and apply BIC and AIC.
 
 ```{r,cache=TRUE}
-fit2 = generic.cormotif(betahat,sebetahat,K=1:6,mess=FALSE)
+fit2 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
 ```
@@
 fit2$bestmotif$pi
+```
+
+The loglikelihood values:
+
+```{r}
+fit2$loglike
 ```

### 2017-09-28 analysis/Trend_of_loglike.Rmd
wflow_publish("/home/zwma/git/cormotif/analysis/Trend_of_loglike.Rmd")

---
title: "Trend of Loglike"
author: "Zhiwei Ma"
date: "9/24/2017"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=TRUE)
```

## Simulation 1

We set unit number $n = 5000$, and study number $R = 4$. We set every observation has the same standard error, $s_{jr}=1$. That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$. The 5000 units come from 4 patterns ($K=4$): 2500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 500 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.

To fit the model, set $K = 1:10$.

```{r,cache=TRUE}
source('../code/function.R')
# Simulate data
K = 4
D = 4
G = 5000
sigma2 = c(16,16,16,16)
pi0 = c(0.1,0.2,0.2,0.5)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,1,1,0)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(111)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
betahat=X
sebetahat=matrix(1,nrow=G,ncol=D)
```

```{r,cache=TRUE}
#fit the model
fit1 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The estimated $\pi$ for $K=4$

```{r}
fit1$allmotif[[4]]$pi
```

we can check the BIC and AIC values obtained by all cluster numbers:

```{r}
fit1$bic
```

The AIC values:

```{r}
fit1$aic
```

The loglikelihood values:

```{r}
fit1$loglike
```

And plot

```{r}
plot(fit1$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
```

## Simulation 2

We set unit number $n = 5000$, and study number $R = 4$. We set every observation has the same standard error, $s_{jr}=1$. That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$. The 5000 units come from 4 patterns ($K=4$): 2500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3, 4$ and $\beta_{jr}=0$ for $r = 1, 2$; 500 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.

To fit the model, set $K = 1:10$.

```{r,cache=TRUE}
#source('../code/function.R')
# Simulate data
K = 4
D = 4
G = 5000
sigma2 = c(16,16,16,16)
pi0 = c(0.1,0.2,0.2,0.5)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,0,1,1)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(222)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
betahat=X
sebetahat=matrix(1,nrow=G,ncol=D)
```

```{r,cache=TRUE}
#fit the model
fit2 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The estimated $\pi$ for $K=4$

```{r}
fit2$allmotif[[4]]$pi
```

we can check the BIC and AIC values obtained by all cluster numbers:

```{r}
fit2$bic
```

The AIC values:

```{r}
fit2$aic
```

The loglikelihood values:

```{r}
fit2$loglike
```

And plot

```{r}
plot(fit2$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-09-29 analysis/Trend_of_loglike.Rmd
wflow_publish("Trend_of_loglike.Rmd")

@@
 That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 5000 units come from 4 patterns ($K=4$): 2500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 500 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.
+The 5000 units come from 4 patterns ($K=4$): 2500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3, 4$ and $\beta_{jr}=0$ for $r = 1, 2$; 500 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.
 
@@
 q0[2,] = c(1,1,0,0)
-q0[3,] = c(0,1,1,0)
+q0[3,] = c(0,0,1,1)
 q0[4,] = c(0,0,0,0)
@@
 A = q0[rows,]
-set.seed(111)
+set.seed(222)
 for(g in 1:G){
@@
 
-We set unit number $n = 5000$, and study number $R = 4$.
-We set every observation has the same standard error, $s_{jr}=1$.
-That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 5000 units come from 4 patterns ($K=4$): 2500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3, 4$ and $\beta_{jr}=0$ for $r = 1, 2$; 500 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.
+With the same setting as simulation 1, just change $\beta_{jr}\sim N(0,10^2)$, we check the same process.
 
@@
 G = 5000
-sigma2 = c(16,16,16,16)
+sigma2 = c(100,100,100,100)
 pi0 = c(0.1,0.2,0.2,0.5)
@@
 A = q0[rows,]
-set.seed(222)
+set.seed(22)
 for(g in 1:G){
@@
 
+## Simulation 3
+
+With the same setting as simulation 1, just change the proportion of 4 patterns to 1250:1250:1250:1250, we check the same process.
+
+To fit the model, set $K = 1:10$.
+
+```{r,cache=TRUE}
+#source('../code/function.R')
+# Simulate data
+K = 4
+D = 4
+G = 5000
+sigma2 = c(16,16,16,16)
+pi0 = c(0.25,0.25,0.25,0.25)
+q0 = matrix(0,nrow=K,ncol=D)
+q0[1,] = c(1,1,1,1)
+q0[2,] = c(1,1,0,0)
+q0[3,] = c(0,0,1,1)
+q0[4,] = c(0,0,0,0)
+X = matrix(0,nrow=G,ncol=D)
+rows = rep(1:K,times=pi0*G)
+A = q0[rows,]
+set.seed(333)
+for(g in 1:G){
+  for(d in 1:D){
+    if(A[g,d]==1){
+      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
+    } else{
+      X[g,d] = rnorm(1,0,1)
+    }
+  }
+}
+betahat=X
+sebetahat=matrix(1,nrow=G,ncol=D)
+```
+
+```{r,cache=TRUE}
+#fit the model
+fit3 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
+```
+
+The estimated $\pi$ for $K=4$
+
+```{r}
+fit3$allmotif[[4]]$pi
+```
+
+we can check the BIC and AIC values obtained by all cluster numbers:
+
+```{r}
+fit3$bic
+```
+
+The AIC values:
+
+```{r}
+fit3$aic
+```
+
+The loglikelihood values:
+
+```{r}
+fit3$loglike
+```
+
+And plot
+
+```{r}
+plot(fit3$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
+```
+
 ## Session information

### 2017-09-29 analysis/Trend_of_loglike.Rmd
wflow_publish("Trend_of_loglike.Rmd")
wflow_publish("/home/zwma/git/cormotif/analysis/Trend_of_loglike.Rmd")

@@
 source('../code/function.R')
+source('../code/lik.R')
+source('../code/set_data.R')
+source('../code/ash.R')
+source('../code/mix.R')
+source('../code/normalmix.R')
+source('../code/mix_opt.R')
+source('../code/output.R')
+source('../code/ashutility.R')
 # Simulate data

### 2017-09-29 code/function.R
function

@@ -124,7 +124,7 @@ gcmfit = function(matrix_llik,K=1,max.iter=300,tol = 1e-4,w_init,pi_thresh,mess=
     tempmax = apply(matrix_llik2,1,max)
     matrix_lik2 = exp(matrix_llik2-tempmax)
     fit.p=mixEM(matrix_lik=matrix_lik2, prior=1,pi_init = p)
-    p.new = fit.p$pihat
+    p.new = normalize(fit.p$pihat + 1e-8)
     #print(p.new)
     
     ## update W

### 2017-09-29 analysis/Trend_of_loglike.Rmd
wflow_publish("Trend_of_loglike.Rmd")

--- 2017-09-29 analysis/Trend_of_loglike.Rmd
@@
 ```{r,cache=TRUE}
+
 source('../code/function.R')
+source('../code/lik.R')
+source('../code/set_data.R')
+source('../code/ash.R')
+source('../code/mix.R')
+source('../code/normalmix.R')
+source('../code/mix_opt.R')
+source('../code/output.R')
+source('../code/ashutility.R')
+
 # Simulate data
@@
 ```{r}
-fit1$bic
+plot(fit1$bic,type = "l",xlab = "K",ylab = "BIC")
 ```
@@
 ```{r}
-fit1$aic
+plot(fit1$aic,type = "l",xlab = "K",ylab = "AIC")
 ```
@@
 sebetahat=matrix(1,nrow=G,ncol=D)
-```
-
-```{r,cache=TRUE}
 #fit the model
@@
 ```{r}
-fit2$bic
+plot(fit2$bic,type = "l",xlab = "K",ylab = "BIC")
 ```
@@
 ```{r}
-fit2$aic
+plot(fit2$aic,type = "l",xlab = "K",ylab = "AIC")
 ```
@@
 ```{r}
-fit3$bic
+plot(fit3$bic,type = "l",xlab = "K",ylab = "BIC")
 ```
@@
 ```{r}
-fit3$aic
+plot(fit3$aic,type = "l",xlab = "K",ylab = "AIC")
 ```

### 2017-09-29 analysis/Trend_of_loglike.Rmd
wflow_publish("Trend_of_loglike.Rmd")

@@
 ```{r,cache=TRUE}
-#source('../code/function.R')
+source('../code/function.R')
 # Simulate data

### 2017-10-03 analysis/Trend_of_loglike.Rmd
wflow_publish("/home/zwma/git/cormotif/analysis/Trend_of_loglike.Rmd")

@@
 ```{r,cache=TRUE}
-
 source('../code/function.R')
-source('../code/lik.R')
-source('../code/set_data.R')
-source('../code/ash.R')
-source('../code/mix.R')
-source('../code/normalmix.R')
-source('../code/mix_opt.R')
-source('../code/output.R')
-source('../code/ashutility.R')
 
@@
 sebetahat=matrix(1,nrow=G,ncol=D)
-```
 
-```{r,cache=TRUE}
 #fit the model

### 2017-10-03 analysis/Trend_of_loglike.Rmd
wflow_publish("Trend_of_loglike.Rmd")

@@
 
+```{r}
+source('../code/lik.R')
+source('../code/set_data.R')
+source('../code/ash.R')
+source('../code/mix.R')
+source('../code/normalmix.R')
+source('../code/mix_opt.R')
+source('../code/output.R')
+source('../code/ashutility.R')
+```
+
 ```{r,cache=TRUE}
@@
 fit1$allmotif[[4]]$pi
+```
+
+The estimated $\pi$ for $K=1：10$
+
+```{r}
+fit1$allmotif[[1]]$pi
+fit1$allmotif[[2]]$pi
+fit1$allmotif[[3]]$pi
+fit1$allmotif[[4]]$pi
+fit1$allmotif[[5]]$pi
+fit1$allmotif[[6]]$pi
+fit1$allmotif[[7]]$pi
+fit1$allmotif[[8]]$pi
+fit1$allmotif[[9]]$pi
+fit1$allmotif[[10]]$pi
 ```
@@
 fit2$allmotif[[4]]$pi
+```
+
+The estimated $\pi$ for $K=1：10$
+
+```{r}
+fit2$allmotif[[1]]$pi
+fit2$allmotif[[2]]$pi
+fit2$allmotif[[3]]$pi
+fit2$allmotif[[4]]$pi
+fit2$allmotif[[5]]$pi
+fit2$allmotif[[6]]$pi
+fit2$allmotif[[7]]$pi
+fit2$allmotif[[8]]$pi
+fit2$allmotif[[9]]$pi
+fit2$allmotif[[10]]$pi
 ```
@@
 
+The estimated $\pi$ for $K=1：10$
+
+```{r}
+fit3$allmotif[[1]]$pi
+fit3$allmotif[[2]]$pi
+fit3$allmotif[[3]]$pi
+fit3$allmotif[[4]]$pi
+fit3$allmotif[[5]]$pi
+fit3$allmotif[[6]]$pi
+fit3$allmotif[[7]]$pi
+fit3$allmotif[[8]]$pi
+fit3$allmotif[[9]]$pi
+fit3$allmotif[[10]]$pi
+```
+
 we can check the BIC and AIC values obtained by all cluster numbers:

### 2017-10-03 analysis/Trend_of_loglike.Rmd
wflow_publish("/home/zwma/git/cormotif/analysis/Trend_of_loglike.Rmd")

@@
 To fit the model, set $K = 1:10$.
-
-```{r}
-source('../code/lik.R')
-source('../code/set_data.R')
-source('../code/ash.R')
-source('../code/mix.R')
-source('../code/normalmix.R')
-source('../code/mix_opt.R')
-source('../code/output.R')
-source('../code/ashutility.R')
-```
 

### 2017-10-08 analysis/estimation_of_effect_size.Rmd
wflow_publish("estimation_of_effect_size.Rmd")

---
title: "estimation of effect sizes and detection of significant effects"
author: "Zhiwei Ma"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
```

## Overview

In this files, we analyze the accuracy of estimaiton of effect sizes and the detection of significant effects. To measure the accuracy, we apply the relative root mean squrared error (RRMSE). An RRMSE < 1 indicates the methods produces estimates that are more accurate than the original obserbation $\hat\beta$. For detection of significant effects, we require the estimated sign of each significant effect to be correct to be considered a "true positive".

## Results

### Simulation 1

We set unit number $n = 5000$, and study number $R = 4$. We set every observation has the same standard error, $s_{jr}=1$. That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$. The 5000 units come from 4 patterns ($K=4$): 4500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 200 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 200 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3, 4$ and $\beta_{jr}=0$ for $r = 1, 2$; 100 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.

To fit the model, set $K = 1:10$.

```{r,cache=TRUE}
source('../code/function.R')
K = 4
D = 4
G = 5000
sigma2 = c(16,16,16,16)
pi0 = c(0.02,0.04,0.04,0.9)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,1,1,0)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
Y = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(111)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(sigma2[d]))
    } else{
      X[g,d] = 0
    }
  }
}
beta = X
sebetahat=matrix(1,nrow=G,ncol=D)
betahat=matrix(10,nrow=G,ncol=D)
for(g in 1:G){
  for(d in 1:D){
    betahat[g,d] = rnorm(1,beta[g,d],sebetahat[g,d])
  }
}

# fit the model
fit1 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The RRMSE for K = 1:10 is

```{r,echo = FALSE}
RRMSE = rep(0,10)
for(i in 1:10){
  RRMSE[i] = sqrt(mean((fit1$post_mean[[i]]-beta)^2)/mean((betahat-beta)^2))
}
plot(RRMSE,type="l",xlab="K",ylab="RRMSE",main="RRMSE for Simulation 1")
```

Changing the threshold for lfsr, we plot the ROC curve for $K=1,2,4,10$. Notice that when $K=1$, it is equvalent to apply ash to each study separately, and $K = 4$ is the true model.

```{r,echo = FALSE}
L = 10000
thres = seq(0.01,1,length=L)
fpr = rep(0,L)
tpr = rep(0,L)
for(i in 1:L){
  tpr[i] = mean(fit1$lfsr[[1]][beta != 0]<thres[i])
  fpr[i] = mean(fit1$lfsr[[1]][beta == 0]<thres[i])
}
plot(fpr,tpr,type="l",xlab="False Positive Rate",ylab="True Positive Rate",main ="ROC curve for Simulation 1",col=1)
for(i in 1:L){
  tpr[i] = mean(fit1$lfsr[[2]][beta != 0]<thres[i])
  fpr[i] = mean(fit1$lfsr[[2]][beta == 0]<thres[i])
}
lines(fpr,tpr,col=2)
for(i in 1:L){
  tpr[i] = mean(fit1$lfsr[[4]][beta != 0]<thres[i])
  fpr[i] = mean(fit1$lfsr[[4]][beta == 0]<thres[i])
}
lines(fpr,tpr,col=3)
for(i in 1:L){
  tpr[i] = mean(fit1$lfsr[[10]][beta != 0]<thres[i])
  fpr[i] = mean(fit1$lfsr[[10]][beta == 0]<thres[i])
}
lines(fpr,tpr,col=4)
legend('bottomright',legend=c("K=1","K=2","K=4","K=10"),lty =1,col=1:4)
```

### Simulation 2

We set unit number $n = 5000$, and study number $R = 8$. We set every observation has the same standard error, $s_{jr}=1$. That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$. The 5000 units come from 5 patterns ($K=5$): 4000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4,5,6,7,8$; 250 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$ and $\beta_{jr}=0$ for $r =5, 6,7,8$; 250 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =5, 6,7,8$ and $\beta_{jr}=0$ for $r = 1,2,3,4$; 250 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3,4,5,6$ and $\beta_{jr}=0$ for $r = 1,2,7,8$; 250 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4,5,6,7,8$.

To fit the model, set $K = 1:10$.

```{r,cache=TRUE}
source('../code/function.R')
K = 5
D = 8
G = 5000
sigma2 = rep(16,D)
pi0 = c(0.05,0.05,0.05,0.05,0.8)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1,1,1,1,1)
q0[2,] = c(1,1,1,1,0,0,0,0)
q0[3,] = c(0,0,0,0,1,1,1,1)
q0[4,] = c(0,0,1,1,1,1,0,0)
q0[5,] = c(0,0,0,0,0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
Y = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(222)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(sigma2[d]))
    } else{
      X[g,d] = 0
    }
  }
}
beta = X
sebetahat=matrix(1,nrow=G,ncol=D)
betahat=matrix(10,nrow=G,ncol=D)
for(g in 1:G){
  for(d in 1:D){
    betahat[g,d] = rnorm(1,beta[g,d],sebetahat[g,d])
  }
}
# fit the model
fit2 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The RRMSE for K = 1:10 is

```{r,echo = FALSE}
RRMSE = rep(0,10)
for(i in 1:10){
  RRMSE[i] = sqrt(mean((fit2$post_mean[[i]]-beta)^2)/mean((betahat-beta)^2))
}
plot(RRMSE,type="l",xlab="K",ylab="RRMSE",main="RRMSE for Simulation 2")
```

We plot the ROC curve for $K=1,2,5,10$. Notice that when $K=1$, it is equvalent to apply ash to each study separately, and $K=5$ is the true model.

```{r,echo = FALSE}
L = 10000
thres = seq(0.01,1,length=L)
fpr = rep(0,L)
tpr = rep(0,L)
for(i in 1:L){
  tpr[i] = mean(fit2$lfsr[[1]][beta != 0]<thres[i])
  fpr[i] = mean(fit2$lfsr[[1]][beta == 0]<thres[i])
}
plot(fpr,tpr,type="l",xlab="False Positive Rate",ylab="True Positive Rate",main="ROC curve for Simulation 2",col=1)
for(i in 1:L){
  tpr[i] = mean(fit2$lfsr[[2]][beta != 0]<thres[i])
  fpr[i] = mean(fit2$lfsr[[2]][beta == 0]<thres[i])
}
lines(fpr,tpr,col=2)
for(i in 1:L){
  tpr[i] = mean(fit2$lfsr[[5]][beta != 0]<thres[i])
  fpr[i] = mean(fit2$lfsr[[5]][beta == 0]<thres[i])
}
lines(fpr,tpr,col=3)
for(i in 1:L){
  tpr[i] = mean(fit2$lfsr[[10]][beta != 0]<thres[i])
  fpr[i] = mean(fit2$lfsr[[10]][beta == 0]<thres[i])
}
lines(fpr,tpr,col=4)
legend('bottomright',legend=c("K=1","K=2","K=4","K=10"),lty =1,col=1:4)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-10-08 analysis/estimation_of_effect_size.Rmd
wflow_publish("estimation_of_effect_size.Rmd")

@@
 author: "Zhiwei Ma"
-date: YYYY-MM-DD
+date: 2017-10-08
 output: html_document
@@
 
-Changing the threshold for lfsr, we plot the ROC curve for $K=1,2,4,10$. Notice that when $K=1$, it is equvalent to apply ash to each study separately, and $K = 4$ is the true model.
+Changing the threshold for lfsr, we plot the ROC curve for $K=1,2,4,10$. Notice that when $K=1$, it is equvalent to applying ash to each study separately, and $K = 4$ is the true model.
 

### 2017-10-08 analysis/estimation_of_effect_size.Rmd
wflow_publish("estimation_of_effect_size.Rmd")

@@
 
-Changing the threshold for lfsr, we plot the ROC curve for $K=1,2,4,10$. Notice that when $K=1$, it is equvalent to applying ash to each study separately, and $K = 4$ is the true model.
+Notice that when $K=1$, it is equvalent to applying ash to each study separately, and $K = 4$ is the true model.
+
+Changing the threshold for lfsr, we plot the ROC curve for $K=1,2,4,10$.
 
@@
 
-We plot the ROC curve for $K=1,2,5,10$. Notice that when $K=1$, it is equvalent to apply ash to each study separately, and $K=5$ is the true model.
+Notice that when $K=1$, it is equvalent to applying ash to each study separately, and $K=5$ is the true model.
+
+We plot the ROC curve for $K=1,2,5,10$.
 
@@
 lines(fpr,tpr,col=4)
-legend('bottomright',legend=c("K=1","K=2","K=4","K=10"),lty =1,col=1:4)
+legend('bottomright',legend=c("K=1","K=2","K=5","K=10"),lty =1,col=1:4)
 ```

### 2017-10-08 code/Untitled.R
change function

@@ -170,20 +170,29 @@ q0[2,] = c(1,1,0,0)
 q0[3,] = c(0,1,1,0)
 q0[4,] = c(0,0,0,0)
 X = matrix(0,nrow=G,ncol=D)
+Y = matrix(0,nrow=G,ncol=D)
 rows = rep(1:K,times=pi0*G)
 A = q0[rows,]
 set.seed(4)
 for(g in 1:G){
   for(d in 1:D){
     if(A[g,d]==1){
-      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
+      X[g,d] = rnorm(1,0,sqrt(sigma2[d]))
     } else{
-      X[g,d] = rnorm(1,0,1)
+      X[g,d] = 0
     }
   }
 }
-betahat=X
+beta = X
 sebetahat=matrix(1,nrow=G,ncol=D)
+betahat=matrix(10,nrow=G,ncol=D)
+for(g in 1:G){
+  for(d in 1:D){
+    betahat[g,d] = rnorm(1,beta[g,d],sebetahat[g,d])
+  }
+}
+
+
 
 ## sim 2
 K = 4
@@ -215,9 +224,9 @@ sebetahat=matrix(1,nrow=G,ncol=D)
 ## sim3
 K = 5
 D = 8
-G = 10000
+G = 1000
 sigma2 = rep(16,D)
-pi0 = c(0.02,0.02,0.02,0.02,0.92)
+pi0 = c(0.05,0.05,0.05,0.05,0.8)
 q0 = matrix(0,nrow=K,ncol=D)
 q0[1,] = c(1,1,1,1,1,1,1,1)
 q0[2,] = c(1,1,1,1,0,0,0,0)
@@ -236,3 +245,32 @@ for(g in 1:G){
     }
   }
 }
+
+L = 1000
+thres = seq(0.01,1,length=L)
+fpr = rep(0,L)
+tpr = rep(0,L)
+for(i in 1:L){
+  tpr[i] = mean(results$lfsr[[1]][beta != 0]<thres[i])
+  fpr[i] = mean(results$lfsr[[1]][beta == 0]<thres[i])
+}
+plot(fpr,tpr,type="l",xlab="False Positive Rate",ylab="True Positive Rate",col=1)
+for(i in 1:L){
+  tpr[i] = mean(results$lfsr[[2]][beta != 0]<thres[i])
+  fpr[i] = mean(results$lfsr[[2]][beta == 0]<thres[i])
+}
[…]

### 2017-10-08 code/function.R

@@ -68,7 +68,16 @@ generic.cormotif=function(betahat,sebetahat,mixcompdist = "normal",optmethod = "
   }
   bestmotif=fitresult[[bestflag]]
   ## lfdr and lsdr
-  ds = comp_lfdsr(bestmotif, g, data)
+  lfdr = list()
+  lfsr = list()
+  post_mean = list()
+  for(i in 1:length(K)){
+    ds = comp_lfdsr(fitresult[[i]], g, data)
+    lfdr[[i]] = ds$lfdr
+    lfsr[[i]] = ds$lfsr
+    post_mean[[i]] = ds$post_mean
+  }
+  #ds = comp_lfdsr(bestmotif, g, data)
   # output = set_output(outputlevel) #sets up flags for what to output
   # resfns = set_resfns(output)
   # result = list()
@@ -82,7 +91,7 @@ generic.cormotif=function(betahat,sebetahat,mixcompdist = "normal",optmethod = "
   #  
   # }
   return(list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
-                aic=cbind(K,aic),loglike=cbind(K,loglike),mixsd=mixsd,allmotif=fitresult,lfdr=ds$lfdr,lfsr=ds$lfsr))
+                aic=cbind(K,aic),loglike=cbind(K,loglike),mixsd=mixsd,allmotif=fitresult,lfdr=lfdr,lfsr=lfsr,post_mean=post_mean))
 }
 
 gcmfit = function(matrix_llik,K=1,max.iter=300,tol = 1e-4,w_init,pi_thresh,mess=TRUE){
@@ -188,6 +197,7 @@ comp_lfdsr= function(bestmotif, g, data){
   clustlike0 = bestmotif$clustlike
   lfdr = matrix(0,n,R)
   lfsr = matrix(0,n,R)
+  post_mean = matrix(0,n,R)
   Theta = array(0,dim=c(R,n,L))
   for(r in 1:R){
     we = array(0,dim=c(K0,n,L))
@@ -202,11 +212,12 @@ comp_lfdsr= function(bestmotif, g, data){
     lfdr[,r] = Theta[r,,1]
     NegativeProb = rowSums(t(comp_cdf_post(g,0,data[[r]]))*Theta[r,,])-lfdr[,r]
     lfsr[,r] = compute_lfsr(NegativeProb,lfdr[,r])
[…]

### 2017-10-17 paper/CorMotif.tex
update paper

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage{setspace}
\doublespacing
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{amsmath,amsfonts}
\usepackage{bm}

\newtheorem{assumption}{Assumption}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

\title{Solving Correlation Motifs via EM algorithm}
\author{Zhiwei Ma}
						% Activate to display a given date or no date

\begin{document}
\maketitle
\section{Introduction}

This article contains introduction about CorMotif model by $Wei$ $and$ $Li$ (2015) and detail implementation of EM algorithm in CorMotif. The original CorMotif is based on $limma$ ($Smyth$, 2004). In this article, we expand CorMotif to any distribution. For Gaussian model, we introduce the iterative formula to estimate its parameters.

\section{Correlation Motif Model}

Suppose there are $n$ genes and $R$ studies for each gene. Let $x_{ir}$ denote the statistics of gene $i$ in study $r$, where $i=1,2,\ldots,n$ and $r = 1,2,\ldots,R$. The collection of all observed data is 
$$X=\{x_{ir}| i=1,2,\ldots,n; r = 1,2,\ldots,R\}.$$

For each study, the statistics $x_{ir}$ may follow two different distribution: one from the $null$ hypothesis, the other from the alternative hypothesis. Let $y_{ir}=1$ or $0$ denote whether $x_{ir}$ comes from the $alternative$ hypothesis. Given the state $y_{ir}$, assume $x_{ir}$ follows
\begin{eqnarray}
\label{eq:1}
f_{r0}(x_{ir}):=f_r(x_{ir}|y_{ir}=0),    \\ 
f_{r1}(x_{ir}):=f_r(x_{ir}|y_{ir}=1).
\label{eq:2}
\end{eqnarray}

The states of gene $i$ can be expressed as $Y_i = (y_{i1},\ldots,y_{iR})^T$. For $R$ studies, there are totally $2^R$ possible configurations. One way to study the correlation between different studies is to find the frequencies of each configuration among $n$ genes. However, we need at least $O(2^R)$ samples to calculate each frequency, which will increase rapidly as $R$ increases. Here we introduce the Correlation Motifs.

CorMotif adopts a hierarchical mixture model. It assumes that all genes fall into $K$ classes. In addtion, it assumes

\begin{assumption} Each gene $i$ is assigned to a class label $z_i$, here $z_i\in \{1,2,\ldots,K\}$. The prior distribution for $z_i$ is $Pr(z_i=k)=\pi_k, k=1,2,\ldots,K $. Thus, we have $\sum_k \pi_k=1$ and denote $ \pi=(\pi_1,\ldots,\pi_K)$.
\end{assumption}
\begin{assumption} Given gene's class label $z_i$, gene's state $y_{ir}$ are independent, following $ Pr(y_{ir}=1|z_i =k)=q_{kr}$. For the $k^{th}$ class, denote $q_k=(q_{k1},\ldots,q_{kR})^T$. 
\end{assumption}
\begin{assumption}Given the gene's state $y_{ir}$, the statistic $x_{ir}$ are independently following (\ref{eq:1}) and (\ref{eq:2}).
\end{assumption}

Let $Z = (z_1,\ldots,z_n)$ denote the class membership and $ Q = (q_1,\ldots,q_K)^T$ is the $K \times R$ matrix. For gene $i$ and study $r$ we have 
\begin{equation*}
p(x_{ir},y_{ir}|z_i=k,\pi,Q) = [q_{kr}f_{r1}(x_{ir})]^{y_{ir}}[(1-q_{kr})f_{r0}(x_{ir})]^{1-y_{ir}}.
\end{equation*}
Thus,
\begin{equation*}
p(X_{i},Y_{i}|z_i=k,\pi,Q) = \prod_{r=1}^R [q_{kr}f_{r1}(x_{ir})]^{y_{ir}}[(1-q_{kr})f_{r0}(x_{ir})]^{1-y_{ir}},
\end{equation*}
here $X_{i}=(x_{i1},\ldots, x_{iR})^T$. We can get
\begin{equation*}
p(X_{i},Y_{i}, z_i|\pi,Q) = \prod_{k=1}^K \{\pi_k \prod_{r=1}^R [q_{kr}f_{r1}(x_{ir})]^{y_{ir}}[(1-q_{kr})f_{r0}(x_{ir})]^{1-y_{ir}}\}^{\mathbb {I}(z_i=k)},
\end{equation*}
where $\mathbb {I}$  is an indicator function. Therefore, based on above formulas, the joint probability distribution of $X$, $Y$ and $Z$ condtional on $\pi$ and $Q$ is
\begin{equation}
p(X,Y, Z|\pi,Q) =\prod_{i=1}^n \prod_{k=1}^K \{\pi_k \prod_{r=1}^R [q_{kr}f_{r1}(x_{ir})]^{y_{ir}}[(1-q_{kr})f_{r0}(x_{ir})]^{1-y_{ir}}\}^{\mathbb {I}(z_i=k)}.
\end{equation}
In the joint probability distribution above, only $X$ is observed, $Y$ and $Z$ are missing values or $latent$ data. $\pi$ and $Q$ are unknown parameters. The maximum likelihood estimate (MLE) of the unknown parameters is determined by the marginal likelihood of the observed data
\begin{eqnarray}
L(\pi,Q;X) &= &p(X|\pi,Q) = \sum_Y \sum_Z p(X,Y, Z|\pi,Q)  \nonumber \\
&=& \prod_{i=1}^n\sum_{k=1}^K \pi_k\prod_{r=1}^R [q_{kr}f_{r1}(x_{ir})+(1-q_{kr})f_{r0}(x_{ir})].
\end{eqnarray}
However, it is unrealistic to obtain the optimal values for $\pi$ and $Q$ by maximizing above formula directly. Instead, we will apply the EM algorithm to handle this problem.

\section{Estimation Method}
\subsection{Expectation-Maximization algorithm}

The Expectation-Maximization (EM) algorithm seeks to find the MLE of the marginal likelihood by iteratively apply these two steps:
\begin{itemize}
\item \textbf{E-step:} Calculate the expected value of log-likelihood function, with respect to the conditional distribution of $Y, Z$ given $X$ under the current estimate of the parameters $(\pi^{(t)}, Q^ {(t)})$:
\begin{equation*}
Q(\pi,Q|\pi^{(t)}, Q^ {(t)})=E_{Y,Z|X,\pi^{(t)}, Q^ {(t)}}[\log L(\pi,Q;X,Y,Z)].
\end{equation*}
\item \textbf{M-step:} Find the parameter that maximizes this quantity:
\begin{equation*}
(\pi^{(t+1)}, Q^ {(t+1)}) =\argmax_{(\pi,Q)} Q(\pi,Q|\pi^{(t)}, Q^ {(t)}).
\end{equation*}
\end{itemize}

In the E-step, we have
\begin{eqnarray*}
&&\log L(\pi,Q;X,Y,Z) = \log p(X,Y, Z|\pi,Q)  \\
&=&  \sum_{i=1}^n \sum_{k=1}^K \mathbb {I}(z_i=k) \{ \log \pi_k + \sum_{r=1}^R y_{ir}[\log q_{kr} + \log f_{r1}(x_{ir})] \\
& +&  \sum_{r=1}^R (1-y_{ir})[\log (1-q_{kr}) + \log f_{r0}(x_{ir})] \}  \\
&=& \sum_{i=1}^n \sum_{k=1}^K \mathbb {I}(z_i=k) \log \pi_k +  \sum_{i=1}^n \sum_{k=1}^K \mathbb {I}(z_i=k) \{ \sum_{r=1}^R y_{ir}[\log q_{kr} + \log f_{r1}(x_{ir})] \\
& +&  \sum_{r=1}^R (1-y_{ir})[\log (1-q_{kr}) + \log f_{r0}(x_{ir})] \}.  \\
\end{eqnarray*}
Therefore,
\begin{eqnarray}
&& Q(\pi,Q|\pi^{(t)}, Q^ {(t)})=E_{Y,Z|X,\pi^{(t)}, Q^ {(t)}}[\log L(\pi,Q;X,Y,Z)] \nonumber \\
&=& \sum_{i=1}^n \sum_{k=1}^K p_{ik} \log \pi_k +\sum_{i=1}^n \sum_{k=1}^K \sum_{r=1}^R p_{ikr1}[\log q_{kr} + \log f_{r1}(x_{ir})] \nonumber \\
&+& \sum_{i=1}^n \sum_{k=1}^K \sum_{r=1}^R p_{ikr0}[\log (1-q_{kr}) + \log f_{r0}(x_{ir})],
\label{eq:Q}
\end{eqnarray}
where we denote 
\begin{eqnarray*}
 p_{ik} &=& E_{Y,Z|X,\pi^{(t)}, Q^ {(t)}}[\mathbb {I}(z_i=k)], \\
 p_{ikr1} &=&E_{Y,Z|X,\pi^{(t)}, Q^ {(t)}}[ \mathbb {I}(z_i=k) y_{ir}], \\
  p_{ikr0} &=&E_{Y,Z|X,\pi^{(t)}, Q^ {(t)}}[ \mathbb {I}(z_i=k)(1-y_{ir})].
 \end{eqnarray*}
It is easy to see that $ p_{ik} = p_{ikr1}+p_{ikr1}$. We can compute $p_{ik}$ and $p_{ikr1}$ by
\begin{eqnarray}
p_{ik} &=& E_{Y,Z|X,\pi^{(t)}, Q^ {(t)}}[\mathbb {I}(z_i=k)]  = Pr(z_i=k|X_i,\pi^{(t)}, Q^ {(t)}) \nonumber\\
&=& \frac{p(z_i=k,X_i|\pi^{(t)}, Q^ {(t)})} {p(X_i|\pi^{(t)}, Q^ {(t)})} \nonumber\\
&=& \frac{ \pi_k^{(t)}\prod_{r=1}^R [q_{kr}^{(t)}f_{r1}(x_{ir})+(1-q_{kr}^{(t)})f_{r0}(x_{ir})] }{\sum_{l=1}^K \pi_l^{(t)}\prod_{r=1}^R [q_{lr}^{(t)}f_{r1}(x_{ir})+(1-q_{lr}^{(t)})f_{r0}(x_{ir})] },
\label{eq:ik}
\end{eqnarray}
\begin{eqnarray}
p_{ikr1} &=& E_{Y,Z|X,\pi^{(t)}, Q^ {(t)}}[ \mathbb {I}(z_i=k) y_{ir}]=Pr(z_i=k, y_{ir}=1|X_i,\pi^{(t)}, Q^ {(t)}) \nonumber\\
&=& Pr(y_{ir}=1|z_i=k, X_i,\pi^{(t)}, Q^ {(t)}) \times Pr(z_i=k|X_i,\pi^{(t)}, Q^ {(t)}) \nonumber\\
&=& \frac{p(y_{ir}=1,x_{ir}|z_i=k,\pi^{(t)}, Q^ {(t)})} {p(x_{ir}|z_i=k, \pi^{(t)}, Q^ {(t)})}\times p_{ik} \nonumber \\
&=& \frac{q_{kr}^{(t)}f_{r1}(x_{ir})}{q_{kr}^{(t)}f_{r1}(x_{ir})+(1-q_{kr}^{(t)})f_{r0}(x_{ir})} \times p_{ik}.
\label{eq:ik1}
\end{eqnarray}
Take (\ref{eq:ik}) and (\ref{eq:ik1}) into (\ref{eq:Q}), we can obtain $Q(\pi,Q|\pi^{(t)}, Q^ {(t)})$. \\

In the M-step, we find $\pi^{(t+1)}$ and $Q^ {(t+1)}$ by maximize $Q(\pi,Q|\pi^{(t)}, Q^ {(t)})$. Notice that $\sum_k \pi_k=1$, we write the Lagrangian of the problem
\begin{equation*}
L(\pi,Q) = Q(\pi,Q|\pi^{(t)}, Q^ {(t)})+ \lambda(\sum_k \pi_k-1).
\end{equation*}
By solving
\begin{eqnarray*}
\frac{\partial L}{\partial \lambda} &=&0, \\
\frac{\partial L}{\partial \pi_k} &=&0, \\
\frac{\partial L}{\partial q_{kr}} &=& 0,
\end{eqnarray*}
we have
\begin{eqnarray}
\pi_k^{(t+1)} = \frac{1}{n} \sum_{i=1}^n p_{ik}, \\
q_{kr}^ {(t+1)}=\frac{\sum_{i=1}^n p_{ikr1}}{\sum_{i=1}^n p_{ik}}.
\end{eqnarray}
Therefore, we could iteratively use the EM algorithm and obtain the estimation for $\pi$ and $Q$.

\subsection{Model Selection: Bayesian Information Criterion}

To determine the motif number of $K$ , we use Bayesian Information Criterion(BIC). The BIC in our setting is written as 
\begin{eqnarray}
{\rm BIC}(K) &=& -2\log L(\hat \pi,\hat Q;X) + (K\times R+K-1+N_f)\times \log n \nonumber \\
&=& -2 \sum_{i=1}^n \log \left \{  \sum_{k=1}^K \hat \pi_k\prod_{r=1}^R [\hat q_{kr}f_{r1}(x_{ir})+(1-\hat q_{kr})f_{r0}(x_{ir})]   \right \} \nonumber\\
&+& (K\times R+K-1+N_f)\times \log n.
\end{eqnarray}
Here $N_f$ is the number of parameters in $f_{r0}$ and $f_{r1}$, $r = 1,\ldots, R$. We choose the K with the smallest BIC, that is 
\begin{equation}
\hat K = \argmin_{K\geq1} {\rm BIC}(K).
\end{equation}

\subsection{Example}

In this section, we will introduce a trivial example applying CorMotif method. Under our setting, suppose
\begin{eqnarray*}
f_{r0}(x_{ir}) &=& N(x_{ir};0,1), \\
f_{r1}(x_{ir}) &=& N(x_{ir};0,1+\sigma_r^2).
\end{eqnarray*}
We have
 $$\log f_{r1}(x_{ir})  = -\frac{1}{2} \log(1+\sigma_r^2)-\frac{x_{ir}^2}{2(1+\sigma_r^2)}+constant.$$
 Solving 
$$\frac{\partial Q}{\partial \sigma_r^2} =0 , $$
we get
$$\sigma_r^{2(t+1)}= \frac{\sum_{i=1}^n \sum_{k=1}^K (x_{ir}^2-1)p_{ikr1}}{\sum_{i=1}^n \sum_{k=1}^K p_{ikr1}}.$$

\begin{thebibliography}{2015}
\bibitem{ref1} Joint analysis of differential gene expression in multiple studies using correlation motifs, Ji HK and Wong WH, Biostatistics, 2015, doi: 10.1093/biostatistics/kxu038 
\bibitem{ref2} Linear models and empirical Bayes methods for assessing differential expression in microarray experiments. Smyth GK, Statistical Applications in Genetics and Molecular Biology 3, 3.
\end{thebibliography}

\end{document}

### 2017-10-17 paper/main.tex
update paper

--- 2017-09-25 paper/main.tex
@@
 
-\begin{assumption} The effects $\beta_1,\beta_2,\ldots,\beta_n$ are independent given $\pi$ and $g$.
-\end{assumption}
+%\begin{assumption} The effects $\beta_1,\beta_2,\ldots,\beta_n$ are independent given $\pi$ and $g$.
+%\end{assumption}
 
-Here we describe the simplest version of our model. First we assume the effect $\beta_{jr}$ is independent of their standard errors $\hat s_{jr}$. Then by Assumption 1, 2 and 3, we have
+Here we describe the simplest version of our model. First we assume the effect $\beta_{jr}$ is independent of their standard errors $\hat s_{jr}$. Then by Assumption 1 and 2, we have
 \begin{equation}
@@
 
-Another advantage of our generic model is that we could compute the posterior distribution for effects $\beta_{jr}$. By Bayes theorem, we have
-\begin{equation}
-    p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}) \propto p(\beta_{jr}|\hat s_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})
-    \label{eq:5}
-\end{equation}
+Another advantage of our generic model is that we could estimate the posterior distribution for effects $\beta_{jr}$, that is $p(\beta_{jr}|\hat\beta,\hat s,\pi,g) $. 
+%By Bayes theorem, we have
+%\begin{equation}
+%    p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}) \propto p(\beta_{jr}|\hat s_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})
+%    \label{eq:5}
+%\end{equation}
 
@@
 \begin{enumerate}
-    \item For a given $K$, estimate the parameters $\pi$ and $g$ by maximizing the likelihood $L(\pi,g)$, given by (\ref{eq:4}).
+    \item For a given $K$, estimate the parameters $\pi$ and $g$ by maximizing the likelihood $L(\pi,g)$, given by (\ref{eq:4}), denote as $\hat\pi$ and $\hat g$.
     \item Choose the best $K$ by minimizing the Bayesian Information Criterion (BIC).
-    \item Compute, for each $j$ and $r$, the posterior distribution $p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)$ by (\ref{eq:5}).
+    \item Compute, for each $j$ and $r$, the posterior distribution $p(\beta_{jr}|\hat\beta,\hat s,\hat\pi,\hat g)$.
 \end{enumerate}
@@
 All the details can be found in Section (\ref{sec:detail}).
-Step 1 is a convex problem, and we apply a EM algorithm to solve it.
+For Step 1 we apply an EM algorithm to solve it.
 Step 2 is straightforward.
-The conditional distributions $p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)$ by (\ref{eq:5}) in Step 3 are analytically available, each a mixture of a point mass on zero and $L$ normal distribution.
+The conditional distributions $p(\beta_{jr}|\hat\beta,\hat s,\hat\pi,\hat g)$ in Step 3 are analytically available, each a mixture of a point mass on zero and $L$ normal distribution.
 
@@
 \begin{equation}
-    \lfsr_{jr} := \min\{p(\beta_{jr}\geq 0|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g), p(\beta_{jr}\leq 0|\hat\beta_{jr},\hat s_{jr},\hat\pi,\hat g)\}
+    \lfsr_{jr} := \min\{p(\beta_{jr}\geq 0|\hat\beta,\hat s,\hat\pi,\hat g), p(\beta_{jr}\leq 0|\hat\beta,\hat s,\hat\pi,\hat g)\}
 \end{equation}
 
-Intutively, $\lfsr_{jr}$ is the probability that we would get the sign of effect $\beta_{jr}$ incorrect if we were to use our best guess of the sign.
+Intuitively, $\lfsr_{jr}$ is the probability that we would get the sign of effect $\beta_{jr}$ incorrect if we were to use our best guess of the sign.
 Therefore, a small $\lfsr$ indicates high confidence in determining the sign of an effect.
@@
 \end{equation}
-where $f_0$ is a point mass on 0. Then we have 
+Denote 
 \begin{equation}
-    p(\beta_{jr}|\pi,g,\hat s) = \sum_{k=1}^K\pi_k g_{kr}(\beta_{jr})=  \sum_{k=1}^K\pi_k\sum_{l=0}^Lw_l^{kr}f_l(\beta_{jr})
-    = \sum_{l=0}^L\eta_{lr}f_l(\beta_{jr}),
+     \tilde f_{l}(\hat\beta_{jr}) := \int f_l(\beta_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})d\beta_{jr},
 \end{equation}
-where $\eta_{lr}=\sum_{k=1}^K\pi_kw_l^{kr}$. Thus, by (\ref{eq:5}), we get
+as the likelihood of $\hat\beta_{jr}$ if the prior of $\beta_{jr}$ follows $f_l(\beta_{jr})$.
+
+By Bayes theorem, 
+\begin{eqnarray}
+&&p(\beta_{jr}|\hat\beta,\hat s,\pi, g) = p(\beta_{jr}|\hat\beta_{j},\hat s_j,\pi, g)\nonumber \\
+&=& \sum_{k=1}^Kp(\beta_{jr}|\hat\beta_{j},\hat s_j,\pi,g,z_j=k)p(z_j=k|\hat\beta_{j},\hat s_j,\pi, g) \nonumber\\
+&=& \sum_{k=1}^Kp(\beta_{jr}|\hat\beta_{jr},\hat s_j,\pi,g,z_j=k)p_{jk}\nonumber \\
+&=& \sum_{k=1}^K\frac{p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})g_{kr}(\beta_{jr})}{\int p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})g_{kr}(\beta_{jr})d\beta_{jr}}\times p_{jk} \nonumber\\
+&=& \sum_{k=1}^K \frac{\sum_{l=0}^Lw_l^{kr}\tilde f_{l}(\hat\beta_{jr})h_{l}(\beta_{jr}) }{\sum_{l'=0}^Lw_{l'}^{kr}\tilde f_{l'}(\hat\beta_{jr})} \times p_{jk} \nonumber\\
+&=& \sum_{l=0}^L \theta_{ljr}h_{l}(\beta_{jr}). 
+\end{eqnarray}
+Here the posterior mixture component $h_l$ is the posterior on $\beta_{jr}$ that would be obtained using prior $f_l(\beta_{jr})$ and likelihood $p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})$, and the mixture weights $\theta_{ljr}$ is
 \begin{equation}
-    p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\pi, g) = \sum_{l=0}^L\theta_{ljr}p_{l}(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}).
+     \theta_{ljr} = \sum_{k=1}^K\frac{w_{l}^{kr}\tilde f_{l}(\hat\beta_{jr})p_{jk}}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}.
 \end{equation}
-The posterior mixture component $p_l$ is the posterior on $\beta_{jr}$ that would be obtained using prior $f_l(\beta_{jr})$ and likelihood $p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})$.
+
+% where $f_0$ is a point mass on 0. Then we have 
+% \begin{equation}
+%     p(\beta_{jr}|\pi,g,\hat s) = \sum_{k=1}^K\pi_k g_{kr}(\beta_{jr})=  \sum_{k=1}^K\pi_k\sum_{l=0}^Lw_l^{kr}f_l(\beta_{jr})
+%     = \sum_{l=0}^L\eta_{lr}f_l(\beta_{jr}),
+% \end{equation}
+% where $\eta_{lr}=\sum_{k=1}^K\pi_kw_l^{kr}$. Thus, by (\ref{eq:5}), we get
+% \begin{equation}
+%     p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\pi, g) = \sum_{l=0}^L\theta_{ljr}p_{l}(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}).
+% \end{equation}
+% The posterior mixture component $p_l$ is the posterior on $\beta_{jr}$ that would be obtained using prior $f_l(\beta_{jr})$ and likelihood $p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})$.
 To compute the mixture weights $\theta_{ljr}$, denote
-\begin{equation}
-    \tilde f_{ljr}(\hat\beta_{jr}) := \int f_l(\beta_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})d\beta_{jr},
-\end{equation}
-the convolution of $f_l$ and a density function. Then $\theta_{ljr}$ can be obtained from
-\begin{equation}
-    \theta_{ljr} = \frac{\eta_{lr}\tilde f_{ljr}(\hat\beta_{jr})}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}.
-\end{equation}
+% \begin{equation}
+%     \tilde f_{ljr}(\hat\beta_{jr}) := \int f_l(\beta_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})d\beta_{jr},
+% \end{equation}
+% the convolution of $f_l$ and a density function. Then $\theta_{ljr}$ can be obtained from
+% \begin{equation}
+%     \theta_{ljr} = \frac{\eta_{lr}\tilde f_{ljr}(\hat\beta_{jr})}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}.
+% \end{equation}
 

### 2017-11-21 paper/main.tex
update paper
paper

@@
 \section{Introduction}
+
+We introduce a new Empirical Bayes approach for jointly analyzing differential gene expression in multiple studies or under multiple experimental conditions (e.g., measured in different cell types), which allows for correlations among studies.
+Compared with existing methods, the new approach improves power and effect-size estimation.
+Model fitting procedures are implemented using expectation maximization (EM).
 %\textit{lfdr}
@@
 \begin{equation}
-     \theta_{ljr} = \sum_{k=1}^K\frac{w_{l}^{kr}\tilde f_{l}(\hat\beta_{jr})p_{jk}}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}.
+     \theta_{ljr} = \sum_{k=1}^K\frac{w_{l}^{kr}\tilde f_{l}(\hat\beta_{jr})p_{jk}}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'}(\hat\beta_{jr})}.
 \end{equation}
@@
 % \begin{equation}
-%     \theta_{ljr} = \frac{\eta_{lr}\tilde f_{ljr}(\hat\beta_{jr})}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}.
+%     \theta_{ljr} = \frac{\eta_{lr}\tilde f_{ljr}(\hat\beta_{jr})}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'} (\hat\beta_{jr})}.
 % \end{equation}

### 2018-01-12 analysis/Simulation_generic_Cormotif.Rmd
add

@@
 # fit the model
-fit1 = generic.cormotif(betahat,sebetahat,K=4,mess=FALSE)
+fit1 = generic.cormotif(betahat,sebetahat,K=4,mess=TRUE)
 ```

### 2018-01-12 code/Untitled.R

@@ -192,7 +192,7 @@ for(g in 1:G){
   }
 }
 
-
+fit2 = generic.cormotif(betahat,sebetahat,K=2:5,mess=TRUE)
 
 ## sim 2
 K = 4

### 2018-01-12 code/function.R

@@ -1,4 +1,5 @@
 library("SQUAREM")
+library("REBayes")
 
 source('../code/lik.R')
 source('../code/set_data.R')
@@ -49,6 +50,7 @@ generic.cormotif=function(betahat,sebetahat,mixcompdist = "normal",optmethod = "
   
   ## fit the model via EM 
   fitresult<-list()
+  print('Start EM algorithm')
   for(i in 1:length(K)){
     fitresult[[i]] = gcmfit(matrix_llik,K=K[i],w_init=w,pi_thresh=pi_thresh,mess=mess)
   }
@@ -133,6 +135,7 @@ gcmfit = function(matrix_llik,K=1,max.iter=300,tol = 1e-4,w_init,pi_thresh,mess=
     tempmax = apply(matrix_llik2,1,max)
     matrix_lik2 = exp(matrix_llik2-tempmax)
     fit.p=mixEM(matrix_lik=matrix_lik2, prior=1,pi_init = p)
+    #fit.p=mixIP(matrix_lik=matrix_lik2, prior=rep(1,K),pi_init = p)
     p.new = normalize(fit.p$pihat + 1e-8)
     #print(p.new)
     
@@ -161,6 +164,7 @@ gcmfit = function(matrix_llik,K=1,max.iter=300,tol = 1e-4,w_init,pi_thresh,mess=
           lnorm = apply(matrix_llik3,1,max)
           matrix_lik3 = exp(matrix_llik3 - lnorm)
           fit.w = w_mixEM(matrix_lik=matrix_lik3, prior=1,pi_init = W[[r]][k,],weights = ww)
+          #fit.w = mixIP(matrix_lik=matrix_lik3, prior=rep(1,L),pi_init = W[[r]][k,],weights = ww)
           W.new[[r]][k,] = fit.w$pihat
         }
       }
@@ -218,6 +222,80 @@ comp_lfdsr= function(bestmotif, g, data){
   list(lfdr=lfdr,lfsr=lfsr,post_mean = post_mean,Theta=Theta)
 }
 
+layout(matrix(1:2,ncol=2))
+R = length(z18_k5$bestmotif$W)
+K = dim(z18_k5$bestmotif$W[[1]])[1]
+q = matrix(0,nrow=K,ncol=R)
+for(r in 1:R){
[…]

### 2018-01-12 analysis/MTB_Expression_Analysis.Rmd
wflow_publish("MTB_Expression_Analysis.Rmd")

---
title: "MTB Expression Analysis"
author: "Zhiwei Ma"
date: 2018-01-12
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
```

<!-- Add your analysis here -->

## Construction of Z-scores

For each bacterial infection at certain timepoint, we have six controls and six treatments, denoted as $X_1,X_2,X_3,X_4,X_5,X_6$ and $Y_1,Y_2,Y_3,Y_4,Y_5,Y_6$. Then the estimated effect size and variance are defined below:
$$
\hat\beta = \bar Y-\bar X; \\
\hat s = \sqrt{\hat\sigma_x^2/6+\hat\sigma_y^2/6.}
$$
Here $\bar X$, $\bar Y$ are sample means, and $\hat\sigma_x$, $\hat\sigma_y$ are sample 
variances.

## Motifs

At 18 hours post-infection, There are eight bacterial infections: "Rv.18", "Rv+.18", "GC.18", "BCG.18", "Smeg.18", "Yers.18", "Salm.18", "Staph.18". Here we set $K=5$, the same like [Blischak, et al, 2015](https://www.nature.com/articles/srep16882). The motifs are shown below.

```{r}
source('../code/function.R')
betahat = read.table("../data/betahat.txt", header = TRUE, sep = "\t",
               stringsAsFactors = FALSE)
sebetahat = read.table("../data/sebetahat.txt", header = TRUE, sep = "\t",
               stringsAsFactors = FALSE)
z18_k5 = generic.cormotif(betahat[,9:16],sebetahat[,9:16],K=5,mess=FALSE)
plotMotif(z18_k5$bestmotif)
```

## Detection of eQTLs

Download the 198 response eQTLs from [Barreiro et al, 2012](https://www.ncbi.nlm.nih.gov/pubmed/22233810) (179 out of the 198 were also expressed in our study). Apply our method, we can compute the $lfsr$ (local false sign rate) for each effect. Set the thresholding at 0.01, we detect 59 differential expression out of 179 genes.

```{r}
dat_cpm <- read.table("../data/table-s1.txt", header = TRUE, sep = "\t",
                      stringsAsFactors = FALSE)
# The first two columns are gene information
rownames(dat_cpm) <- dat_cpm$id
gene_names <- dat_cpm$name
names(gene_names) <- dat_cpm$id

id = rownames(dat_cpm)
lfsr18 = z18_k5$lfsr[[1]][,1] 
list = id[lfsr18<0.01]
a = intersect(S3[,2],list)
print(gene_names[a])
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2018-01-12 analysis/MTB_Expression_Analysis.Rmd
wflow_publish("MTB_Expression_Analysis.Rmd")

@@
 For each bacterial infection at certain timepoint, we have six controls and six treatments, denoted as $X_1,X_2,X_3,X_4,X_5,X_6$ and $Y_1,Y_2,Y_3,Y_4,Y_5,Y_6$.
-Then the estimated effect size and variance are defined below:
+Then the estimated effect size and standard error are defined below:
 $$
@@
 Here $\bar X$, $\bar Y$ are sample means, and $\hat\sigma_x$, $\hat\sigma_y$ are sample 
-variances.
+variances. Then we have
+$$
+\hat\beta|\hat s\sim N(\beta,\hat s^2).
+$$
 
@@
 
-```{r}
+```{r,cache=TRUE}
 source('../code/function.R')
@@
 # The first two columns are gene information
-rownames(dat_cpm) <- dat_cpm$id
 gene_names <- dat_cpm$name
@@
 
+S3 = read.csv("../data/S3.csv",
+              header = FALSE,stringsAsFactors = FALSE)
 id = rownames(dat_cpm)

### 2018-01-12 analysis/MTB_Expression_Analysis.Rmd
wflow_publish("MTB_Expression_Analysis.Rmd")

@@
 At 18 hours post-infection, There are eight bacterial infections: "Rv.18", "Rv+.18", "GC.18", "BCG.18", "Smeg.18", "Yers.18", "Salm.18", "Staph.18".
-Here we set $K=5$, the same like [Blischak, et al, 2015](https://www.nature.com/articles/srep16882).
-The motifs are shown below.
+Here we set $K=5$, the same as [Blischak et al, 2015](https://www.nature.com/articles/srep16882).
+The learned motifs are shown below.
 
@@
 
-S3 = read.csv("../data/S3.csv",
-              header = FALSE,stringsAsFactors = FALSE)
-id = rownames(dat_cpm)
+S3 = read.csv("../data/S3.csv",header = FALSE,stringsAsFactors = FALSE)
+id = names(gene_names)
+ll = intersect(S3[,2],id)
 lfsr18 = z18_k5$lfsr[[1]][,1] 

### 2018-01-12 code/Untitled.R
add

@@ -161,9 +161,10 @@ for(r in 1:R){
 ## sim1
 K = 4
 D = 4
-G = 1000
+G = 5000
 sigma2 = c(16,16,16,16)
-pi0 = c(0.02,0.04,0.04,0.9)
+#pi0 = c(0.02,0.04,0.04,0.9)
+pi0 = c(0.1,0.2,0.2,0.5)
 q0 = matrix(0,nrow=K,ncol=D)
 q0[1,] = c(1,1,1,1)
 q0[2,] = c(1,1,0,0)
@@ -192,7 +193,7 @@ for(g in 1:G){
   }
 }
 
-fit2 = generic.cormotif(betahat,sebetahat,K=2:5,mess=TRUE)
+fit2 = generic.cormotif(betahat,sebetahat,K=4,mess=TRUE)
 
 ## sim 2
 K = 4

### 2018-01-12 code/function.R

@@ -222,41 +222,7 @@ comp_lfdsr= function(bestmotif, g, data){
   list(lfdr=lfdr,lfsr=lfsr,post_mean = post_mean,Theta=Theta)
 }
 
-layout(matrix(1:2,ncol=2))
-R = length(z18_k5$bestmotif$W)
-K = dim(z18_k5$bestmotif$W[[1]])[1]
-q = matrix(0,nrow=K,ncol=R)
-for(r in 1:R){
-  q[,r] = 1-z18_k5$bestmotif$W[[r]][,1]
-}
-u = 1:R
-v = 1:K
-image(u,v,t(q),
-      col=gray(seq(from=1,to=0,by=-0.1)),xlab="Bacterial infection",yaxt = "n",
-      ylab="Expression pattern",main="pattern")
-axis(2,at=1:length(v))
-axis(1,at=1:length(u))
-for(i in 1:(length(u)+1)){
-  abline(v=(i-0.5))
-}
-for(i in 1:(length(v)+1)) {
-  abline(h=(i-0.5))
-}
-
-Ng=10000
-if(is.null(z18_k5$bestmotif$clustlike)!=TRUE)
-  Ng=nrow(z18_k5$bestmotif$clustlike)
-genecount=floor(z18_k5$bestmotif$pi*Ng)
-NK=K
-plot(0,0.7,pch=".",xlim=c(0,1.2),ylim=c(0.75,NK+0.25),
-     frame.plot=FALSE,axes=FALSE,xlab="Number of genes",ylab="", main="frequency")
-segments(0,0.7,z18_k5$bestmotif$pi[1],0.7)
-rect(0,1:NK-0.3,z18_k5$bestmotif$pi,1:NK+0.3,
-     col="dark grey")
-mtext(1:NK,at=1:NK,side=2,cex=0.8)
-text(z18_k5$bestmotif$pi+0.15,1:NK,
-     labels=floor(z18_k5$bestmotif$pi*Ng))
-
+## Plot motifs
 plotMotif = function(bestmotif){
   layout(matrix(1:2,ncol=2))
   R = length(bestmotif$W)
@@ -295,7 +261,5 @@ plotMotif = function(bestmotif){
 }
 
 #plotMotif(z18_k5$bestmotif)
-  
-  
 #results = generic.cormotif(betahat,sebetahat,K=1:4)
 

### 2018-01-29 analysis/Simulation_paper.Rmd
add

--- 2017-10-08 analysis/estimation_of_effect_size.Rmd
@@
 ---
-title: "estimation of effect sizes and detection of significant effects"
+title: "Simulation in Paper"
 author: "Zhiwei Ma"
-date: 2017-10-08
+date: 2018-01-28
 output: html_document
@@
 
-<!-- Add your analysis here -->
-
 ```{r setup, include=FALSE}
@@
 
-## Overview
-
-In this files, we analyze the accuracy of estimaiton of effect sizes and the detection of significant effects.
-To measure the accuracy, we apply the relative root mean squrared error (RRMSE).
-An RRMSE < 1 indicates the methods produces estimates that are more accurate than the original obserbation $\hat\beta$.
-For detection of significant effects, we require the estimated sign of each significant effect to be correct to be considered a "true positive".
+<!-- Add your analysis here -->
 
@@
 
-We set unit number $n = 5000$, and study number $R = 4$.
-We set every observation has the same standard error, $s_{jr}=1$.
+We set unit number $n = 10000$, and study number $R = 4$.
+We set every observation to be the same standard error, $s_{jr}=1$.
 That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 5000 units come from 4 patterns ($K=4$): 4500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 200 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 200 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3, 4$ and $\beta_{jr}=0$ for $r = 1, 2$; 100 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.
+The 10000 units come from 4 patterns ($K=4$): 9000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 400 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 400 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r = 1, 4$; 200 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$.
 
@@
 
-```{r,cache=TRUE}
+```{r}
 source('../code/function.R')
@@
 D = 4
-G = 5000
+G = 10000
 sigma2 = c(16,16,16,16)
@@
 
+The loglikelihood value plot:
+
+```{r}
+plot(fit1$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
+```
+
 The RRMSE for K = 1:10 is
 
-```{r,echo = FALSE}
+```{r}
 RRMSE = rep(0,10)
@@
 }
-plot(RRMSE,type="l",xlab="K",ylab="RRMSE",main="RRMSE for Simulation 1")
+plot(RRMSE,type="l",xlab="K",ylab="RRMSE")
 ```
@@
 
+The plot for lfsr
+
+```{r}
+plot(as.vector(fit1$lfsr[[4]][,1]),ylim=c(-0.1,1),type="l",xlab="Index",ylab="lfsr",col=1)
+lines(as.vector(fit1$lfsr[[10]][,1]-fit1$lfsr[[4]][,1]),lty=2,col=2)
+legend('right',legend=c("lfsr K=4","diff K=4,10"),lty =c(1,2),col=1:2)
+```
+
 Changing the threshold for lfsr, we plot the ROC curve for $K=1,2,4,10$.
 
-```{r,echo = FALSE}
+```{r}
 L = 10000
@@
 for(i in 1:L){
-  tpr[i] = mean(fit1$lfsr[[1]][beta != 0]<thres[i])
-  fpr[i] = mean(fit1$lfsr[[1]][beta == 0]<thres[i])
+  nT = sum(beta != 0)
+  nCSS = sum((fit1$lfsr[[1]]<=thres[i])&(fit1$post_mean[[1]]*beta>0))
+  nN = sum(beta == 0)
+  nNS = sum((fit1$lfsr[[1]]<=thres[i])&(beta == 0))
+  tpr[i] = nCSS/nT
+  fpr[i] = nNS/nN
 }
-plot(fpr,tpr,type="l",xlab="False Positive Rate",ylab="True Positive Rate",main ="ROC curve for Simulation 1",col=1)
+plot(fpr,tpr,type="l",xlab="False Positive Rate",ylab="True Positive Rate",col=1)
 for(i in 1:L){
-  tpr[i] = mean(fit1$lfsr[[2]][beta != 0]<thres[i])
-  fpr[i] = mean(fit1$lfsr[[2]][beta == 0]<thres[i])
+  nT = sum(beta != 0)
+  nCSS = sum((fit1$lfsr[[2]]<=thres[i])&(fit1$post_mean[[2]]*beta>0))
+  nN = sum(beta == 0)
+  nNS = sum((fit1$lfsr[[2]]<=thres[i])&(beta == 0))
+  tpr[i] = nCSS/nT
+  fpr[i] = nNS/nN
 }
@@
 for(i in 1:L){
-  tpr[i] = mean(fit1$lfsr[[4]][beta != 0]<thres[i])
-  fpr[i] = mean(fit1$lfsr[[4]][beta == 0]<thres[i])
+  nT = sum(beta != 0)
+  nCSS = sum((fit1$lfsr[[4]]<=thres[i])&(fit1$post_mean[[4]]*beta>0))
+  nN = sum(beta == 0)
+  nNS = sum((fit1$lfsr[[4]]<=thres[i])&(beta == 0))
+  tpr[i] = nCSS/nT
+  fpr[i] = nNS/nN
 }
@@
 for(i in 1:L){
-  tpr[i] = mean(fit1$lfsr[[10]][beta != 0]<thres[i])
-  fpr[i] = mean(fit1$lfsr[[10]][beta == 0]<thres[i])
+  nT = sum(beta != 0)
+  nCSS = sum((fit1$lfsr[[10]]<=thres[i])&(fit1$post_mean[[10]]*beta>0))
+  nN = sum(beta == 0)
+  nNS = sum((fit1$lfsr[[10]]<=thres[i])&(beta == 0))
+  tpr[i] = nCSS/nT
+  fpr[i] = nNS/nN
 }

### 2018-01-29 analysis/Simulation_paper2.Rmd
add

--- 2018-01-29 analysis/Simulation_paper.Rmd
@@
 That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 10000 units come from 4 patterns ($K=4$): 9000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 400 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 400 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r = 1, 4$; 200 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$.
+The 10000 units come from 4 patterns ($K=4$): 9100 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 300 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 300 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3, 4$ and $\beta_{jr}=0$ for $r = 1, 2$; 300 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$.
 
@@
 sigma2 = c(16,16,16,16)
-pi0 = c(0.02,0.04,0.04,0.9)
+pi0 = c(0.03,0.03,0.03,0.9)
 q0 = matrix(0,nrow=K,ncol=D)
@@
 
-### Simulation 2
-
-We set unit number $n = 5000$, and study number $R = 8$.
-We set every observation has the same standard error, $s_{jr}=1$.
-That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
-The 5000 units come from 5 patterns ($K=5$): 4000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4,5,6,7,8$; 250 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$ and $\beta_{jr}=0$ for $r =5, 6,7,8$; 250 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =5, 6,7,8$ and $\beta_{jr}=0$ for $r = 1,2,3,4$; 250 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3,4,5,6$ and $\beta_{jr}=0$ for $r = 1,2,7,8$; 250 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4,5,6,7,8$.
-
-To fit the model, set $K = 1:10$.
-
-```{r,cache=TRUE}
-source('../code/function.R')
-K = 5
-D = 8
-G = 5000
-sigma2 = rep(16,D)
-pi0 = c(0.05,0.05,0.05,0.05,0.8)
-q0 = matrix(0,nrow=K,ncol=D)
-q0[1,] = c(1,1,1,1,1,1,1,1)
-q0[2,] = c(1,1,1,1,0,0,0,0)
-q0[3,] = c(0,0,0,0,1,1,1,1)
-q0[4,] = c(0,0,1,1,1,1,0,0)
-q0[5,] = c(0,0,0,0,0,0,0,0)
-X = matrix(0,nrow=G,ncol=D)
-Y = matrix(0,nrow=G,ncol=D)
-rows = rep(1:K,times=pi0*G)
-A = q0[rows,]
-set.seed(222)
-for(g in 1:G){
-  for(d in 1:D){
-    if(A[g,d]==1){
-      X[g,d] = rnorm(1,0,sqrt(sigma2[d]))
-    } else{
-      X[g,d] = 0
-    }
-  }
-}
-beta = X
-sebetahat=matrix(1,nrow=G,ncol=D)
-betahat=matrix(10,nrow=G,ncol=D)
-for(g in 1:G){
-  for(d in 1:D){
-    betahat[g,d] = rnorm(1,beta[g,d],sebetahat[g,d])
-  }
-}
-# fit the model
-fit2 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
-```
-
-The RRMSE for K = 1:10 is
-
-```{r,echo = FALSE}
-RRMSE = rep(0,10)
-for(i in 1:10){
-  RRMSE[i] = sqrt(mean((fit2$post_mean[[i]]-beta)^2)/mean((betahat-beta)^2))
-}
-plot(RRMSE,type="l",xlab="K",ylab="RRMSE",main="RRMSE for Simulation 2")
-```
-
-Notice that when $K=1$, it is equvalent to applying ash to each study separately, and $K=5$ is the true model.
-
-We plot the ROC curve for $K=1,2,5,10$.
-
-```{r,echo = FALSE}
-L = 10000
-thres = seq(0.01,1,length=L)
-fpr = rep(0,L)
-tpr = rep(0,L)
-for(i in 1:L){
-  tpr[i] = mean(fit2$lfsr[[1]][beta != 0]<thres[i])
-  fpr[i] = mean(fit2$lfsr[[1]][beta == 0]<thres[i])
-}
-plot(fpr,tpr,type="l",xlab="False Positive Rate",ylab="True Positive Rate",main="ROC curve for Simulation 2",col=1)
-for(i in 1:L){
-  tpr[i] = mean(fit2$lfsr[[2]][beta != 0]<thres[i])
-  fpr[i] = mean(fit2$lfsr[[2]][beta == 0]<thres[i])
-}
-lines(fpr,tpr,col=2)
-for(i in 1:L){
-  tpr[i] = mean(fit2$lfsr[[5]][beta != 0]<thres[i])
-  fpr[i] = mean(fit2$lfsr[[5]][beta == 0]<thres[i])
-}
-lines(fpr,tpr,col=3)
-for(i in 1:L){
-  tpr[i] = mean(fit2$lfsr[[10]][beta != 0]<thres[i])
-  fpr[i] = mean(fit2$lfsr[[10]][beta == 0]<thres[i])
-}
-lines(fpr,tpr,col=4)
-legend('bottomright',legend=c("K=1","K=2","K=5","K=10"),lty =1,col=1:4)
-```
-
 ## Session information

### 2018-01-29 analysis/Simulation_paper3.Rmd
add

---
title: "Simulation in Paper"
author: "Zhiwei Ma"
date: 2018-01-28
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=TRUE)
```

<!-- Add your analysis here -->

## Results

### Simulation 1

We set unit number $n = 10000$, and study number $R = 8$. We set every observation has the same standard error, $s_{jr}=1$. That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$. The 10000 units come from 5 patterns ($K=5$): 8000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4,5,6,7,8$; 500 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$ and $\beta_{jr}=0$ for $r =5, 6,7,8$; 500 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =5, 6,7,8$ and $\beta_{jr}=0$ for $r = 1,2,3,4$; 500 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3,4,5,6$ and $\beta_{jr}=0$ for $r = 1,2,7,8$; 500 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4,5,6,7,8$.

To fit the model, set $K = 1:10$.

```{r,cache=TRUE}
source('../code/function.R')
K = 5
D = 8
G = 10000
sigma2 = rep(16,D)
pi0 = c(0.05,0.05,0.05,0.05,0.8)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1,1,1,1,1)
q0[2,] = c(1,1,1,1,0,0,0,0)
q0[3,] = c(0,0,0,0,1,1,1,1)
q0[4,] = c(0,0,1,1,1,1,0,0)
q0[5,] = c(0,0,0,0,0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
Y = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(222)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(sigma2[d]))
    } else{
      X[g,d] = 0
    }
  }
}
beta = X
sebetahat=matrix(1,nrow=G,ncol=D)
betahat=matrix(10,nrow=G,ncol=D)
for(g in 1:G){
  for(d in 1:D){
    betahat[g,d] = rnorm(1,beta[g,d],sebetahat[g,d])
  }
}
# fit the model
fit1 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The loglikelihood value plot:

```{r}
plot(fit1$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
```

The RRMSE for K = 1:10 is

```{r}
RRMSE = rep(0,10)
for(i in 1:10){
  RRMSE[i] = sqrt(mean((fit1$post_mean[[i]]-beta)^2)/mean((betahat-beta)^2))
}
plot(RRMSE,type="l",xlab="K",ylab="RRMSE")
```

Notice that when $K=1$, it is equvalent to applying ash to each study separately, and $K = 4$ is the true model.

The plot for lfsr

```{r}
plot(as.vector(fit1$lfsr[[5]][,1]),ylim=c(-0.1,1),type="l",xlab="Index",ylab="lfsr",col=1)
lines(as.vector(fit1$lfsr[[10]][,1]-fit1$lfsr[[5]][,1]),lty=2,col=2)
legend('right',legend=c("lfsr K=5","diff K=5,10"),lty =c(1,2),col=1:2)
```

Changing the threshold for lfsr, we plot the ROC curve for $K=1,2,5,10$.

```{r}
L = 10000
thres = seq(0.001,1,length=L)
fpr = rep(0,L)
tpr = rep(0,L)
for(i in 1:L){
  nT = sum(beta != 0)
  nCSS = sum((fit1$lfsr[[1]]<=thres[i])&(fit1$post_mean[[1]]*beta>0))
  nN = sum(beta == 0)
  nNS = sum((fit1$lfsr[[1]]<=thres[i])&(beta == 0))
  tpr[i] = nCSS/nT
  fpr[i] = nNS/nN
}
plot(fpr,tpr,type="l",xlab="False Positive Rate",ylab="True Positive Rate",col=1)
for(i in 1:L){
  nT = sum(beta != 0)
  nCSS = sum((fit1$lfsr[[2]]<=thres[i])&(fit1$post_mean[[2]]*beta>0))
  nN = sum(beta == 0)
  nNS = sum((fit1$lfsr[[2]]<=thres[i])&(beta == 0))
  tpr[i] = nCSS/nT
  fpr[i] = nNS/nN
}
lines(fpr,tpr,col=2)
for(i in 1:L){
  nT = sum(beta != 0)
  nCSS = sum((fit1$lfsr[[5]]<=thres[i])&(fit1$post_mean[[5]]*beta>0))
  nN = sum(beta == 0)
  nNS = sum((fit1$lfsr[[5]]<=thres[i])&(beta == 0))
  tpr[i] = nCSS/nT
  fpr[i] = nNS/nN
}
lines(fpr,tpr,col=3)
for(i in 1:L){
  nT = sum(beta != 0)
  nCSS = sum((fit1$lfsr[[10]]<=thres[i])&(fit1$post_mean[[10]]*beta>0))
  nN = sum(beta == 0)
  nNS = sum((fit1$lfsr[[10]]<=thres[i])&(beta == 0))
  tpr[i] = nCSS/nT
  fpr[i] = nNS/nN
}
lines(fpr,tpr,col=4)
legend('bottomright',legend=c("K=1","K=2","K=5","K=10"),lty =1,col=1:4)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2018-01-29 code/function.R

@@ -260,6 +260,44 @@ plotMotif = function(bestmotif){
        labels=floor(bestmotif$pi*Ng))
 }
 
+## Plot motifs
+plotMotif.mean = function(bestmotif, mixsd){
+  layout(matrix(1:2,ncol=2))
+  R = length(bestmotif$W)
+  K = dim(bestmotif$W[[1]])[1]
+  q = matrix(0,nrow=K,ncol=R)
+  for(r in 1:R){
+    q[,r] = bestmotif$W[[r]]%*%mixsd*sqrt(2/pi)
+  }
+  u = 1:R
+  v = 1:K
+  image(u,v,t(q),
+        col=gray(seq(from=1,to=0,by=-0.1)),xlab="Bacterial infection",yaxt = "n",
+        ylab="Expression pattern",main="pattern")
+  axis(2,at=1:length(v))
+  axis(1,at=1:length(u))
+  for(i in 1:(length(u)+1)){
+    abline(v=(i-0.5))
+  }
+  for(i in 1:(length(v)+1)) {
+    abline(h=(i-0.5))
+  }
+  
+  Ng=10000
+  if(is.null(bestmotif$clustlike)!=TRUE)
+    Ng=nrow(bestmotif$clustlike)
+  genecount=floor(bestmotif$pi*Ng)
+  NK=K
+  plot(0,0.7,pch=".",xlim=c(0,1.2),ylim=c(0.75,NK+0.25),
+       frame.plot=FALSE,axes=FALSE,xlab="Number of genes",ylab="", main="frequency")
+  segments(0,0.7,bestmotif$pi[1],0.7)
+  rect(0,1:NK-0.3,bestmotif$pi,1:NK+0.3,
+       col="dark grey")
+  mtext(1:NK,at=1:NK,side=2,cex=0.8)
+  text(bestmotif$pi+0.15,1:NK,
+       labels=floor(bestmotif$pi*Ng))
+}
+
 #plotMotif(z18_k5$bestmotif)
 #results = generic.cormotif(betahat,sebetahat,K=1:4)
 

### 2018-01-29 analysis/Simulation_paper2.Rmd
correct

@@
 sigma2 = c(16,16,16,16)
-pi0 = c(0.03,0.03,0.03,0.9)
+pi0 = c(0.03,0.03,0.03,0.91)
 q0 = matrix(0,nrow=K,ncol=D)

### 2018-01-30 analysis/Simulation_paper.Rmd
wflow_publish("/home/zwma/git/cormotif/analysis/Simulation_paper.Rmd")

@@
 D = 4
-G = 10000
+G = 50
 sigma2 = c(16,16,16,16)

### 2018-03-07 paper/main.tex
paper

--- 2017-10-17 paper/main.tex
@@
-\documentclass[11pt]{article}   	
-\usepackage{geometry}                		
-\geometry{letterpaper}                   		
-\usepackage{setspace}
-\doublespacing
+\documentclass[12pt]{article}
+\topmargin= -0.4in
+\textheight = +8.9in
+\oddsidemargin = 0.05in
+\evensidemargin = 0.05in
+\textwidth = 6.5in
 
+%%%%%%%%%%%%%
 \usepackage{graphicx}					
@@
 \usepackage{bm}
+\usepackage{hyperref}
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{enumerate}
 
@@
 \DeclareMathOperator*{\argmin}{argmin}
-
-\title{Generic Cormotif}
-\author{Zhiwei Ma}
-\date{September 2017}
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\maketitle
+\noindent
+\thispagestyle{empty}
+\underline{\bf Master's Paper of the Department of Statistics, the
+  University of Chicago} 
+\\~~(Internal document only, not for circulation)
 
+\vspace{1.8in}
+\begin{center}
+{\bf\LARGE  Joint analysis of differential gene expression in multiple studies
+using Bayesian method}
+%\\~\\
+%{\bf\Large --- A Sample Format}
+
+\vspace{1.4in}
+{\Large Zhiwei Ma}
+
+\vspace{1.3in}
+{\Large Advisor(s): Matthew Stephens} %\\{\small (at least
+  %one statistics faculty member)}}
+
+\end{center}
+
+\vspace{.6in}
+{\Large Approved} ~\underline{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}
+
+\vspace{.2in}
+{\Large Date} ~\underline{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}
+
+\vfill
+\begin{center}{\large February 15, 2018}\end{center}
+
+\newpage
+\pagestyle{plain}
+\setcounter{page}{1}
+
+\begin{abstract}
+
+\vspace{7mm}\noindent  We introduce a new Empirical Bayes approach for jointly
+analyzing differential gene expression in multiple studies or under
+multiple experimental conditions. Our method searches for a small number
+of latent prior distribution vectors to capture the dependence among
+multiple studies. Compared with independent analysis of each study, our joint analysis approach improves both power to detect significant effects and effect-size estimation. Model fitting procedures
+are implemented using expectation maximization (EM) algorithm. 
+\end{abstract}
+
+%\newpage
+\vspace{1.5in}
+\tableofcontents
+
+\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-%\textit{lfdr}
+%%%%%%%%%%%%%%%%%%%%%%
+
+\vspace{4mm} Detecting differentially expressed genes across multiple conditions or studies is an important task in the analysis of gene expression data.
+Examples including detecting the genes which are specific to MTB (Mycobacterium tuberculosis) infection \cite{Blischak2015}; or studying vertebrate sonic hedgehog (SHH) pathway \cite{Wei2015}; or identifying eQTLs in multiple cell-type or tissues \cite{Urbut2017}.
+
+\vspace{4mm}\noindent The simplest and the most common solution to this problem is to analyze data in different studies one at a time, and then identify the overlap of significant results in multiple studies.
+Methods for single study including \emph{limma} \cite{Smyth} or \emph{DESeq} \cite{Anders2010} to estimate effect sizes, followed by
+methods like \emph{qvalue} \cite{Storey2003} or \emph{ash} \cite{Stephens2017} to estimate false discovery rates.
+However when data from multiple related gene expression studies are available, separately analyzing each study is not ideal as it may fail to detect important genes with consistent but relatively weak differential signals in multiple studies.
+
+\vspace{4mm}\noindent To address the deficiencies of study-by-study analyses, several methods have been developed for joint analysis of effects under multiple studies.
+\emph{Cormotif} \cite{Wei2015} used a small number of latent probability vectors called ``correlation motifs'' to model the major correlation patterns among the studies.
+This approach can handle all possible study-specific differential patterns in a computationally-tractable way.
+However, its simplicity substantially reduce flexibility: firstly, \emph{Cormotif} adopts a two-stage design and uses moderated $t$-statistics produced by \emph{limma} \cite{Smyth}.
+The restrictive assumptions to the estimated effect sizes greatly weakens the adaptability in different situations; secondly, \emph{Cormotif} focus only on testing for the significant effects in each condition, but not on estimating effect sizes, which might be interesting in assessing heterogeneity of effects among conditions.
+
+\vspace{4mm}\noindent By applying an Empirical Bayesian method, \emph{mash} \cite{Urbut2017} builds a hierarchical model for jointly analyzing effect sizes.
+\emph{mash} is an extension of \emph{ash} by setting the priors of effect sizes as a mixture of multivariate normal distribution with both data-driven and canonical covariance matrices.
+However, \emph{mash} fails to include some common situations, such as when two studies have both positive or negative effects at the same time for some genes.
+In addition, simple normal distribution may not be able to capture the prior distributions for the effect sizes well.
+
+\vspace{4mm}\noindent Here we introduce a more flexible method that combines the attractive features of existing methods.
+Specifically, we extend the models in \emph{Cormotif} \cite{Wei2015} by allowing general unimodal distributions for the effect size distributions in each component.
+These general unimodal distributions are modeled using finite mixtures as in \emph{ash} \cite{Stephens2017}.
+Compared with \emph{Cormotif}, our approach can handle both effect size estimation and testing for significant effects by computing the posterior distributions of the effect sizes.
+Compared with \emph{mash}, our model includes the situation when two studies are conditional independent and we give more freedom for choosing the prior distribution.
+However, unlike \emph{mash} we do not allow for correlations in effect sizes across studies within each mixture component.
+Another feature of our approach is that our method is adaptive to both the amount of signal in the data and the measurement precision of each observation.
+
+\vspace{4mm}\noindent Our approach is a general method and can be applied to various problems in analyzing differential gene expression in multiple studies.
+We provide implementation in an R package, {\tt miximash}, available at \url{https://github.com/stephenslab/miximash}.
+Code and instructions for reproducing analyses and figures in this paper are at \url{https://github.com/stephenslab/cormotif}.
+
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\vspace{5mm}
 \subsection{Model Outline}
 
-Our method is designed to estimate the effects of multiple units in multiple studies.
+\vspace{4mm} Our method is designed to estimate the effects of multiple units in multiple studies.
 Suppose there are totally $n$ units and $R$ studies.
-Let $\beta = [\beta_{jr}]_{J\times R}$ denote "effects" of interest.
-For instance, $\beta_{jr}$ could be the difference in the mean expression of gene $j$ at study $r$ under two conditions.
+Let $\beta = [\beta_{jr}]_{J\times R}$ denote ``effects" of interest.
+For instance, $\beta_{jr}$ could be the difference in the mean (log) expression of gene $j$ under study $r$ in two conditions.
 
-Assume that the available data are estimates $\hat \beta = [\hat \beta_{jr}]_{J\times R}$ of the effects, and the corresponding estimated standard errors $\hat s = [\hat s_{jr}]_{J\times R}$.
+\vspace{4mm}\noindent Assume that the available data are estimates $\hat \beta = [\hat \beta_{jr}]_{J\times R}$ of the effects, and the corresponding estimated standard errors $\hat s = [\hat s_{jr}]_{J\times R}$.
 Let $\beta_j:=(\beta_{j1},\ldots,\beta_{jr})'$, where $j=1,2,\ldots,n$.
-Similarly for $\hat \beta_j$ and $\hat s_j$,
+Similar for $\hat \beta_j$ and $\hat s_j$.
 
-Similar to the original \textit{Cormotif} \cite{cormotif}, we assume that all units fall into $K$ different classes.
-Besides, we have the following assumptions, which distinguish our model from existing ones:
+\vspace{4mm}\noindent Similar to the original \emph{Cormotif} \cite{Wei2015}, we assume that all units fall into $K$ different classes to reduce the complexity of parameter space.
+Specifically, we have the following key assumptions:
 
@@
 
-%\begin{assumption} The effects $\beta_1,\beta_2,\ldots,\beta_n$ are independent given $\pi$ and $g$.
-%\end{assumption}
-
-Here we describe the simplest version of our model. First we assume the effect $\beta_{jr}$ is independent of their standard errors $\hat s_{jr}$. Then by Assumption 1 and 2, we have
+\vspace{4mm}\noindent Here we describe the simplest version of our model, more embellishments can be found in Detailed Method.
+First we assume the effect $\beta_{jr}$ is independent of their standard errors $\hat s_{jr}$.
+Then by Assumption 1 and 2, we have
 \begin{equation}
@@
 \end{equation}
-
-where $g$ represents $\{g_{kr}|k = 1,\ldots,K$, $r=1,\ldots,R\}$.
-A simple way to implement the unimodal assumption (UA) is to assume that $g_{kr}$ is a mixture of a point mass at 0 and a mixture of $zero$-$mean$ normal distribution \cite{ash}:
+where $g$ represents $\{g_{kr}|k = 1,\ldots,K$, $r=1,\ldots,R\}$, the set of unimodal distributions.
+A simple way to implement the unimodal assumption (UA) is to assume that $g_{kr}$ is a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distribution \cite{Stephens2017}:
 \begin{equation}
@@
 where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot;\mu,\sigma^2)$ denotes the density of Normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here we assume  $\sigma_1,\sigma_2,\ldots,\sigma_L$ are known and fixed positive numbers forming a wide and dense grid, which are built via a data-driven approach.
+Here we assume  $\sigma_1,\sigma_2,\ldots,\sigma_L$ are known and fixed positive numbers forming a wide and dense grid.
+By using a sufficiently large $L$ the finite mixture (\ref{eq:2}) can approximate any scale mixture of zero-centered normals to arbitrary accuracy.
+$\sigma_1,\sigma_2,\ldots,\sigma_L$ are built via a data-driven approach--see Implementation Details.
+The mixture proportions $w^{kr}=(w^{kr}_0,\ldots,w^{kr}_L)$ for $k=1,\ldots,K$, $r=1,\ldots,R$ and $\pi=(\pi_1,\ldots,\pi_K)$ are the parameters to be estimated.
 
-For likelihood $p(\hat \beta|\beta,\hat s)$, we assume a Normal approximation:
+\vspace{4mm}\noindent For likelihood $p(\hat \beta|\beta,\hat s)$, we assume a Normal approximation \cite{Wakefield2009}:
 \begin{equation}
@@
 
-Together, (\ref{eq:1})-(\ref{eq:3}) imply that 
+\vspace{4mm}\noindent Together, (\ref{eq:1})-(\ref{eq:3}) imply that 
 \begin{eqnarray}
@@
 \end{eqnarray}
-Here $N_{jr}$ denotes $N(\cdot;0,\hat s_{jr}^2)$ and $*$ means the convolution of two functions. We define $\sigma_0:=0$.
+Here $N_{jr}$ denotes $N(\cdot;0,\hat s_{jr}^2)$ and $*$ means the convolution of two functions. We define $\sigma_0:=0$ in the formula above.
 
-Model (\ref{eq:4}) is the extension of original  \textit{Cormotif}: set $L=1$ and assume for any fixed $r$, $\hat s_{jr}^2$s are identical for all $j$, (\ref{eq:4}) is just Cormotif under Normal distribution.
-Similar to Cormotif, our model can capture the correlation among multiple studies.
+\vspace{4mm}\noindent Actually, model (\ref{eq:4}) is the extension of the original \emph{Cormotif}: set $L=1$ and assume for any fixed $r$, $\hat s_{jr}^2$s are identical for all $j$, (\ref{eq:4}) is just \emph{Cormotif} model under Normal distribution.
+The difference here is that by using a larger $L$ we obtain more flexible models than \emph{Cormotif}.
+Similar to \emph{Cormotif}, our model can capture the dependence among multiple studies.
 To see this, consider the likelihood for unit $j$.
@@
 It is clear that $p(\hat\beta_j|\pi,g,\hat s) \neq \prod_{r=1}^Rp(\hat\beta_{jr}|\pi,g,\hat s)$, so different studies are dependent.
+Compared with \emph{Cormotif}, another advantage of our generic model is its focus on both testing for significant effects and estimating the effect sizes, which are realized by computing the posterior distribution of effect $\beta_{jr}$, $p(\beta_{jr}|\hat\beta,\hat s,\pi,g) $.
 
-When setting $K =1$, our model is equivalent to applying {\tt ash} \cite{ash} to each study separately.
-The advantage of our model is that it allows correlations among studies.
-This advantage leads to higher accuracy for effect estimation in multiple studies.
+\vspace{4mm}\noindent When setting $K =1$, our model is equivalent to applying \emph{ash} \cite{Stephens2017} to each study separately.
+Compared with \emph{ash}, our model allows dependence among studies.
+As we show in Simulation, this advantage improves both effect estimation and detection of significant effects in multiple studies.
 
-Another advantage of our generic model is that we could estimate the posterior distribution for effects $\beta_{jr}$, that is $p(\beta_{jr}|\hat\beta,\hat s,\pi,g) $. 
-%By Bayes theorem, we have
-%\begin{equation}
-%    p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}) \propto p(\beta_{jr}|\hat s_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})
-%    \label{eq:5}
-%\end{equation}
+%
+%\vspace{4mm}\noindent The model has several additional assumptions can be relaxed, specifically,
+%\begin{enumerate}
+%\item In (\ref{eq:2}) we assume $g_{kr}$ is symmetric about 0. We can release the setting by using mixtures of uniforms. In this way, $g_{kr}$ can approximate any unimodal distribution. 
+%\item The normal distribution in (\ref{eq:4}) can be generalized to a t distribution. 
+%\end{enumerate}
 
+\vspace{5mm}
 \subsection{Fitting the model}
 
-Our method involves three steps:
+\vspace{4mm} Our method involves two steps:
 \begin{enumerate}
-    \item For a given $K$, estimate the parameters $\pi$ and $g$ by maximizing the likelihood $L(\pi,g)$, given by (\ref{eq:4}), denote as $\hat\pi$ and $\hat g$.
-    \item Choose the best $K$ by minimizing the Bayesian Information Criterion (BIC).
+    \item For a large enough $K$, estimate the parameters $\pi$ and $g$ by maximizing the likelihood $L(\pi,g)$, given by (\ref{eq:4}), denote as $\hat\pi$ and $\hat g$.
     \item Compute, for each $j$ and $r$, the posterior distribution $p(\beta_{jr}|\hat\beta,\hat s,\hat\pi,\hat g)$. 
 \end{enumerate}
+All the details can be found in Implementation Details.
+For Step 1, we  have totally $K\times R\times L+K-1$ parameters to estimate.
+We apply an EM algorithm to solve it (Alternatively interior point (IP) method \cite{Koenker2014} can be applied here as well).
+The conditional distributions $p(\beta_{jr}|\hat\beta,\hat s,\hat\pi,\hat g)$ in Step 2 are analytically available, each a mixture of a point mass on zero and $L$ normal distribution.
+With the posterior distributions for the effect, one can estimate the effect size by $E(\beta_{jr}|\hat\beta,\hat s,\hat\pi,\hat g)$ and test for non-zero effects.
 
-All the details can be found in Section (\ref{sec:detail}).
-For Step 1 we apply an EM algorithm to solve it.
-Step 2 is straightforward.
-The conditional distributions $p(\beta_{jr}|\hat\beta,\hat s,\hat\pi,\hat g)$ in Step 3 are analytically available, each a mixture of a point mass on zero and $L$ normal distribution.
-
+\vspace{5mm}
 \subsection{Local False Sign Rate}
 
-To measure "significance" of an effect $\beta_{jr}$ we use the local false sign rate ($\lfsr$), which is defined as:
+\vspace{4mm} To measure ``significance" of an effect $\beta_{jr}$ we use the local false sign rate ($\lfsr$) \cite{Stephens2017}, which is defined as:
 \begin{equation}
-    \lfsr_{jr} := \min\{p(\beta_{jr}\geq 0|\hat\beta,\hat s,\hat\pi,\hat g), p(\beta_{jr}\leq 0|\hat\beta,\hat s,\hat\pi,\hat g)\}
+    \lfsr_{jr} := \min\{p(\beta_{jr}\geq 0|\hat\beta,\hat s,\hat\pi,\hat g), p(\beta_{jr}\leq 0|\hat\beta,\hat s,\hat\pi,\hat g)\}.
+\label{eq:lfsr}
 \end{equation}
 
-Intuitively, $\lfsr_{jr}$ is the probability that we would get the sign of effect $\beta_{jr}$ incorrect if we were to use our best guess of the sign.
+\vspace{4mm}\noindent Intuitively, $\lfsr_{jr}$ is the probability that we would get the sign of effect $\beta_{jr}$ incorrect if we were to use our best guess of the sign.
 Therefore, a small $\lfsr$ indicates high confidence in determining the sign of an effect.
-Notice that $\lfsr$ is more conservative than the local false discovery rate ($\lfdr$), since we can infer $\lfsr_{jr}\geq\lfdr_{jr}$ from the definition.
+Notice that $\lfsr$ is more conservative than the local false discovery rate ($\lfdr$) since $\lfsr_{jr}\geq\lfdr_{jr}$.
 
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulation} 
+%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/pa3.png}
+    \caption{True pattern} 
+\label{fig:pa3}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/pattern3.png}
+    \caption{Estimated pattern} \label{fig:pattern3}
+\end{subfigure}
+\caption{Patterns learned in Simulation 1 (constant precision, $\hat s_{jr}=1$).} \label{fig:pat3}
+\end{figure}
+
+\begin{figure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/llik3.png}
+    \caption{Log likelihood function for different $K$} 
+\label{fig:llik3}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/RRMSE3.png}
+    \caption{Accuracy of effect estimates (RRMSE)} \label{fig:rrmse3}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/lfsr3.png}
+    \caption{$\lfsr$ for the first study} \label{fig:lfsr3}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/ROC3.png}
+    \caption{Detection of non-null effects (ROC curves)} \label{fig:roc3}
+\end{subfigure}
+\caption{Results for Simulation 1 } \label{fig:sim3}
+\end{figure}
+
+\vspace{5mm}
+\subsection{Estimation of patterns}
+
+\vspace{4mm} When set $K=1$ in our model, our model is equivalent to applying \emph{ash} to each study separately.
+For comparison, we would like to check the performance for $K=1$ and $K\geq 2$ separately.
+The increase in the performance can be viewed as the gain by considering the correlation among studies.
+
+\vspace{4mm}\noindent In simulation 1, we set unit number $n = 10000$, and study number $R = 8$.
+Assume every observation has the same standard error, $s_{jr}=1$.
+That is, $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
+The 10000 units come from 5 patterns ($K=5$): 8000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for all $r = 1,2,3,4,5,6,7,8$; 500 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$ and $\beta_{jr}=0$ for $r =5, 6,7,8$; 500 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =5, 6,7,8$ and $\beta_{jr}=0$ for $r = 1,2,3,4$; 500 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3,4,5,6$ and $\beta_{jr}=0$ for $r = 1,2,7,8$; 500 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4,5,6,7,8$.
+
+\vspace{4mm}\noindent Figure \ref{fig:pa3} shows the true patterns and its frequencies.
+In the left plot, the value in $k$-th row and $r$-th column is $p(\beta_{jr}\neq 0|z_j=k,g)=1-w_0^{kr}$.
+The right plot shows the frequencies of each pattern.
+For estimation, we set $K=5$ and the estimated patterns are displayed in Figure \ref{fig:pattern3}.
+The estimated pattern are very similar to the true underlying differential patterns in Figure \ref{fig:pa3}.
+
+\vspace{5mm}
+\subsection{Improved effect size estimates}
+
+\vspace{4mm}\noindent We fit our model for $K$ increasing from 1 to 10.
+Since our model is nested for $K$ (for larger $K$, we can always set some $\pi_k$ to be zero and get the same log likelihood function), the log likelihood function for larger $K$ is no less than smaller ones.
+Figure \ref{fig:llik3} shows the log likelihood function in (\ref{eq:4}) for different $K$.
+The log likelihood function has a non-decreasing trend.
+From $K=1$ to $2$, the log likelihood function has the most significant increase.
+And for $K\geq 5$, the log likelihood function remains approximately constant.
+The increasing trend shows that our hierarchical model fits better than applying {\tt ash} independently across studies (K=1).
+Then trend of log likelihood values after $K=5$ is due to that only five $\hat\pi_k$ get significant non-zeros values, while all others are estimated to be almost zeros.
+
+\vspace{4mm}\noindent Figure \ref{fig:rrmse3} compares the accuracy of effect size estimates by relative root mean squared error (RRMSE).
+Denote estimates for $\beta_{jr}$ is $\tilde\beta_{jr}$ by our method.
+The RRMSE for estimates $\tilde\beta_{jr}$ is computed as
+\begin{equation}
+\textrm{RRMSE} = \frac{\sqrt{\frac{1}{NR}\sum_{j=1}^N\sum_{r=1}^R(\tilde\beta_{jr}-\beta_{jr})}}{\sqrt{\frac{1}{NR}\sum_{j=1}^N\sum_{r=1}^R(\hat\beta_{jr}-\beta_{jr})}}, 
+\end{equation}
+which is the RMSE of estimates $\tilde\beta_{jr}$ divided by the RMSE achieved by simply using the original observed estimates $\hat\beta_{jr}$.
+Therefore, RRMSE less than 1 indicates that the estimates is more accurate than original observation  $\hat\beta_{jr}$.
+From Figure \ref{fig:rrmse3}, we can find that even $K=1$ gives approximate 0.5 RRMSE.
+In addition, the RRMSE has a decline trend as $K$ increases and reaches approximate constance for $K\geq 5$.
+Compared with \emph{ash} ($K=1$), our method has a substantial improvement in accuracy for large enough $K$.
+
+\vspace{5mm}
+\subsection{Improved detection of significant effects}
+
+\vspace{4mm} In addition to effect estimates, our method also provides a measure of significance for each effect by \lfsr defined in (\ref{eq:lfsr}).
+We used the same simulation as above to illustrate the gains in power to detect significant effects coming from our method.
+Figure \ref{fig:lfsr3} shows the $\lfsr$ values computed by our method.
+The black dot represent the estimated $\lfsr$ in the first study for $K=5$.
+Notice that in the first study, we only set the first 1000 units to own significant effects ($\beta_{jr}\sim N(0,4^2)$) out of totally 10000 units.
+For the first 1000 units, the estimated $\lfsr$ tend to gather to 0, which means we are confident that they are nonzero.
+The following units have $\lfsr$ gathering at 1, which means they have zero effects.
+The red line shows the negative absolutely values of the differences of $\lfsr$ between $K=5$ and $K=10$.
+Except for the first 2000 units, the differences are very small in each unit, and almost all differences are less than 0.1.
+This again illustrate that large enough $K$ will give similar results as the exact one.
+
+\vspace{4mm}\noindent Figure \ref{fig:roc3} shows the trade-off between false positive and true positive discoveries for different $K$ (ROC curves).
+The True Positive Rate and False Positive Rate are computed at any given threshold $\gamma$ as
+\begin{eqnarray}
+\textmd{True Positive Rate} &:=& \frac{|CS\cap S|}{|T|}\\
+\textmd{False Positive Rate} &:=& \frac{|N\cap S|}{|N|}
+\end{eqnarray}
+where $S$ is the set of significant effects at threshold $\gamma$, $CS$ is the set of correctly signed effects, $T$ is the set of true (non-zeros) effects and $N$ is the set of null (zero) effects:
+\begin{eqnarray}
+S &:=& \{(j,r)|\lfsr_{jr}\leq\gamma \}\\
+CS &:=& \{(j,r)|E(\beta_{jr}|D)\times\beta_{jr}>0\}\\
+N &:=& \{(j,r)|\beta_{jr}=0 \}\\
+T &:=& \{(j,r)|\beta_{jr}\neq 0 \}
+\end{eqnarray}
+A true positive needs to be correctly signed, not only significant.
+The performance precisely mirrors the RRMSE results: our method outperforms \emph{ash} and the performance is similar for large enough $K$.
+
+\vspace{4mm}\noindent Using a similar approach, we performed simulations 2, which involved different study number and differential expression patterns, see Appendix Figure \ref{fig:sim2}.
+The conclusion is similar to simulation 1.
+
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Detailed Method} 
+%%%%%%%%%%%%%%%%%%%%%%%%
 
+\vspace{5mm}
 \subsection{Embellishments}
 
+\vspace{4mm} We can extend $g_{kr}$ in (\ref{eq:2}) to other symmetric or even asymmetric unimodal distributions. For a more general case, suppose
+\begin{equation}
+g_{kr}(\cdot;w) = \sum_{l=0}^L w_l^{kr}f_l(\cdot),
+\label{eq:geng}
+\end{equation}
+where $f_0$ is a point mass at 0, and $f_l$ $(l=1,\ldots,L)$ are component distribution with one of the following forms:
+\begin{enumerate}[(i)]
+\item $f_l(\cdot) = N(\cdot; 0, \sigma^2_l)$,
+\item $f_l(\cdot) = U[\cdot; -a_l,a_l]$,
+\item $f_l(\cdot) = U[\cdot; -a_l,0] \text{ and/or } U[\cdot; 0,a_l]$,
+\end{enumerate}
+where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
+Actually, with dense enough $a_l$, (iii) can approximate any unimodal distribution about 0.
+To better approximate the unimodal distribution, larger $K$ should be better.
+However in practice, a moderate $K$ can provide reasonable performance.
+The detailed method for estimating $\sigma_l^2$ and $a_l$ form data can be found in the Implementation Details in \cite{Stephens2017}.
+
+\vspace{5mm}
 \subsection{Implementation Details} \label{sec:detail}
 
+\vspace{5mm}
+\subsubsection{Likelihood}
+
+\vspace{4mm} Using the prior $\beta_{jr}|z_j=k \sim \sum_{l=0}^L w_l^{kr}f_l(\beta_{jr})$ given by (\ref{eq:geng}) and the normal likelihood in (\ref{eq:3}), integrating over $\beta_{jr}$ yields
+\begin{equation}
+p(\hat\beta_{jr}|z_j=k,\pi,g,\hat s) = \sum_{l=0}^L w_l^{kr}\tilde f_l(\hat\beta_{jr}),
+\end{equation}
+where
+\begin{equation}
+     \tilde f_{l}(\hat\beta_{jr}) := \int f_l(\beta_{jr})N(\hat\beta_{jr};\beta_{jr},\hat s_{jr})d\beta_{jr},
+\end{equation}
+denotes the likelihood of $\hat\beta_{jr}$ if the prior of $\beta_{jr}$ follows $f_l(\beta_{jr})$.
+
+\vspace{4mm}\noindent These convolutions are straightforward to evaluate when $f_l$ is a normal or uniform density. Specifically, 
+\begin{equation} \label{eqn:uconv}
+\tilde{f}_l(\hat\beta_{jr})  = 
+\begin{cases}
+N(\hat\beta_{jr};0, \hat s_{jr}^2 +  \sigma_l^2) & \text{if $f_l(\cdot) = N(\cdot; 0, \sigma_l^2)$}, \\
+\frac{\Phi((\hat\beta_{jr}-a_l)/\hat s_{jr}) - \Phi((\hat\beta_{jr}-b_l)/\hat s_{jr})}{b_l-a_l} & \text{if $f_l(\cdot) = U(\cdot; a_l,b_l)$},
+\end{cases}
+\end{equation}
+where $\Phi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
+
+\vspace{5mm}
 \subsubsection{Optimization}
-This section We presents the EM algorithm used to estimate both $\pi$ and $g$. First compute the log likelihood function for $\hat\beta$ and group label $z=(z_1,\ldots,z_n)'$:
+
+\vspace{4mm}\noindent  This section We presents the EM algorithm used to estimate both $\pi$ and $g$. First compute the log likelihood function for $\hat\beta$ and group label $z=(z_1,\ldots,z_n)'$:
 \begin{eqnarray}
@@
     +\sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\log p(z_j=k|\pi)\nonumber\\
-    &=& \sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\sum_{r=1}^R\log \left[(g_{kr}*N_{jr})(\hat\beta_{jr}) \right]+\sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\log\pi_k.
+    &=& \sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\sum_{r=1}^R\log \left[p(\hat\beta_{jr}|z_j=k,\pi,g,\hat s) \right]+\sum_{j=1}^n\sum_{k=1}^K\mathbb{I}(z_j=k)\log\pi_k.
 \end{eqnarray}
-
 Here $z$ is a latent variable. The EM algorithm seeks to find the MLE of the marginal likelihood (\ref{eq:4}) by iteratively applying the E-step and the M-step.
 
-In the E-step, one evaluates the $Q$-function $Q(\pi,g|\pi^{(t)},g^{(t)})$, here $(\pi^{(t)},g^{(t)})$ is the current estimation. We have
+\vspace{4mm}\noindent In the E-step, one evaluates the $Q$-function $Q(\pi,g|\pi^{(t)},g^{(t)})$, here $(\pi^{(t)},g^{(t)})$ is the current estimation. We have
 \begin{eqnarray}
     &&Q(\pi,g|\pi^{(t)},g^{(t)}) = E_{z|\hat\beta,\hat s,\pi^{(t)},g^{(t)}}\left[\log p(\hat\beta,z|\pi,g,\hat s)\right]\nonumber\\
-    &=&\sum_{j=1}^n\sum_{k=1}^K\sum_{r=1}^Rp_{jk}\log \left[(g_{kr}*N_{jr})(\hat\beta_{jr}) \right]+\sum_{j=1}^n\sum_{k=1}^Kp_{jk}\log\pi_k,
+    &=&\sum_{j=1}^n\sum_{k=1}^K\sum_{r=1}^Rp_{jk}\log \left[p(\hat\beta_{jr}|z_j=k,\pi,g,\hat s)\right]+\sum_{j=1}^n\sum_{k=1}^Kp_{jk}\log\pi_k,
 \label{eq:Q}
@@
     &=& \frac{p(\hat\beta_j,z_j=k|\hat s,\pi^{(t)},g^{(t)})}{p(\hat\beta_j|\hat s,\pi^{(t)},g^{(t)})}\nonumber\\
-    &=& \frac{\pi_k^{(t)}\prod_{r=1}^R(g_{kr}^{(t)}*N_{jr})(\hat\beta_{jr})}{\sum_{k'=1}^K\pi_{k'}^{(t)}\prod_{r=1}^R(g_{k'r}^{(t)}*N_{jr})(\hat\beta_{jr})}
+    &=& \frac{\pi_k^{(t)}\prod_{r=1}^Rp(\hat\beta_{jr}|z_j=k,\pi^{(t)},g^{(t)},\hat s)}{\sum_{k'=1}^K\pi_{k'}^{(t)}\prod_{r=1}^Rp(\hat\beta_{jr}|z_j=k,\pi^{(t)},g^{(t)},\hat s)}
 \end{eqnarray}
 
-In the M-step, one finds $\pi$ and $g$ that maximize the $Q$-function $Q(\pi,g|\pi^{(t)},g^{(t)})$, and denote them as $\pi^{(t+1)}$ and $g^{(t+1)}$, that is 
+\vspace{4mm}\noindent In the M-step, one finds $\pi$ and $g$ that maximize the $Q$-function $Q(\pi,g|\pi^{(t)},g^{(t)})$, and denote them as $\pi^{(t+1)}$ and $g^{(t+1)}$, that is 
 \begin{equation}
@@
 Notice in (\ref{eq:Q}), we could separately optimize $g_{kr}$ for fixed $k$ and $r$, that is 
-\begin{equation}
-    g_{kr}^ {(t+1)}=\argmax_{g_{kr}}\sum_{j=1}^np_{jk}\log \left[(g_{kr}*N_{jr})(\hat\beta_{jr}) \right].
+\begin{eqnarray}
+    g_{kr}^ {(t+1)}&=&\argmax_{g_{kr}}\sum_{j=1}^np_{jk}\log \left[p(\hat\beta_{jr}|z_j=k,\pi,g,\hat s) \right] \nonumber \\
+ &=&\argmax_{w^{kr}}\sum_{j=1}^np_{jk}\log\left[\sum_{l=0}^L w_l^{kr}\tilde f_l(\hat\beta_{jr}) \right].
 \label{eq:g_kr}
-\end{equation}
-Optimizing (\ref{eq:g_kr}) is a convex problem, which we solve using an EM algorithm, accelerated using R package {\tt SQUAREM} .
+\end{eqnarray}
+Optimizing (\ref{eq:g_kr}) is a convex problem, which we solve using an EM algorithm, accelerated using R package {\tt SQUAREM}.
+A simple interior point (IP) methods could also be applied here using R package {\tt REBayes}.
 
-\subsubsection{Model Selection: Bayesian Information Criterion}
-
-To determine the class number of $K$ , we use Bayesian Information Criterion(BIC). The BIC in our setting is written as 
-\begin{eqnarray}
-{\rm BIC}(K) &=& -2\log p(\hat\beta|\pi,g,\hat s) + (K\times R\times L+K-1)\times \log n \nonumber \\
-&=& -2 \sum_{j=1}^n\log\left[\sum_{k=1}^K\pi_k\prod_{r=1}^R(g_{kr}*N_{jr})(\hat\beta_{jr})\right] \nonumber\\
-&+& (K\times R\times L+K-1)\times \log n.
-\end{eqnarray}
-Here $K-1$ is the number of parameters for $\pi$, $K\times R\times L$ is the number of parameters involved in $g$ and $n$ is the unit number. We choose the K with the smallest BIC, that is 
-\begin{equation}
-\hat K = \argmin_{K\geq1} {\rm BIC}(K).
-\end{equation}
-
+\vspace{5mm}
 \subsubsection{Posterior distribution}
 
-For a more general case, suppose
-\begin{equation}
-g_{kr}(\cdot;w) = \sum_{l=0}^L w_l^{kr}f_l(\cdot),
-\label{eq:geng}
-\end{equation}
-Denote 
-\begin{equation}
-     \tilde f_{l}(\hat\beta_{jr}) := \int f_l(\beta_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})d\beta_{jr},
-\end{equation}
-as the likelihood of $\hat\beta_{jr}$ if the prior of $\beta_{jr}$ follows $f_l(\beta_{jr})$.
-
-By Bayes theorem, 
+\vspace{4mm} By Bayes theorem, 
 \begin{eqnarray}
@@
 \begin{equation}
-     \theta_{ljr} = \sum_{k=1}^K\frac{w_{l}^{kr}\tilde f_{l}(\hat\beta_{jr})p_{jk}}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}.
+     \theta_{ljr} = \sum_{k=1}^K\frac{w_{l}^{kr}\tilde f_{l}(\hat\beta_{jr})p_{jk}}{\sum_{l'=0}^Lw_{l'}^{kr}\tilde f_{l'}(\hat\beta_{jr})}.
 \end{equation}
 
-% where $f_0$ is a point mass on 0. Then we have 
-% \begin{equation}
-%     p(\beta_{jr}|\pi,g,\hat s) = \sum_{k=1}^K\pi_k g_{kr}(\beta_{jr})=  \sum_{k=1}^K\pi_k\sum_{l=0}^Lw_l^{kr}f_l(\beta_{jr})
-%     = \sum_{l=0}^L\eta_{lr}f_l(\beta_{jr}),
-% \end{equation}
-% where $\eta_{lr}=\sum_{k=1}^K\pi_kw_l^{kr}$. Thus, by (\ref{eq:5}), we get
-% \begin{equation}
-%     p(\beta_{jr}|\hat\beta_{jr},\hat s_{jr},\pi, g) = \sum_{l=0}^L\theta_{ljr}p_{l}(\beta_{jr}|\hat\beta_{jr},\hat s_{jr}).
-% \end{equation}
-% The posterior mixture component $p_l$ is the posterior on $\beta_{jr}$ that would be obtained using prior $f_l(\beta_{jr})$ and likelihood $p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})$.
-To compute the mixture weights $\theta_{ljr}$, denote
-% \begin{equation}
-%     \tilde f_{ljr}(\hat\beta_{jr}) := \int f_l(\beta_{jr})p(\hat\beta_{jr}|\beta_{jr},\hat s_{jr})d\beta_{jr},
-% \end{equation}
-% the convolution of $f_l$ and a density function. Then $\theta_{ljr}$ can be obtained from
-% \begin{equation}
-%     \theta_{ljr} = \frac{\eta_{lr}\tilde f_{ljr}(\hat\beta_{jr})}{\sum_{l'=0}^L\eta_{l'r}\tilde f_{l'jr}(\hat\beta_{jr})}.
-% \end{equation}
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{thebibliography}{99}
+\vspace{4mm} In this paper, we introduce an Empirical Bayesian method for jointly analyzing differential gene expression in multiple studies.
+To capture the dependence among different studies, we introduce a small number of latent patterns for prior distributions of effect sizes.
+Our approach focus on both estimation of effect sizes and testing for significant effects.
+Simulation results show that our method outperforms the single study approach \emph{ash} in both effect-size estimation and detecting significant effects.
+In addition, in simple settings our approach can accurately estimate the true patterns in the data.
 
-\bibitem{cormotif} 
-Wei, Y. and Ji, H. (2014). Joint analysis of differential gene expression in multiple studies using correlation motif. {\it Biostatistics}, doi:10.1093/biostatistics/
-kxu038.
+\newpage
 
-\bibitem{ash} 
-Stephens, M. (2016). False discovery rate: A new deal. {\it Biostatistics} 18 (2): 275-294
+\begin{appendix}
+\section{Appendix}
 
-\bibitem{mash}
-Urbut, S.M., Wang, G. and Stephens, M. (2017). Flexible statistical methods for estimating and testing effects in genomic studies with multiple conditions. {\it bioRxiv}, doi:10.1101/096552.
+\begin{figure}[ht]
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/llik2.png}
+    \caption{Log likelihood function for different $K$} 
+\label{fig:llik2}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/RRMSE2.png}
+    \caption{Accuracy of effect estimates (RRMSE)} \label{fig:rrmse2}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/lfsr2.png}
+    \caption{$\lfsr$ for the first study} \label{fig:lfsr2}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{./figures/ROC2.png}
+    \caption{Detection of non-null effects (ROC curves)} \label{fig:roc2}
+\end{subfigure}
+\caption{Results for Simulation 2 (constant precision, $\hat s_{jr}=1$).} 
+\medskip
+\small
+Note: In simulation 2, we set unit number $n = 10000$, and study number $R = 4$.
+Assume every observation has the same standard error $\hat s_{jr}=1$.
+That is, $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$.
+The 10000 units come from 4 patterns ($K=4$): 9100 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 300 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 300 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3, 4$ and $\beta_{jr}=0$ for $r = 1, 2$; 300 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$.
+\label{fig:sim2}
+\end{figure}
 
-\end{thebibliography}
+\end{appendix}
+
+\newpage
+
+\bibliographystyle{apalike}
+\bibliography{Motif.bib}
 

### 2018-03-08 code/function.R
git

@@ -1,5 +1,5 @@
 library("SQUAREM")
-library("REBayes")
+#library("REBayes")
 
 source('../code/lik.R')
 source('../code/set_data.R')

### 2018-03-08 analysis/Simulation_paper2.Rmd

--- 2018-01-29 analysis/Simulation_paper2.Rmd
@@
 ---
-title: "Simulation in Paper"
+title: "Simulation in Paper2"
 author: "Zhiwei Ma"
